### 2015-05-04 doc/paper.tex

% ****** Start of file aipsamp.tex ******
%
%   This file is part of the AIP files in the AIP distribution for REVTeX 4.
%   Version 4.1 of REVTeX, October 2009
%
%   Copyright (c) 2009 American Institute of Physics.
%
%   See the AIP README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex  aipsamp
%  2)  bibtex aipsamp
%  3)  latex  aipsamp
%  4)  latex  aipsamp
%
% Use this file as a source of example code for your aip document.
% Use the file aiptemplate.tex as a template for your document.
\documentclass[%
 aip,
 jmp,%
 amsmath,amssymb,
%preprint,%
 reprint,%
%author-year,%
%author-numerical,%
]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage[utf8]{inputenc}% bold math
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

\begin{document}

\preprint{AIP/123-QED}

\title[Sample title]{Selección de Instancias}% Force line breaks with \\

\author{Formica, Gabriel}
\author{Ponte, José Antonio}%
\affiliation{ 
Universidad Simón Bolívar%\\This line break forced with \textbackslash\textbackslash
}%

\author{C. Author}
 \homepage{http://www.Second.institution.edu/~Charlie.Author.}
\affiliation{%
Universidad Simón Bolívar%\\This line break forced% with \\
}%

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
\end{abstract}

\pacs{Valid PACS appear here}% PACS, the Physics and Astronomy
                             % Classification Scheme.
\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

\begin{quotation}
The ``lead paragraph'' is encapsulated with the \LaTeX\ 
\verb+quotation+ environment and is formatted as a single paragraph before the first section heading. 
(The \verb+quotation+ environment reverts to its usual meaning after the first sectioning command.) 
Note that numbered references are allowed in the lead paragraph.
%
The lead paragraph will only be found in an article being prepared for the journal \textit{Chaos}.
\end{quotation}

\section{\label{sec:level1}First-level heading}

This sample document demonstrates proper use of REV\TeX~4.1 (and
\LaTeXe) in manuscripts prepared for submission to AIP
journals. Further information can be found in the documentation included in the distribution or available at
\url{http://authors.aip.org} and in the documentation for 
REV\TeX~4.1 itself.

When commands are referred to in this example file, they are always
shown with their required arguments, using normal \TeX{} format. In
this format, \verb+#1+, \verb+#2+, etc. stand for required
author-supplied arguments to commands. For example, in
\verb+\section{#1}+ the \verb+#1+ stands for the title text of the
author's section heading, and in \verb+\title{#1}+ the \verb+#1+
stands for the title text of the paper.

Line breaks in section headings at all levels can be introduced using
\textbackslash\textbackslash. A blank input line tells \TeX\ that the
paragraph has ended.

\subsection{\label{sec:level2}Second-level heading: Formatting}

This file may be formatted in both the \texttt{preprint} (the default) and
\texttt{reprint} styles; the latter format may be used to 
mimic final journal output. Either format may be used for submission
purposes; however, for peer review and production, AIP will format the
article using the \texttt{preprint} class option. Hence, it is
essential that authors check that their manuscripts format acceptably
under \texttt{preprint}. Manuscripts submitted to AIP that do not
format correctly under the \texttt{preprint} option may be delayed in
both the editorial and production processes.

The \texttt{widetext} environment will make the text the width of the
full page, as on page~\pageref{eq:wideeq}. (Note the use the
\verb+\pageref{#1}+ to get the page number right automatically.) The
width-changing commands only take effect in \texttt{twocolumn}
formatting. It has no effect if \texttt{preprint} formatting is chosen
instead.

\subsubsection{\label{sec:level3}Third-level heading: Citations and Footnotes}

Citations in text refer to entries in the Bibliography;
they use the commands \verb+\cite{#1}+ or \verb+\onlinecite{#1}+. 
Because REV\TeX\ uses the \verb+natbib+ package of Patrick Daly, 
its entire repertoire of commands are available in your document;
see the \verb+natbib+ documentation for further details.
The argument of \verb+\cite+ is a comma-separated list of \emph{keys};
a key may consist of letters and numerals.

By default, citations are numerical; \cite{feyn54} author-year citations are an option. 
To give a textual citation, use \verb+\onlinecite{#1}+: (Refs.~\onlinecite{witten2001,epr,Bire82}). 
REV\TeX\ ``collapses'' lists of consecutive numerical citations when appropriate. 
REV\TeX\ provides the ability to properly punctuate textual citations in author-year style;
this facility works correctly with numerical citations only with \texttt{natbib}'s compress option turned off. 
To illustrate, we cite several together \cite{feyn54,witten2001,epr,Berman1983}, 
and once again (Refs.~\onlinecite{epr,feyn54,Bire82,Berman1983}). 
Note that, when numerical citations are used, the references were sorted into the same order they appear in the bibliography.

A reference within the bibliography is specified with a \verb+\bibitem{#1}+ command,
where the argument is the citation key mentioned above. 
\verb+\bibitem{#1}+ commands may be crafted by hand or, preferably,
generated by using Bib\TeX. 
The AIP styles for REV\TeX~4 include Bib\TeX\ style files
\verb+aipnum.bst+ and \verb+aipauth.bst+, appropriate for
numbered and author-year bibliographies,
respectively. 
REV\TeX~4 will automatically choose the style appropriate for 
the document's selected class options: the default is numerical, and
you obtain the author-year style by specifying a class option of \verb+author-year+.

This sample file demonstrates a simple use of Bib\TeX\ 
via a \verb+\bibliography+ command referencing the \verb+aipsamp.bib+ file.
Running Bib\TeX\ (in this case \texttt{bibtex
aipsamp}) after the first pass of \LaTeX\ produces the file
\verb+aipsamp.bbl+ which contains the automatically formatted
\verb+\bibitem+ commands (including extra markup information via
\verb+\bibinfo+ commands). If not using Bib\TeX, the
\verb+thebibiliography+ environment should be used instead.

\paragraph{Fourth-level heading is run in.}%
Footnotes are produced using the \verb+\footnote{#1}+ command. 
Numerical style citations put footnotes into the 
bibliography\footnote{Automatically placing footnotes into the bibliography requires using BibTeX to compile the bibliography.}.
Author-year and numerical author-year citation styles (each for its own reason) cannot use this method. 
Note: due to the method used to place footnotes in the bibliography, \emph{you
must re-run BibTeX every time you change any of your document's
footnotes}.

\section{Math and Equations}
Inline math may be typeset using the \verb+$+ delimiters. Bold math
symbols may be achieved using the \verb+bm+ package and the
\verb+\bm{#1}+ command it supplies. For instance, a bold $\alpha$ can
be typeset as \verb+$\bm{\alpha}$+ giving $\bm{\alpha}$. Fraktur and
Blackboard (or open face or double struck) characters should be
typeset using the \verb+\mathfrak{#1}+ and \verb+\mathbb{#1}+ commands
respectively. Both are supplied by the \texttt{amssymb} package. For
example, \verb+$\mathbb{R}$+ gives $\mathbb{R}$ and
\verb+$\mathfrak{G}$+ gives $\mathfrak{G}$

In \LaTeX\ there are many different ways to display equations, and a
few preferred ways are noted below. Displayed math will center by
default. Use the class option \verb+fleqn+ to flush equations left.

Below we have numbered single-line equations, the most common kind: 
\begin{eqnarray}
\chi_+(p)\alt{\bf [}2|{\bf p}|(|{\bf p}|+p_z){\bf ]}^{-1/2}
\left(
\begin{array}{c}
|{\bf p}|+p_z\\
px+ip_y
\end{array}\right)\;,
\\
\left\{%
 \openone234567890abc123\alpha\beta\gamma\delta1234556\alpha\beta
 \frac{1\sum^{a}_{b}}{A^2}%
\right\}%
\label{eq:one}.
\end{eqnarray}
Note the open one in Eq.~(\ref{eq:one}).

Not all numbered equations will fit within a narrow column this
way. The equation number will move down automatically if it cannot fit
on the same line with a one-line equation:
\begin{equation}
\left\{
 ab12345678abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
 \frac{1\sum^{a}_{b}}{A^2}%
\right\}.
\end{equation}

When the \verb+\label{#1}+ command is used [cf. input for
Eq.~(\ref{eq:one})], the equation can be referred to in text without
knowing the equation number that \TeX\ will assign to it. Just
use \verb+\ref{#1}+, where \verb+#1+ is the same name that used in
the \verb+\label{#1}+ command.

Unnumbered single-line equations can be typeset
using the \verb+\[+, \verb+\]+ format:
\[g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
q^+g^+g^+ \dots ~. \]

\subsection{Multiline equations}

Multiline equations are obtained by using the \verb+eqnarray+
environment.  Use the \verb+\nonumber+ command at the end of each line
to avoid assigning a number:
\begin{eqnarray}
{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
\delta_{\sigma_1,-\sigma_2}
(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
&&\times
[\epsilon_jl_i\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1),
\end{eqnarray}
\begin{eqnarray}
\sum \vert M^{\text{viol}}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
        (N^2-1)\nonumber \\
 & &\times \left( \sum_{i<j}\right)
  \sum_{\text{perm}}
 \frac{1}{S_{12}}
 \frac{1}{S_{12}}
 \sum_\tau c^f_\tau~.
\end{eqnarray}
\textbf{Note:} Do not use \verb+\label{#1}+ on a line of a multiline
equation if \verb+\nonumber+ is also used on that line. Incorrect
cross-referencing will result. Notice the use \verb+\text{#1}+ for
using a Roman font within a math environment.

To set a multiline equation without \emph{any} equation
numbers, use the \verb+\begin{eqnarray*}+,
\verb+\end{eqnarray*}+ format:
\begin{eqnarray*}
\sum \vert M^{\text{viol}}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
        (N^2-1)\\
 & &\times \left( \sum_{i<j}\right)
 \left(
  \sum_{\text{perm}}\frac{1}{S_{12}S_{23}S_{n1}}
 \right)
 \frac{1}{S_{12}}~.
\end{eqnarray*}
To obtain numbers not normally produced by the automatic numbering,
use the \verb+\tag{#1}+ command, where \verb+#1+ is the desired
equation number. For example, to get an equation number of
(\ref{eq:mynum}),
\begin{equation}
g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
q^+g^+g^+ \dots ~. \tag{2.6$'$}\label{eq:mynum}
\end{equation}

A few notes on \verb=\tag{#1}=. \verb+\tag{#1}+ requires
\texttt{amsmath}. The \verb+\tag{#1}+ must come before the
\verb+\label{#1}+, if any. The numbering set with \verb+\tag{#1}+ is
\textit{transparent} to the automatic numbering in REV\TeX{};
therefore, the number must be known ahead of time, and it must be
manually adjusted if other equations are added. \verb+\tag{#1}+ works
with both single-line and multiline equations. \verb+\tag{#1}+ should
only be used in exceptional case - do not use it to number all
equations in a paper.

Enclosing single-line and multiline equations in
\verb+\begin{subequations}+ and \verb+\end{subequations}+ will produce
a set of equations that are ``numbered'' with letters, as shown in
Eqs.~(\ref{subeq:1}) and (\ref{subeq:2}) below:
\begin{subequations}
\label{eq:whole}
\begin{equation}
\left\{
 abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
 \frac{1\sum^{a}_{b}}{A^2}
\right\},\label{subeq:1}
\end{equation}
\begin{eqnarray}
{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
&&\times
[\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1).\label{subeq:2}
\end{eqnarray}
\end{subequations}
Putting a \verb+\label{#1}+ command right after the
\verb+\begin{subequations}+, allows one to
reference all the equations in a subequations environment. For
example, the equations in the preceding subequations environment were
Eqs.~(\ref{eq:whole}).

\subsubsection{Wide equations}
The equation that follows is set in a wide format, i.e., it spans
across the full page. The wide format is reserved for long equations
that cannot be easily broken into four lines or less:
\begin{widetext}
\begin{equation}
{\cal R}^{(\text{d})}=
 g_{\sigma_2}^e
 \left(
   \frac{[\Gamma^Z(3,21)]_{\sigma_1}}{Q_{12}^2-M_W^2}
  +\frac{[\Gamma^Z(13,2)]_{\sigma_1}}{Q_{13}^2-M_W^2}
 \right)
 + x_WQ_e
 \left(
   \frac{[\Gamma^\gamma(3,21)]_{\sigma_1}}{Q_{12}^2-M_W^2}
  +\frac{[\Gamma^\gamma(13,2)]_{\sigma_1}}{Q_{13}^2-M_W^2}
 \right)\;. \label{eq:wideeq}
\end{equation}
\end{widetext}
This is typed to show the output is in wide format.
(Since there is no input line between \verb+\equation+ and
this paragraph, there is no paragraph indent for this paragraph.)

\end{document}
%
% ****** End of file aipsamp.tex ******

### 2015-05-04 doc/paper.tex
paper

@@
 \begin{abstract}
+  El problema de \textit{Selección de Instancias} (IS por sus siglas en inglés), busca
+  escoger un subconjunto de instancias de menor cardinalidad del conjunto original
+  que mantenga o mejore la clasificación de las muestras a ser usadas como un conjunto
+  de entrenamiento. En este articulo se considera la resolución de este problema
+  mediante varios algoritmos metaheurísticos, finalizando con los resultados 
+  experimentales y comparaciones entre ellos.
 \end{abstract}
@@
 
-\section{\label{sec:level1}First-level heading}
+\section{\label{sec:level1} Definición del Problema}
 
-This sample document demonstrates proper use of REV\TeX~4.1 (and
-\LaTeXe) in manuscripts prepared for submission to AIP
-journals. Further information can be found in the documentation included in the distribution or available at
-\url{http://authors.aip.org} and in the documentation for 
-REV\TeX~4.1 itself.
+Un conjunto de datos se define en función de un conjunto de clases $\Omega$ y
+un conjunto de $n$ observaciones $T$. Un problema de SI se define entonces como dicho 
+conjunto de instancias 
+$T = \{t_{1}, t_{2},..., t_{n}\}$, donde una instancia (observación) $t_{i}$ es
+una tupla $t_{i} = (v_{i,1}, v_{i,2},...,v_{i,m})$ de $m$ valores/mediciones
+(un punto en un espacio m-dimensional). Además, cada instancia $t_{i}$
+pertenece a una clase $w_{t_{i}}  \epsilon  \Omega$.
 
-When commands are referred to in this example file, they are always
-shown with their required arguments, using normal \TeX{} format. In
-this format, \verb+#1+, \verb+#2+, etc. stand for required
-author-supplied arguments to commands. For example, in
-\verb+\section{#1}+ the \verb+#1+ stands for the title text of the
-author's section heading, and in \verb+\title{#1}+ the \verb+#1+
-stands for the title text of the paper.
-
-Line breaks in section headings at all levels can be introduced using
-\textbackslash\textbackslash. A blank input line tells \TeX\ that the
-paragraph has ended.
+La idea es seleccionar un subconjuto de menor cardinalidad 
+$R \subseteq T$ tal que $R$ mantega o mejore la capacidad 
+de representación del conjunto $T$.
 
@@
 
-\subsubsection{\label{sec:level3}Third-level heading: Citations and Footnotes}
+\subsubsection{\label{sec:level3}Representación}
 
-Citations in text refer to entries in the Bibliography;
-they use the commands \verb+\cite{#1}+ or \verb+\onlinecite{#1}+. 
-Because REV\TeX\ uses the \verb+natbib+ package of Patrick Daly, 
-its entire repertoire of commands are available in your document;
-see the \verb+natbib+ documentation for further details.
-The argument of \verb+\cite+ is a comma-separated list of \emph{keys};
-a key may consist of letters and numerals.
-
-By default, citations are numerical; \cite{feyn54} author-year citations are an option. 
-To give a textual citation, use \verb+\onlinecite{#1}+: (Refs.~\onlinecite{witten2001,epr,Bire82}). 
-REV\TeX\ ``collapses'' lists of consecutive numerical citations when appropriate. 
-REV\TeX\ provides the ability to properly punctuate textual citations in author-year style;
-this facility works correctly with numerical citations only with \texttt{natbib}'s compress option turned off. 
-To illustrate, we cite several together \cite{feyn54,witten2001,epr,Berman1983}, 
-and once again (Refs.~\onlinecite{epr,feyn54,Bire82,Berman1983}). 
-Note that, when numerical citations are used, the references were sorted into the same order they appear in the bibliography.
-
-A reference within the bibliography is specified with a \verb+\bibitem{#1}+ command,
-where the argument is the citation key mentioned above. 
-\verb+\bibitem{#1}+ commands may be crafted by hand or, preferably,
-generated by using Bib\TeX. 
-The AIP styles for REV\TeX~4 include Bib\TeX\ style files
-\verb+aipnum.bst+ and \verb+aipauth.bst+, appropriate for
-numbered and author-year bibliographies,
-respectively. 
-REV\TeX~4 will automatically choose the style appropriate for 
-the document's selected class options: the default is numerical, and
-you obtain the author-year style by specifying a class option of \verb+author-year+.
-
-This sample file demonstrates a simple use of Bib\TeX\ 
-via a \verb+\bibliography+ command referencing the \verb+aipsamp.bib+ file.
-Running Bib\TeX\ (in this case \texttt{bibtex
-aipsamp}) after the first pass of \LaTeX\ produces the file
-\verb+aipsamp.bbl+ which contains the automatically formatted
-\verb+\bibitem+ commands (including extra markup information via
-\verb+\bibinfo+ commands). If not using Bib\TeX, the
-\verb+thebibiliography+ environment should be used instead.
-
-\paragraph{Fourth-level heading is run in.}%
-Footnotes are produced using the \verb+\footnote{#1}+ command. 
-Numerical style citations put footnotes into the 
-bibliography\footnote{Automatically placing footnotes into the bibliography requires using BibTeX to compile the bibliography.}.
-Author-year and numerical author-year citation styles (each for its own reason) cannot use this method. 
-Note: due to the method used to place footnotes in the bibliography, \emph{you
-must re-run BibTeX every time you change any of your document's
-footnotes}.
+Sea un conjunto inicial $T$, una solución al problema de SI está
+dado por un subconjuto de instancias $R$,
+que puede ser fácilmente representado por una cadena de bits de 
+tamaño $n = |T|$ donde cada bit tiene valor 1/0, indicando respectivamente,
+la pertenencia o no de una instancia $t_{i} \epsilon T$ en la solución.
 

### 2015-05-04 doc/paper.tex

@@
-% ****** Start of file aipsamp.tex ******
 %
-%   This file is part of the AIP files in the AIP distribution for REVTeX 4.
-%   Version 4.1 of REVTeX, October 2009
+% File acl2014.tex
 %
-%   Copyright (c) 2009 American Institute of Physics.
-%
-%   See the AIP README file for restrictions and more information.
-%
-% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
-% as well as the rest of the prerequisites for REVTeX 4.1
-%
-% It also requires running BibTeX. The commands are as follows:
-%
-%  1)  latex  aipsamp
-%  2)  bibtex aipsamp
-%  3)  latex  aipsamp
-%  4)  latex  aipsamp
-%
-% Use this file as a source of example code for your aip document.
-% Use the file aiptemplate.tex as a template for your document.
-\documentclass[%
- aip,
- jmp,%
- amsmath,amssymb,
-%preprint,%
- reprint,%
-%author-year,%
-%author-numerical,%
-]{revtex4-1}
+% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
+%%
+%% Based on the style files for ACL-2013, which were, in turn,
+%% Based on the style files for ACL-2012, which were, in turn,
+%% based on the style files for ACL-2011, which were, in turn, 
+%% based on the style files for ACL-2010, which were, in turn, 
+%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
+%% based on the style files for EACL-2009 and IJCNLP-2008...
 
-\usepackage{graphicx}% Include figure files
-\usepackage{dcolumn}% Align table columns on decimal point
-\usepackage[utf8]{inputenc}% bold math
-%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
-%\linenumbers\relax % Commence numbering lines
+%% Based on the style files for EACL 2006 by 
+%%e.agirre@ehu.es or Sergi.Balari@uab.es
+%% and that of ACL 08 by Joakim Nivre and Noah Smith
+
+\documentclass[11pt]{article}
+\usepackage{paper}
+\usepackage{times}
+\usepackage{url}
+\usepackage{latexsym}
+\usepackage[utf8]{inputenc}
+
+%\setlength\titlebox{5cm}
+
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
+
+\title{Selección de Instancias}
+
+\author{Formica, Gabriel\\
+  Universidad Simón Bolívar\\
+  Caracas, Venezuela\\
+  {gabrielformica93@gmail.com} \\\And
+  Ponte, José Antonio\\
+  Universidad Simón Bolívar\\
+  Caracas, Venezuela\\
+  {pontezambrano@gmail.com} \\}
+
+\date{}
 
 \begin{document}
-
-\preprint{AIP/123-QED}
-
-\title[Sample title]{Selección de Instancias}% Force line breaks with \\
-
-\author{Formica, Gabriel}
-\author{Ponte, José Antonio}%
-\affiliation{ 
-Universidad Simón Bolívar%\\This line break forced with \textbackslash\textbackslash
-}%
-
-\author{C. Author}
- \homepage{http://www.Second.institution.edu/~Charlie.Author.}
-\affiliation{%
-Universidad Simón Bolívar%\\This line break forced% with \\
-}%
-
-\date{\today}% It is always \today, today,
-             %  but any date may be explicitly specified
-
+\maketitle
 \begin{abstract}
@@
 
-\pacs{Valid PACS appear here}% PACS, the Physics and Astronomy
-                             % Classification Scheme.
-\keywords{Suggested keywords}%Use showkeys class option if keyword
-                              %display desired
-\maketitle
-
-\begin{quotation}
-The ``lead paragraph'' is encapsulated with the \LaTeX\ 
-\verb+quotation+ environment and is formatted as a single paragraph before the first section heading. 
-(The \verb+quotation+ environment reverts to its usual meaning after the first sectioning command.) 
-Note that numbered references are allowed in the lead paragraph.
-%
-The lead paragraph will only be found in an article being prepared for the journal \textit{Chaos}.
-\end{quotation}
-
-\section{\label{sec:level1} Definición del Problema}
+\section{Definición del Problema}
 
@@
 
-\subsection{\label{sec:level2}Second-level heading: Formatting}
+\section{Descripción de la solución}
 
-This file may be formatted in both the \texttt{preprint} (the default) and
-\texttt{reprint} styles; the latter format may be used to 
-mimic final journal output. Either format may be used for submission
-purposes; however, for peer review and production, AIP will format the
-article using the \texttt{preprint} class option. Hence, it is
-essential that authors check that their manuscripts format acceptably
-under \texttt{preprint}. Manuscripts submitted to AIP that do not
-format correctly under the \texttt{preprint} option may be delayed in
-both the editorial and production processes.
+En esta sección se muestra al lector el proceso de construcción de una 
+solución. Especificamente, se muestra cómo se representa una solución,
+cómo se generan las soluciones iniciales, y finalmente, se incluye 
+el pseudocódigo de la metaheurística implementada.
 
-The \texttt{widetext} environment will make the text the width of the
-full page, as on page~\pageref{eq:wideeq}. (Note the use the
-\verb+\pageref{#1}+ to get the page number right automatically.) The
-width-changing commands only take effect in \texttt{twocolumn}
-formatting. It has no effect if \texttt{preprint} formatting is chosen
-instead.
-
-\subsubsection{\label{sec:level3}Representación}
+\subsection{Representación}
 
@@
 
-\section{Math and Equations}
-Inline math may be typeset using the \verb+$+ delimiters. Bold math
-symbols may be achieved using the \verb+bm+ package and the
-\verb+\bm{#1}+ command it supplies. For instance, a bold $\alpha$ can
-be typeset as \verb+$\bm{\alpha}$+ giving $\bm{\alpha}$. Fraktur and
-Blackboard (or open face or double struck) characters should be
-typeset using the \verb+\mathfrak{#1}+ and \verb+\mathbb{#1}+ commands
-respectively. Both are supplied by the \texttt{amssymb} package. For
-example, \verb+$\mathbb{R}$+ gives $\mathbb{R}$ and
-\verb+$\mathfrak{G}$+ gives $\mathfrak{G}$
+\subsection{Generación de soluciones iniciales}
 
-In \LaTeX\ there are many different ways to display equations, and a
-few preferred ways are noted below. Displayed math will center by
-default. Use the class option \verb+fleqn+ to flush equations left.
+Siendo una representación una cadena de bits, para generar una solución
+inicial seguimos una estrategia muy sencilla: cada bit tiene una probabilidad $\rho$
+de estar en la solución. Si bien en la literatura, lo común es
+que $\rho$ sea igual a $0.5$, de manera de comenzar con un solución
+de $50\%$, en nuestro caso hemos hecho $\rho = 0.1$. Esto nos genera
+solucción con una reducción del $90\%$.
 
-Below we have numbered single-line equations, the most common kind: 
-\begin{eqnarray}
-\chi_+(p)\alt{\bf [}2|{\bf p}|(|{\bf p}|+p_z){\bf ]}^{-1/2}
-\left(
-\begin{array}{c}
-|{\bf p}|+p_z\\
-px+ip_y
-\end{array}\right)\;,
-\\
-\left\{%
- \openone234567890abc123\alpha\beta\gamma\delta1234556\alpha\beta
- \frac{1\sum^{a}_{b}}{A^2}%
-\right\}%
-\label{eq:one}.
-\end{eqnarray}
-Note the open one in Eq.~(\ref{eq:one}).
+\subsection{Format of Electronic Manuscript}
+\label{sect:pdf}
 
-Not all numbered equations will fit within a narrow column this
-way. The equation number will move down automatically if it cannot fit
-on the same line with a one-line equation:
-\begin{equation}
-\left\{
- ab12345678abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
- \frac{1\sum^{a}_{b}}{A^2}%
-\right\}.
-\end{equation}
+For the production of the electronic manuscript you must use Adobe's
+Portable Document Format (PDF). PDF files are usually produced from
+\LaTeX\ using the \textit{pdflatex} command. If your version of
+\LaTeX\ produces Postscript files, you can convert these into PDF
+using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
+Adobe Distiller to generate PDF.
 
-When the \verb+\label{#1}+ command is used [cf. input for
-Eq.~(\ref{eq:one})], the equation can be referred to in text without
-knowing the equation number that \TeX\ will assign to it. Just
-use \verb+\ref{#1}+, where \verb+#1+ is the same name that used in
-the \verb+\label{#1}+ command.
+Please make sure that your PDF file includes all the necessary fonts
+(especially tree diagrams, symbols, and fonts with Asian
+characters). When you print or create the PDF file, there is usually
+an option in your printer setup to include none, all or just
+non-standard fonts.  Please make sure that you select the option of
+including ALL the fonts. \textbf{Before sending it, test your PDF by
+  printing it from a computer different from the one where it was
+  created.} Moreover, some word processors may generate very large PDF
+files, where each page is rendered as an image. Such images may
+reproduce poorly. In this case, try alternative ways to obtain the
+PDF. One way on some systems is to install a driver for a postscript
+printer, send your document to the printer specifying ``Output to a
+file'', then convert the file to PDF.
 
-Unnumbered single-line equations can be typeset
-using the \verb+\[+, \verb+\]+ format:
-\[g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
-q^+g^+g^+ \dots ~. \]
+It is of utmost importance to specify the \textbf{A4 format} (21 cm
+x 29.7 cm) when formatting the paper. When working with
+{\tt dvips}, for instance, one should specify {\tt -t a4}.
 
-\subsection{Multiline equations}
+Print-outs of the PDF file on A4 paper should be identical to the
+hardcopy version. If you cannot meet the above requirements about the
+production of your electronic submission, please contact the
+publication chairs as soon as possible.
 
-Multiline equations are obtained by using the \verb+eqnarray+
-environment.  Use the \verb+\nonumber+ command at the end of each line
-to avoid assigning a number:
-\begin{eqnarray}
-{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
-\delta_{\sigma_1,-\sigma_2}
-(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
-&&\times
-[\epsilon_jl_i\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1),
-\end{eqnarray}
-\begin{eqnarray}
-\sum \vert M^{\text{viol}}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
-        (N^2-1)\nonumber \\
- & &\times \left( \sum_{i<j}\right)
-  \sum_{\text{perm}}
- \frac{1}{S_{12}}
- \frac{1}{S_{12}}
- \sum_\tau c^f_\tau~.
-\end{eqnarray}
-\textbf{Note:} Do not use \verb+\label{#1}+ on a line of a multiline
-equation if \verb+\nonumber+ is also used on that line. Incorrect
-cross-referencing will result. Notice the use \verb+\text{#1}+ for
-using a Roman font within a math environment.
+\subsection{Layout}
+\label{ssec:layout}
 
-To set a multiline equation without \emph{any} equation
-numbers, use the \verb+\begin{eqnarray*}+,
-\verb+\end{eqnarray*}+ format:
-\begin{eqnarray*}
-\sum \vert M^{\text{viol}}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
-        (N^2-1)\\
- & &\times \left( \sum_{i<j}\right)
- \left(
-  \sum_{\text{perm}}\frac{1}{S_{12}S_{23}S_{n1}}
- \right)
- \frac{1}{S_{12}}~.
-\end{eqnarray*}
-To obtain numbers not normally produced by the automatic numbering,
-use the \verb+\tag{#1}+ command, where \verb+#1+ is the desired
-equation number. For example, to get an equation number of
-(\ref{eq:mynum}),
-\begin{equation}
-g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
-q^+g^+g^+ \dots ~. \tag{2.6$'$}\label{eq:mynum}
-\end{equation}
+Format manuscripts two columns to a page, in the manner these
+instructions are formatted. The exact dimensions for a page on A4
+paper are:
 
-A few notes on \verb=\tag{#1}=. \verb+\tag{#1}+ requires
-\texttt{amsmath}. The \verb+\tag{#1}+ must come before the
-\verb+\label{#1}+, if any. The numbering set with \verb+\tag{#1}+ is
-\textit{transparent} to the automatic numbering in REV\TeX{};
-therefore, the number must be known ahead of time, and it must be
-manually adjusted if other equations are added. \verb+\tag{#1}+ works
-with both single-line and multiline equations. \verb+\tag{#1}+ should
-only be used in exceptional case - do not use it to number all
-equations in a paper.
+\begin{itemize}
+\item Left and right margins: 2.5 cm
+\item Top margin: 2.5 cm
+\item Bottom margin: 2.5 cm
+\item Column width: 7.7 cm
+\item Column height: 24.7 cm
+\item Gap between columns: 0.6 cm
+\end{itemize}
 
-Enclosing single-line and multiline equations in
-\verb+\begin{subequations}+ and \verb+\end{subequations}+ will produce
-a set of equations that are ``numbered'' with letters, as shown in
-Eqs.~(\ref{subeq:1}) and (\ref{subeq:2}) below:
-\begin{subequations}
-\label{eq:whole}
-\begin{equation}
-\left\{
- abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
- \frac{1\sum^{a}_{b}}{A^2}
-\right\},\label{subeq:1}
-\end{equation}
-\begin{eqnarray}
-{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
-(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
-&&\times
-[\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1).\label{subeq:2}
-\end{eqnarray}
-\end{subequations}
-Putting a \verb+\label{#1}+ command right after the
-\verb+\begin{subequations}+, allows one to
-reference all the equations in a subequations environment. For
-example, the equations in the preceding subequations environment were
-Eqs.~(\ref{eq:whole}).
+\noindent Papers should not be submitted on any other paper size.
+ If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chairs above as soon as possible.
 
-\subsubsection{Wide equations}
-The equation that follows is set in a wide format, i.e., it spans
-across the full page. The wide format is reserved for long equations
-that cannot be easily broken into four lines or less:
-\begin{widetext}
-\begin{equation}
-{\cal R}^{(\text{d})}=
- g_{\sigma_2}^e
- \left(
-   \frac{[\Gamma^Z(3,21)]_{\sigma_1}}{Q_{12}^2-M_W^2}
-  +\frac{[\Gamma^Z(13,2)]_{\sigma_1}}{Q_{13}^2-M_W^2}
- \right)
- + x_WQ_e
- \left(
-   \frac{[\Gamma^\gamma(3,21)]_{\sigma_1}}{Q_{12}^2-M_W^2}
-  +\frac{[\Gamma^\gamma(13,2)]_{\sigma_1}}{Q_{13}^2-M_W^2}
- \right)\;. \label{eq:wideeq}
-\end{equation}
-\end{widetext}
-This is typed to show the output is in wide format.
-(Since there is no input line between \verb+\equation+ and
-this paragraph, there is no paragraph indent for this paragraph.)
+\subsection{Fonts}
+
+For reasons of uniformity, Adobe's {\bf Times Roman} font should be
+used. In \LaTeX2e{} this is accomplished by putting
+
+\begin{quote}
+\begin{verbatim}
+\usepackage{times}
+\usepackage{latexsym}
+\end{verbatim}
+\end{quote}
+in the preamble. If Times Roman is unavailable, use {\bf Computer
+  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
+  10\% less dense than Adobe's Times Roman font.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|rl|}
+\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
+paper title & 15 pt & bold \\
+author names & 12 pt & bold \\
+author affiliation & 12 pt & \\
+the word ``Abstract'' & 12 pt & bold \\
+section titles & 12 pt & bold \\
+document text & 11 pt  &\\
+captions & 11 pt & \\
+abstract text & 10 pt & \\
+bibliography & 10 pt & \\
+footnotes & 9 pt & \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{font-table} Font guide. }
+\end{table}
+
+\subsection{The First Page}
+\label{ssec:first}
+
+Center the title, author's name(s) and affiliation(s) across both
+columns. Do not use footnotes for affiliations. Do not include the
+paper ID number assigned during the submission process. Use the
+two-column format only when you begin the abstract.
+
+{\bf Title}: Place the title centered at the top of the first page, in
+a 15-point bold font. (For a complete guide to font sizes and styles,
+see Table~\ref{font-table}) Long titles should be typed on two lines
+without a blank line intervening. Approximately, put the title at 2.5
+cm from the top of the page, followed by a blank line, then the
+author's names(s), and the affiliation on the following line. Do not
+use only initials for given names (middle initials are allowed). Do
+not format surnames in all capitals (e.g., use ``Schlangen'' not
+``SCHLANGEN'').  Do not format title and section headings in all
+capitals as well except for proper names (such as ``BLEU'') that are
+conventionally in all capitals.  The affiliation should contain the
+author's complete address, and if possible, an electronic mail
+address. Start the body of the first page 7.5 cm from the top of the
+page.
+
+The title, author names and addresses should be completely identical
+to those entered to the electronical paper submission website in order
+to maintain the consistency of author information among all
+publications of the conference. If they are different, the publication
+chairs may resolve the difference without consulting with you; so it
+is in your own interest to double-check that the information is
+consistent.
+
+{\bf Abstract}: Type the abstract at the beginning of the first
+column. The width of the abstract text should be smaller than the
+width of the columns for the text in the body of the paper by about
+0.6 cm on each side. Center the word {\bf Abstract} in a 12 point bold
+font above the body of the abstract. The abstract should be a concise
+summary of the general thesis and conclusions of the paper. It should
+be no longer than 200 words. The abstract text should be in 10 point font.
+
+{\bf Text}: Begin typing the main body of the text immediately after
+the abstract, observing the two-column format as shown in 
+the present document. Do not include page numbers.
+
+{\bf Indent} when starting a new paragraph. Use 11 points for text and 
+subsection headings, 12 points for section headings and 15 points for
+the title.
+
+\subsection{Sections}
+
+{\bf Headings}: Type and label section and subsection headings in the
+style shown on the present document.  Use numbered sections (Arabic
+numerals) in order to facilitate cross references. Number subsections
+with the section number and the subsection number separated by a dot,
+in Arabic numerals. Do not number subsubsections.
+
+{\bf Citations}: Citations within the text appear in parentheses
+as~\cite{Gusfield:97} or, if the author's name appears in the text
+itself, as Gusfield~\shortcite{Gusfield:97}.  Append lowercase letters
+to the year in cases of ambiguity.  Treat double authors as
+in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
+authors are involved. Collapse multiple citations as
+in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
+as sentence constituents. We suggest that instead of
+\begin{quote}
+  ``\cite{Gusfield:97} showed that ...''
+\end{quote}
+you use
+\begin{quote}
+``Gusfield \shortcite{Gusfield:97}   showed that ...''
+\end{quote}
+
+If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
+can use the command \verb|\newcite| to get ``author (year)'' citations.
+
+As reviewing will be double-blind, the submitted version of the papers
+should not include the authors' names and affiliations. Furthermore,
+self-references that reveal the author's identity, e.g.,
+\begin{quote}
+``We previously showed \cite{Gusfield:97} ...''  
+\end{quote}
+should be avoided. Instead, use citations such as 
+\begin{quote}
+``Gusfield \shortcite{Gusfield:97}
+previously showed ... ''
+\end{quote}
+
+\textbf{Please do not use anonymous citations} and do not include
+acknowledgements when submitting your papers. Papers that do not
+conform to these requirements may be rejected without review.
+
+\textbf{References}: Gather the full set of references together under
+the heading {\bf References}; place the section before any Appendices,
+unless they contain references. Arrange the references alphabetically
+by first author, rather than by order of occurrence in the text.
+Provide as complete a citation as possible, using a consistent format,
+such as the one for {\em Computational Linguistics\/} or the one in the 
+{\em Publication Manual of the American 
+Psychological Association\/}~\cite{APA:83}.  Use of full names for
+authors rather than initials is preferred.  A list of abbreviations
+for common computer science journals can be found in the ACM 
+{\em Computing Reviews\/}~\cite{ACM:83}.
+
+The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
+American Psychological Association format, allowing regular citations, 
+short citations and multiple citations as described above.
+
+{\bf Appendices}: Appendices, if any, directly follow the text and the
+references (but see above).  Letter them in sequence and provide an
+informative title: {\bf Appendix A. Title of Appendix}.
+
+\subsection{Footnotes}
+
+{\bf Footnotes}: Put footnotes at the bottom of the page and use 9
+points text. They may be numbered or referred to by asterisks or other
+symbols.\footnote{This is how a footnote should appear.} Footnotes
+should be separated from the text by a line.\footnote{Note the line
+separating the footnotes from the text.}
+
+\subsection{Graphics}
+
+{\bf Illustrations}: Place figures, tables, and photographs in the
+paper near where they are first discussed, rather than at the end, if
+possible.  Wide illustrations may run across both columns.  Color
+illustrations are discouraged, unless you have verified that  
+they will be understandable when printed in black ink.
+
+{\bf Captions}: Provide a caption for every illustration; number each one
+sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
+Caption of the Table.''  Type the captions of the figures and 
+tables below the body, using 11 point text.
+
+\section{XML conversion and supported \LaTeX\ packages}
+
+ACL 2014 innovates over earlier years in that we will attempt to
+automatically convert your \LaTeX\ source files to machine-readable
+XML with semantic markup. This will facilitate future research that
+uses the ACL proceedings themselves as a corpus.
+
+We encourage you to submit a ZIP file of your \LaTeX\ sources along
+with the camera-ready version of your paper. We will then convert them
+to XML automatically, using the LaTeXML tool
+(\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
+a number of \LaTeX\ packages, including the ACL 2014 stylefile. These
+bindings allow LaTeXML to render the commands from these packages
+correctly in XML. For best results, we encourage you to use the
+packages that are officially supported by LaTeXML, listed at
+\url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}
+
+\section{Translation of non-English Terms}
+
+It is also advised to supplement non-English characters and terms
+with appropriate transliterations and/or translations
+since not all readers understand all such characters and terms.
+Inline transliteration or translation can be represented in
+the order of: original-form transliteration ``translation''.
+
+\section{Length of Submission}
+\label{sec:length}
+
+Long papers may consist of up to 9 pages of content, plus two extra
+pages for references. Short papers may consist of up to 5 pages of
+content, plus two extra pages for references.  Papers that do not
+conform to the specified length and formatting requirements may be
+rejected without review.
+
+\section*{Acknowledgments}
+
+The acknowledgments should go immediately before the references.  Do
+not number the acknowledgments section. Do not include this section
+when submitting your paper for review.
+
+% include your own bib file like this:
+%\bibliographystyle{acl}
+%\bibliography{acl2014}
+
+\begin{thebibliography}{}
+
+\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
+Alfred~V. Aho and Jeffrey~D. Ullman.
+\newblock 1972.
+\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
+\newblock Prentice-{Hall}, Englewood Cliffs, NJ.
+
+\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
+{American Psychological Association}.
+\newblock 1983.
+\newblock {\em Publications Manual}.
+\newblock American Psychological Association, Washington, DC.
+
+\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
+{Association for Computing Machinery}.
+\newblock 1983.
+\newblock {\em Computing Reviews}, 24(11):503--512.
+
+\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
+Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
+\newblock 1981.
+\newblock Alternation.
+\newblock {\em Journal of the Association for Computing Machinery},
+  28(1):114--133.
+
+\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
+Dan Gusfield.
+\newblock 1997.
+\newblock {\em Algorithms on Strings, Trees and Sequences}.
+\newblock Cambridge University Press, Cambridge, UK.
+
+\end{thebibliography}
 
 \end{document}
-%
-% ****** End of file aipsamp.tex ******

### 2015-05-04 doc/paper.tex
working

@@
 \usepackage{latexsym}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage[utf8]{inputenc}
@@
 solucción con una reducción del $90\%$.
+
+\section{Función de evaluación}
+En [referencia a cano] se describe una función de evaluación que considera la tasa de instancias clasificadas correctamente tomando en cuenta el conjunto reducido como conjunto de entrenamiento y una tasa de reducción del conjunto original.
+Se hace uso de una constante $\alpha$ donde $\alpha \in [0,1]$ para darle ponderación a ambos parámetros.
+La fórmula usada para medir la calidad de una solución es:
+
+\begin{center}
+    $ evaluar(R) = \alpha \times tasa\_clas + (1 - \alpha) \times porcen\_redc $
+\end{center}
+
+El objetivo de un algoritmo es maximizar esta función, entonces si $evaluar(A) < evaluar(B)$, se tiene que $B$ es mejor solución que $A$.
+Para los experimentos se usó la tasa sugerida por [referencia a cano] de $\alpha = 0.5$ dándole la misma ponderación a $tasa\_clas$ y a $porcen\_redc$.
+
+\section{Algoritmos}
+
+\subsection{Búsqueda local}
+    Para la búsqueda local se hace uso de un algoritmo inspirado en \emph{hill-climbing} el cual toma una solución inicial y tiene como objetivo mejorarla haciendo uso de un operador $op: S \to S$.
+
+    \begin{algorithmic}
+        \Function{BusquedaLocal}{$T: \text{Problema}$}
+            \State $S \gets \text{solucionAleatoria()}$
+            \While{\text{$\lnot condicionDeParada$}}
+                \State $R \gets S$
+                \State $R \gets operador(R)$
+                \If{\Call{calidad}{R,T} $<$ \Call{calidad}{S,T}}
+                    \State $S \gets R$
+                \EndIf
+            \EndWhile
+            \State return $S$
+        \EndFunction
+    \end{algorithmic}
+
+\subsection{1NN}
+    Para el cálculo de la calidad de una solución, se hace uso del clasificador $1NN$ que clasifica un conjunto usando un conjunto definido como entrenamiento.
+Sobre los resultados obtenidos de $1NN$ se corre una función de evaluación que asinga un valor a la solución.
+
+    \begin{algorithmic}
+        \Function{Calidad}{$S: \text{Solución}$, $T: \text{Problema}$}
+            \State \text{Problema entrenamiento}
+            \State \text{Problema resultado}
+            \State $resultado \gets \Call{1NN}{entrenamiento, T}$
+            \State retorna eval(resultado)
+        \EndFunction
+    \end{algorithmic}
 

### 2015-05-04 doc/paper.tex
working paper

--- 2015-05-04 doc/paper.tex
@@
 que $\rho$ sea igual a $0.5$, de manera de comenzar con un solución
-de $50\%$, en nuestro caso hemos hecho $\rho = 0.1$. Esto nos genera
-solucción con una reducción del $90\%$.
+de $50\%$, en nuestro caso hemos hecho varias ``corridas'' con distintos valores
+de $\rho$, generando soluciones iniciales de distintos tamaños, de manera de poder
+evaluar el impacto de cada una de estas reducciones.
 

### 2015-05-04 doc/paper.tex
working

--- 2015-05-04 doc/paper.tex
@@
 \usepackage{times}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{url}
 \usepackage{latexsym}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
 \usepackage[utf8]{inputenc}
@@
 
+~\
+
 \begin{center}
+    {\fontsize{10}{10}\selectfont
     $ evaluar(R) = \alpha \times tasa\_clas + (1 - \alpha) \times porcen\_redc $
+    }
 \end{center}
+
+~\
 
@@
 \subsection{Búsqueda local}
-    Para la búsqueda local se hace uso de un algoritmo inspirado en \emph{hill-climbing} el cual toma una solución inicial y tiene como objetivo mejorarla haciendo uso de un operador $op: S \to S$.
+    Para la búsqueda local se hace uso de un algoritmo inspirado en \emph{hill-climbing} el cual toma una solución inicial y tiene como objetivo mejorarla haciendo uso de un operador $op: S \to S$. \\
 
+    {\fontsize{10}{10}\selectfont
     \begin{algorithmic}
-        \Function{BusquedaLocal}{$T: \text{Problema}$}
-            \State $S \gets \text{solucionAleatoria()}$
-            \While{\text{$\lnot condicionDeParada$}}
+        \Function{BusquedaLocal}{$T: Problema$}
+            \State $S \gets solucionAleatoria()$
+            \While{$\lnot condicionDeParada$}
                 \State $R \gets S$
@@
     \end{algorithmic}
+    }
 
@@
 Sobre los resultados obtenidos de $1NN$ se corre una función de evaluación que asinga un valor a la solución.
+\\
 
+    {\fontsize{10}{10}\selectfont
     \begin{algorithmic}
-        \Function{Calidad}{$S: \text{Solución}$, $T: \text{Problema}$}
-            \State \text{Problema entrenamiento}
-            \State \text{Problema resultado}
+        \Function{Calidad}{$S: Solucion$, $T: Problema$}
+            \State $Problema$  $entrenamiento$
+            \State $Problema$  $resultado$
             \State $resultado \gets \Call{1NN}{entrenamiento, T}$
-            \State retorna eval(resultado)
+            \State $retorna$ $eval(resultado)$
         \EndFunction
     \end{algorithmic}
+    }
 

### 2015-05-04 doc/paper.tex
working paper

@@
 
-\subsection{Format of Electronic Manuscript}
-\label{sect:pdf}
+\subsection{Operador}
+Ante la necesidad de movernos en el
+espacio de busqueda, surge la idea de un operador 
+que modifique de manera sutil (y rápida) una solución
+dada. 
+En nuestro caso, nuestro operador selecciona al
+azar un bit ``prendido'' y lo ``apaga'', que se traduce
+en ``sacar'' una instancia del conjunto de instancias solución.
 
-For the production of the electronic manuscript you must use Adobe's
-Portable Document Format (PDF). PDF files are usually produced from
-\LaTeX\ using the \textit{pdflatex} command. If your version of
-\LaTeX\ produces Postscript files, you can convert these into PDF
-using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
-Adobe Distiller to generate PDF.
-
-Please make sure that your PDF file includes all the necessary fonts
-(especially tree diagrams, symbols, and fonts with Asian
-characters). When you print or create the PDF file, there is usually
-an option in your printer setup to include none, all or just
-non-standard fonts.  Please make sure that you select the option of
-including ALL the fonts. \textbf{Before sending it, test your PDF by
-  printing it from a computer different from the one where it was
-  created.} Moreover, some word processors may generate very large PDF
-files, where each page is rendered as an image. Such images may
-reproduce poorly. In this case, try alternative ways to obtain the
-PDF. One way on some systems is to install a driver for a postscript
-printer, send your document to the printer specifying ``Output to a
-file'', then convert the file to PDF.
-
-It is of utmost importance to specify the \textbf{A4 format} (21 cm
-x 29.7 cm) when formatting the paper. When working with
-{\tt dvips}, for instance, one should specify {\tt -t a4}.
-
-Print-outs of the PDF file on A4 paper should be identical to the
-hardcopy version. If you cannot meet the above requirements about the
-production of your electronic submission, please contact the
-publication chairs as soon as possible.
+De esta manera, se consiguen soluciones cada vez más pequeñas y una tasa
+de aciertos $tasa\_clas$ cada vez más precisas: esto viene de si tenemos 
+$x$ aciertos y $w$ instancias que no están en la solución, aumetar el tamaño
+de una solución conllevaría a una posible $tasa\_clas = x / (w - 1) $,
+que supondría una mayor $tasa\_clas$ y por lo tanto una mejor solución. Sin embargo, 
+esto no puede estar más alejado de la realidad.
 

### 2015-05-04 doc/paper.tex
* Added stop condition * worked on paper

@@
 \usepackage{algpseudocode}
+\usepackage{multirow}
 \usepackage{url}
@@
 \section{Función de evaluación}
-En [referencia a cano] se describe una función de evaluación que considera la tasa de instancias clasificadas correctamente tomando en cuenta el conjunto reducido como conjunto de entrenamiento y una tasa de reducción del conjunto original.
+En (Cano et.
+al 2003) se describe una función de evaluación que considera la tasa de instancias clasificadas correctamente tomando en cuenta el conjunto reducido como conjunto de entrenamiento y una tasa de reducción del conjunto original.
 Se hace uso de una constante $\alpha$ donde $\alpha \in [0,1]$ para darle ponderación a ambos parámetros.
@@
 El objetivo de un algoritmo es maximizar esta función, entonces si $evaluar(A) < evaluar(B)$, se tiene que $B$ es mejor solución que $A$.
-Para los experimentos se usó la tasa sugerida por [referencia a cano] de $\alpha = 0.5$ dándole la misma ponderación a $tasa\_clas$ y a $porcen\_redc$.
+Para los experimentos se usó la tasa sugerida por (Cano et.
+al 2003) de $\alpha = 0.5$ dándole la misma ponderación a $tasa\_clas$ y a $porcen\_redc$.
 
@@
 
+    ~\
+
+    Como condición de parada se establece que el mínimo de calidad de una solución sea $0.95$ y una cantidad determinada de iteraciones en las cuales la solución no cambie.
+
 \subsection{1NN}
@@
 
-\subsection{Operador}
-Ante la necesidad de movernos en el
-espacio de busqueda, surge la idea de un operador 
-que modifique de manera sutil (y rápida) una solución
-dada. 
-En nuestro caso, nuestro operador selecciona al
-azar un bit ``prendido'' y lo ``apaga'', que se traduce
-en ``sacar'' una instancia del conjunto de instancias solución.
+\section{Experimentos}
 
-De esta manera, se consiguen soluciones cada vez más pequeñas y una tasa
-de aciertos $tasa\_clas$ cada vez más precisas: esto viene de si tenemos 
-$x$ aciertos y $w$ instancias que no están en la solución, aumetar el tamaño
-de una solución conllevaría a una posible $tasa\_clas = x / (w - 1) $,
-que supondría una mayor $tasa\_clas$ y por lo tanto una mejor solución. Sin embargo, 
-esto no puede estar más alejado de la realidad.
+Se seleccionaron instancias de tamaño pequeño con menos de 1000 instancias y mediano con más de 1000 pero menos de 5000 instancias.
+Cada problema cuenta con atributos y clases en el dominio de los reales.
 
-\subsection{Layout}
-\label{ssec:layout}
+\subsection{Resultados}
 
-Format manuscripts two columns to a page, in the manner these
-instructions are formatted. The exact dimensions for a page on A4
-paper are:
-
-\begin{itemize}
-\item Left and right margins: 2.5 cm
-\item Top margin: 2.5 cm
-\item Bottom margin: 2.5 cm
-\item Column width: 7.7 cm
-\item Column height: 24.7 cm
-\item Gap between columns: 0.6 cm
-\end{itemize}
-
-\noindent Papers should not be submitted on any other paper size.
- If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chairs above as soon as possible.
-
-\subsection{Fonts}
-
-For reasons of uniformity, Adobe's {\bf Times Roman} font should be
-used. In \LaTeX2e{} this is accomplished by putting
-
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\end{quote}
-in the preamble. If Times Roman is unavailable, use {\bf Computer
-  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
-  10\% less dense than Adobe's Times Roman font.
+ Se seleccionan tres casos pequeños y se observa que el porcentaje de reducción es pequeño.
 
 \begin{table}[h]
-\begin{center}
-\begin{tabular}{|l|rl|}
-\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
-paper title & 15 pt & bold \\
-author names & 12 pt & bold \\
-author affiliation & 12 pt & \\
-the word ``Abstract'' & 12 pt & bold \\
-section titles & 12 pt & bold \\
-document text & 11 pt  &\\
-captions & 11 pt & \\
-abstract text & 10 pt & \\
-bibliography & 10 pt & \\
-footnotes & 9 pt & \\
+\begin{tabular}{ |l|l|l|l| }
     \hline
+    Caso & Instancias & Atributos & \% red. \\ \hline
+    Haberman & 306 & 4 & 13.7 \\ \hline
+    Heart & 270 & 14 & 46 \\ \hline
+    Fertility & 100 & 10 & 34 \\ \hline
 \end{tabular}
-\end{center}
-\caption{\label{font-table} Font guide. }
+\caption{Datos pequeños}
+\label{tabla:1}
 \end{table}
 
-\subsection{The First Page}
-\label{ssec:first}
+    ~\
 
-Center the title, author's name(s) and affiliation(s) across both
-columns. Do not use footnotes for affiliations. Do not include the
-paper ID number assigned during the submission process. Use the
-two-column format only when you begin the abstract.
+    Para los casos medianos, se observa que el porcentaje de reducción es mucho mayor
 
-{\bf Title}: Place the title centered at the top of the first page, in
-a 15-point bold font. (For a complete guide to font sizes and styles,
-see Table~\ref{font-table}) Long titles should be typed on two lines
-without a blank line intervening. Approximately, put the title at 2.5
-cm from the top of the page, followed by a blank line, then the
-author's names(s), and the affiliation on the following line. Do not
-use only initials for given names (middle initials are allowed). Do
-not format surnames in all capitals (e.g., use ``Schlangen'' not
-``SCHLANGEN'').  Do not format title and section headings in all
-capitals as well except for proper names (such as ``BLEU'') that are
-conventionally in all capitals.  The affiliation should contain the
-author's complete address, and if possible, an electronic mail
-address. Start the body of the first page 7.5 cm from the top of the
-page.
+    ~\
 
-The title, author names and addresses should be completely identical
-to those entered to the electronical paper submission website in order
-to maintain the consistency of author information among all
-publications of the conference. If they are different, the publication
-chairs may resolve the difference without consulting with you; so it
-is in your own interest to double-check that the information is
-consistent.
-
-{\bf Abstract}: Type the abstract at the beginning of the first
-column. The width of the abstract text should be smaller than the
-width of the columns for the text in the body of the paper by about
-0.6 cm on each side. Center the word {\bf Abstract} in a 12 point bold
-font above the body of the abstract. The abstract should be a concise
-summary of the general thesis and conclusions of the paper. It should
-be no longer than 200 words. The abstract text should be in 10 point font.
-
-{\bf Text}: Begin typing the main body of the text immediately after
-the abstract, observing the two-column format as shown in 
-the present document. Do not include page numbers.
-
-{\bf Indent} when starting a new paragraph. Use 11 points for text and 
-subsection headings, 12 points for section headings and 15 points for
-the title.
-
-\subsection{Sections}
-
-{\bf Headings}: Type and label section and subsection headings in the
-style shown on the present document.  Use numbered sections (Arabic
-numerals) in order to facilitate cross references. Number subsections
-with the section number and the subsection number separated by a dot,
-in Arabic numerals. Do not number subsubsections.
-
-{\bf Citations}: Citations within the text appear in parentheses
-as~\cite{Gusfield:97} or, if the author's name appears in the text
-itself, as Gusfield~\shortcite{Gusfield:97}.  Append lowercase letters
-to the year in cases of ambiguity.  Treat double authors as
-in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
-authors are involved. Collapse multiple citations as
-in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
-as sentence constituents. We suggest that instead of
-\begin{quote}
-  ``\cite{Gusfield:97} showed that ...''
-\end{quote}
-you use
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}   showed that ...''
-\end{quote}
-
-If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
-can use the command \verb|\newcite| to get ``author (year)'' citations.
-
-As reviewing will be double-blind, the submitted version of the papers
-should not include the authors' names and affiliations. Furthermore,
-self-references that reveal the author's identity, e.g.,
-\begin{quote}
-``We previously showed \cite{Gusfield:97} ...''  
-\end{quote}
-should be avoided. Instead, use citations such as 
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}
-previously showed ... ''
-\end{quote}
-
-\textbf{Please do not use anonymous citations} and do not include
-acknowledgements when submitting your papers. Papers that do not
-conform to these requirements may be rejected without review.
-
-\textbf{References}: Gather the full set of references together under
-the heading {\bf References}; place the section before any Appendices,
-unless they contain references. Arrange the references alphabetically
-by first author, rather than by order of occurrence in the text.
-Provide as complete a citation as possible, using a consistent format,
-such as the one for {\em Computational Linguistics\/} or the one in the 
-{\em Publication Manual of the American 
-Psychological Association\/}~\cite{APA:83}.  Use of full names for
-authors rather than initials is preferred.  A list of abbreviations
-for common computer science journals can be found in the ACM 
-{\em Computing Reviews\/}~\cite{ACM:83}.
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
-American Psychological Association format, allowing regular citations, 
-short citations and multiple citations as described above.
-
-{\bf Appendices}: Appendices, if any, directly follow the text and the
-references (but see above).  Letter them in sequence and provide an
-informative title: {\bf Appendix A. Title of Appendix}.
-
-\subsection{Footnotes}
-
-{\bf Footnotes}: Put footnotes at the bottom of the page and use 9
-points text. They may be numbered or referred to by asterisks or other
-symbols.\footnote{This is how a footnote should appear.} Footnotes
-should be separated from the text by a line.\footnote{Note the line
-separating the footnotes from the text.}
-
-\subsection{Graphics}
-
-{\bf Illustrations}: Place figures, tables, and photographs in the
-paper near where they are first discussed, rather than at the end, if
-possible.  Wide illustrations may run across both columns.  Color
-illustrations are discouraged, unless you have verified that  
-they will be understandable when printed in black ink.
-
-{\bf Captions}: Provide a caption for every illustration; number each one
-sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
-Caption of the Table.''  Type the captions of the figures and 
-tables below the body, using 11 point text.
-
-\section{XML conversion and supported \LaTeX\ packages}
-
-ACL 2014 innovates over earlier years in that we will attempt to
-automatically convert your \LaTeX\ source files to machine-readable
-XML with semantic markup. This will facilitate future research that
-uses the ACL proceedings themselves as a corpus.
-
-We encourage you to submit a ZIP file of your \LaTeX\ sources along
-with the camera-ready version of your paper. We will then convert them
-to XML automatically, using the LaTeXML tool
-(\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
-a number of \LaTeX\ packages, including the ACL 2014 stylefile. These
-bindings allow LaTeXML to render the commands from these packages
-correctly in XML. For best results, we encourage you to use the
-packages that are officially supported by LaTeXML, listed at
-\url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}
-
-\section{Translation of non-English Terms}
-
-It is also advised to supplement non-English characters and terms
-with appropriate transliterations and/or translations
-since not all readers understand all such characters and terms.
-Inline transliteration or translation can be represented in
-the order of: original-form transliteration ``translation''.
-
-\section{Length of Submission}
-\label{sec:length}
-
-Long papers may consist of up to 9 pages of content, plus two extra
-pages for references. Short papers may consist of up to 5 pages of
-content, plus two extra pages for references.  Papers that do not
-conform to the specified length and formatting requirements may be
-rejected without review.
-
-\section*{Acknowledgments}
-
-The acknowledgments should go immediately before the references.  Do
-not number the acknowledgments section. Do not include this section
-when submitting your paper for review.
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l| }
+    \hline
+    Caso & Instancias & Atributos & \% red. \\ \hline
+    Phishing & 2456 & 31 & 84 \\ \hline
+    Segment & 2310 & 20 & 87 \\ \hline
+    Bank & 1372 & 5 & 90 \\ \hline
+\end{tabular}
+\caption{Datos medianos}
+\label{tabla:2}
+\end{table}
 
@@
 
-\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
-Alfred~V. Aho and Jeffrey~D. Ullman.
-\newblock 1972.
-\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
-\newblock Prentice-{Hall}, Englewood Cliffs, NJ.
+\bibitem[\protect\citename{Cano et al}2003]{Cano:03}
+José Ramón Cano, Francisco Herrera, and Manuel Lozano.
+\newblock 2003.
+\newblock {\em Using evo-
+lutionary algorithms as instance selection for data reduction in kdd: an
+experimental study.}.
+\newblock Evolutionary Computation, IEEE Transactions.
 
-\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
-{American Psychological Association}.
-\newblock 1983.
-\newblock {\em Publications Manual}.
-\newblock American Psychological Association, Washington, DC.
-
-\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
-{Association for Computing Machinery}.
-\newblock 1983.
-\newblock {\em Computing Reviews}, 24(11):503--512.
-
-\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
-Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
-\newblock 1981.
-\newblock Alternation.
-\newblock {\em Journal of the Association for Computing Machinery},
-  28(1):114--133.
-
-\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
-Dan Gusfield.
-\newblock 1997.
-\newblock {\em Algorithms on Strings, Trees and Sequences}.
-\newblock Cambridge University Press, Cambridge, UK.
+\bibitem[\protect\citename{Toussaint}200]{Toussaint:02}
+Godfried Toussaing.
+\newblock 2002.
+\newblock {\em Proximity Graphs for Nearest Neighbor Decision Problem}.
+\newblock School of computer science. McGill University
 

### 2015-05-04 src/is.cpp

@@ -62,7 +62,8 @@ void IS::Solution::tweak() {
 void IS::Metaheuristic::optimize(const IS::Problem &T, IS::Solution &S) {
     S.setSize(T.size());
     S.generateRandom();
-    double q_max;
+    double q_max, old_q = -1;
+    int iter_old = 0;
     while (1) {
         IS::Solution R(S);
         R.tweak(); 
@@ -74,7 +75,13 @@ void IS::Metaheuristic::optimize(const IS::Problem &T, IS::Solution &S) {
             S.setBits(R.getBits());
         
         q_max = std::max(q1, q2);
-        if (q_max > 0.95) break;
+        if (old_q == q_max) {
+          iter_old ++;  
+        } else {
+            iter_old = 0;
+        }
+        if (q_max > 0.95 or iter_old == 1000) break;
+        old_q = q_max;
     }
     cout << "---> " << q_max << endl;
 }
@@ -102,15 +109,15 @@ double IS::Metaheuristic::quality(const IS::Problem &T,
     double clas_rate = (1.0 * count) / (1.0 * result.size());
     double perc_redc = (1.0 * result.size()) / (1.0 * T.size());
 
-    //cout << "Hubo un total de " << count << " aciertos" << endl;
-    //cout << "Result es de tamanio: " << result.size() << endl;
-    //cout << "La relación es: " << clas_rate << endl;
+    // cout << "Hubo un total de " << count << " aciertos" << endl;
+    // cout << "Result es de tamanio: " << result.size() << endl;
+    // cout << "La relación es: " << clas_rate << endl;
 
-    //cout << "El tamaño de T es: " << T.size() << endl;
-    //cout << "El porcentaje de reducción es: " << perc_redc << endl;
+    // cout << "El tamaño de T es: " << T.size() << endl;
+    // cout << "El porcentaje de reducción es: " << perc_redc << endl;
 
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
-    cout << "El fitness es : " << fitness << std::endl;
+    // cout << "El fitness es : " << fitness << std::endl;
     assert(fitness <= 1.0);
     return fitness;
 }

### 2015-05-28 doc/paper.tex

@@
 
+\subsection{Operador}
+Ante la necesidad de movernos en el
+espacio de busqueda, surge la idea de un operador 
+que modifique de manera sutil (y rápida) una solución
+dada. 
+En nuestro caso, nuestro operador selecciona al
+azar un bit ``prendido'' y lo ``apaga'', que se traduce
+en ``sacar'' una instancia del conjunto de instancias solución.
+
+De esta manera, se consiguen soluciones cada vez más pequeñas y una tasa
+de aciertos $tasa\_clas$ cada vez más precisas: esto viene de si tenemos 
+$x$ aciertos y $w$ instancias que no están en la solución, aumetar el tamaño
+de una solución conllevaría a una posible $tasa\_clas = x / (w - 1) $,
+que supondría una mayor $tasa\_clas$ y por lo tanto una mejor solución. Sin embargo, 
+esto no puede estar más alejado de la realidad.
+
 \section{Experimentos}

### 2015-05-30 src/is.cpp
* refactor * simulated annealing

@@ -9,38 +9,44 @@
 
 const double epsilon = 0.00001;
 
-double IS::Metaheuristic::find_nearest(const Problem &p, const Instance &a) {
-    double imin = 1.0 * MAX;
-    double category;
-    for (Problem::const_iterator it = p.begin(); it < p.end(); ++it) {
-        double dist = calc_distance(*it, a);
-        if (dist < epsilon) {
-            category = (*it).getCategory();
-            break;
-        }
-        if (dist < imin) {
-            imin = dist;                  
-            category = (*it).getCategory();
-        }
-    }
-    return category;
-}
+/*
+ *double IS::Metaheuristic::find_nearest(const Problem &p, const Instance &a) {
+ *    double imin = 1.0 * MAX;
+ *    double category;
+ *    for (Problem::const_iterator it = p.begin(); it < p.end(); ++it) {
+ *        double dist = calc_distance(*it, a);
+ *        if (dist < epsilon) {
+ *            category = (*it).getCategory();
+ *            break;
+ *        }
+ *        if (dist < imin) {
+ *            imin = dist;                  
+ *            category = (*it).getCategory();
+ *        }
+ *    }
+ *    return category;
+ *}
+ */
 
-void IS::Metaheuristic::oneNN(const Problem &training, 
-                              const Problem &result, 
-                              std::vector<double> &category) {
+/*
+ *void IS::Metaheuristic::oneNN(const Problem &training, 
+ *                              const Problem &result, 
+ *                              std::vector<double> &category) {
+ *
+ *    for (int i = 0; i < result.size(); i++)
+ *        category[i] = find_nearest(training, result[i]);
+ *}
+ */
 
-    for (int i = 0; i < result.size(); i++)
-        category[i] = find_nearest(training, result[i]);
-}
-
-double IS::Metaheuristic::calc_distance(const Instance &a, const Instance &b) {
-    double dist = 0.0;
-    assert(a.size() == b.size());
-    for (int i = 0; i < a.size(); ++i)
-        dist += (a[i] - b[i]) * (a[i] - b[i]);
-    return sqrt(dist);
-}
+/*
+ *double IS::Metaheuristic::calc_distance(const Instance &a, const Instance &b) {
+ *    double dist = 0.0;
+ *    assert(a.size() == b.size());
[…]

### 2015-05-30 src/is.h

@@ -33,9 +33,9 @@ namespace IS {
         int getSize() const { return size; };
         void setSize(int n) { size = n; };
         std::bitset<MAX> getBits() const { return bits; };
-        void setBits(std::bitset<MAX> a) { bits = a; }
+        void setBits(const std::bitset<MAX> &a) { bits = a; }
         void generateRandom(); 
-        void tweak();
+    //    void tweak();
         std::string to_str() {
             std::string str = "";
             for (int i = 0; i < size; i++)
@@ -56,15 +56,15 @@ namespace IS {
     };
 
 
-    class Metaheuristic {
-      public:
-        void optimize(const Problem &T, Solution &R);
-        double quality(const IS::Problem &T, const IS::Solution &S, double alpha);
-        double find_nearest(const Problem &p, const Instance &a);
-        void oneNN(const Problem &training, const Problem &result, std::vector<double> &category);
-        double calc_distance(const Instance &a, const Instance &b); 
-        void tweak(Solution S);
-    };
+ //   class Metaheuristic {
+ //     public:
+ //       void optimize(const Problem &T, Solution &R);
+ //       double quality(const IS::Problem &T, const IS::Solution &S, double alpha);
+ //       double find_nearest(const Problem &p, const Instance &a);
+ //       void oneNN(const Problem &training, const Problem &result, std::vector<double> &category);
+ //       double calc_distance(const Instance &a, const Instance &b); 
+ //       void tweak(Solution S);
+ //   };
 }
 
 #endif

### 2015-05-30 src/main.cpp

@@ -3,6 +3,7 @@
 #include <string>
 #include <cstdio>
 #include "is.h"
+#include "metaheuristic.h"
 
 IS::Problem load_data(char *file_name) {
     FILE *input = fopen(file_name, "r");
@@ -41,7 +42,7 @@ IS::Problem load_data(char *file_name) {
 int main(int argc, char *argv[]) {
 
     IS::Problem problem = load_data(argv[1]);
-    IS::Metaheuristic ls;
+    HillClimbing ls;
     IS::Solution sol(problem.size());
 
     std::cout << ">> Local search: " << std::endl;

### 2015-05-30 src/metaheuristic.cpp

@@ -1,0 +1,168 @@
+#include "metaheuristic.h"
+#include "is.h"
+
+using namespace std;
+
+/* Metaheuristics methods */
+
+double Metaheuristic::calc_distance(const IS::Instance &a, const IS::Instance &b) {
+    double dist = 0.0;
+    assert(a.size() == b.size());
+    for (int i = 0; i < a.size(); ++i)
+        dist += (a[i] - b[i]) * (a[i] - b[i]);
+    return sqrt(dist);
+}
+
+double Metaheuristic::find_nearest(const IS::Problem &p, const IS::Instance &a) {
+    double imin = 1.0 * MAX;
+    double category;
+    for (IS::Problem::const_iterator it = p.begin(); it < p.end(); ++it) {
+        double dist = calc_distance(*it, a);
+        if (dist < epsilon) {
+            category = (*it).getCategory();
+            break;
+        }
+        if (dist < imin) {
+            imin = dist;                  
+            category = (*it).getCategory();
+        }
+    }
+    return category;
+}
+
+void Metaheuristic::oneNN(const IS::Problem &training, 
+                          const IS::Problem &result, 
+                          std::vector<double> &category) {
+
+    for (int i = 0; i < result.size(); i++)
+        category[i] = find_nearest(training, result[i]);
+}
+
+double Metaheuristic::quality(const IS::Problem &T, 
+                              const IS::Solution &S,
+                              double alpha) {
+
+    std::bitset<MAX> bits = S.getBits();
+    IS::Problem training, result;
+    std::vector<double> category;
+    int j = 0;
+    for (int i = 0; i < T.size(); i++) {
+        if (bits.test(i)) training.push_back(T[i]); 
+        else result.push_back(T[i]);
+    }
+
+    category.assign(result.size(), -1);
+    oneNN(training, result, category);
+
+    int count = 0;
+    for (int i = 0; i < result.size(); i++)
+        if (result[i].getCategory() == category[i])
+            count++;
+
+    double clas_rate = (1.0 * count) / (1.0 * result.size());
+    double perc_redc = (1.0 * result.size()) / (1.0 * T.size());
+
+    cout << "Hubo un total de " << count << " aciertos" << endl;
+    cout << "Result es de tamanio: " << result.size() << endl;
[…]

### 2015-05-30 src/metaheuristic.h

@@ -1,0 +1,36 @@
+#ifndef METAHEURISTIC_H
+#define METAHEURISTIC_H
+
+#include <iostream>
+#include "is.h"
+
+const double epsilon = 0.00001;
+
+class Metaheuristic {
+    public:
+        virtual void optimize(const IS::Problem &T, IS::Solution &R) = 0;
+        double quality(const IS::Problem &T, const IS::Solution &S, double alpha);
+        double find_nearest(const IS::Problem &p, const IS::Instance &a);
+        void oneNN(const IS::Problem &training, 
+                   const IS::Problem &result, 
+                   std::vector<double> &category);
+        double calc_distance(const IS::Instance &a, const IS::Instance &b); 
+        virtual void tweak(IS::Solution &S) = 0;
+};
+
+class HillClimbing : public Metaheuristic {
+    public: 
+        void optimize(const IS::Problem &T, IS::Solution &R);
+        void tweak(IS::Solution &S);
+};
+
+class SimulatedAnnealing : public Metaheuristic {
+    private:
+    public: 
+        SimulatedAnnealing(): temperature(160000), dec_factor(10) { };
+        SimulatedAnnealing(int t, int d): temperature(t), dec_factor(d) { };
+        void optimize(const IS::Problem &T, IS::Solution &R);
+        void tweak(IS::Solution &S);
+};
+
+#endif

### 2015-05-30 src/is.cpp
* refactor * Tweak operator (toggle random bit) in simulated annealing

@@ -7,47 +7,6 @@
 #include <iostream>
 #include "is.h"
 
-const double epsilon = 0.00001;
-
-/*
- *double IS::Metaheuristic::find_nearest(const Problem &p, const Instance &a) {
- *    double imin = 1.0 * MAX;
- *    double category;
- *    for (Problem::const_iterator it = p.begin(); it < p.end(); ++it) {
- *        double dist = calc_distance(*it, a);
- *        if (dist < epsilon) {
- *            category = (*it).getCategory();
- *            break;
- *        }
- *        if (dist < imin) {
- *            imin = dist;                  
- *            category = (*it).getCategory();
- *        }
- *    }
- *    return category;
- *}
- */
-
-/*
- *void IS::Metaheuristic::oneNN(const Problem &training, 
- *                              const Problem &result, 
- *                              std::vector<double> &category) {
- *
- *    for (int i = 0; i < result.size(); i++)
- *        category[i] = find_nearest(training, result[i]);
- *}
- */
-
-/*
- *double IS::Metaheuristic::calc_distance(const Instance &a, const Instance &b) {
- *    double dist = 0.0;
- *    assert(a.size() == b.size());
- *    for (int i = 0; i < a.size(); ++i)
- *        dist += (a[i] - b[i]) * (a[i] - b[i]);
- *    return sqrt(dist);
- *}
- */
-
 void IS::Solution::generateRandom() {
     srand(time(NULL));
     for (int i = 0; i < size; i++) {
@@ -55,75 +14,3 @@ void IS::Solution::generateRandom() {
         this->bits[i] = (p > 10) ? 1 : 0;
     }
 }
-
-//void IS::Solution::tweak() {
-//    srand(time(NULL));
-//    int n = (rand() % bits.count()) + 1;
-//    int i, j;
-//    for (j = 0, i = 0; j < size && i < n; j++) if (bits.test(j)) i++;
-//    assert(bits.test(j-1));
-//    this->bits[j-1] = 0;
-//}
-
-//void IS::Metaheuristic::optimize(const IS::Problem &T, IS::Solution &S) {
-//    S.setSize(T.size());
-//    S.generateRandom();
-//    double q_max, old_q = -1;
-//    int iter_old = 0;
-//    while (1) {
-//        IS::Solution R(S);
-//        R.tweak(); 
-//        assert((S.getBits() ^ R.getBits()).count() == 1);
-//
-//        // Using 0.5 because paper
[…]

### 2015-05-30 src/metaheuristic.h

@@ -7,30 +7,31 @@
 const double epsilon = 0.00001;
 
 class Metaheuristic {
-    public:
-        virtual void optimize(const IS::Problem &T, IS::Solution &R) = 0;
-        double quality(const IS::Problem &T, const IS::Solution &S, double alpha);
-        double find_nearest(const IS::Problem &p, const IS::Instance &a);
-        void oneNN(const IS::Problem &training, 
-                   const IS::Problem &result, 
-                   std::vector<double> &category);
-        double calc_distance(const IS::Instance &a, const IS::Instance &b); 
-        virtual void tweak(IS::Solution &S) = 0;
+  public:
+    virtual void optimize(const IS::Problem &T, IS::Solution &R) = 0;
+    double quality(const IS::Problem &T, const IS::Solution &S, double alpha);
+    double find_nearest(const IS::Problem &p, const IS::Instance &a);
+    void oneNN(const IS::Problem &training, 
+               const IS::Problem &result, 
+               std::vector<double> &category);
+    double calc_distance(const IS::Instance &a, const IS::Instance &b); 
+    virtual void tweak(IS::Solution &S) = 0;
 };
 
 class HillClimbing : public Metaheuristic {
-    public: 
-        void optimize(const IS::Problem &T, IS::Solution &R);
-        void tweak(IS::Solution &S);
+  public: 
+    void optimize(const IS::Problem &T, IS::Solution &R);
+    void tweak(IS::Solution &S);
 };
 
 class SimulatedAnnealing : public Metaheuristic {
-    private:
-    public: 
-        SimulatedAnnealing(): temperature(160000), dec_factor(10) { };
-        SimulatedAnnealing(int t, int d): temperature(t), dec_factor(d) { };
-        void optimize(const IS::Problem &T, IS::Solution &R);
-        void tweak(IS::Solution &S);
+  private:
+    int temperature, dec_factor;
+  public: 
+    SimulatedAnnealing(): temperature(160000), dec_factor(10) { };
+    SimulatedAnnealing(int t, int d): temperature(t), dec_factor(d) { };
+    void optimize(const IS::Problem &T, IS::Solution &R);
+    void tweak(IS::Solution &S);
 };
 
 #endif

### 2015-05-31 src/is.cpp
*Minor refactor: Problem class is now Dataset *Added arguments for ILS

@@ -30,12 +30,12 @@ double IS::Instance::calcDistance(const IS::Instance &b) const {
     return sqrt(dist);
 } 
 
-void IS::Problem::getDispersions(std::vector<double> &dispersion) const {
+void IS::Dataset::getDispersions(std::vector<double> &dispersion) const {
     dispersion.assign(0.0, size());
     IS::Instance dummy;
     dummy.assign(0.0, size());
 
-    const IS::Problem &a = *this;
+    const IS::Dataset &a = *this;
     for (int i = 0; i < a.size(); i++)
         dispersion[i] = a[i].calcDistance(dummy);
 }

### 2015-05-31 src/is.h

@@ -48,13 +48,13 @@ namespace IS {
         }
     };
 
-    class Problem : public std::vector<Instance> {
+    class Dataset : public std::vector<Instance> {
       private:
         int attrs;
       public:
-        Problem() { };
-        Problem(const Problem& a) : vector<Instance>(a.begin(), a.end()) { };
-        Problem(int N) : std::vector<Instance>(N) {};
+        Dataset() { };
+        Dataset(const Dataset& a) : vector<Instance>(a.begin(), a.end()) { };
+        Dataset(int N) : std::vector<Instance>(N) {};
         void setAttrs(int a) { attrs = a; }
         int getAttrs() const { return attrs; }
         void getDispersions(std::vector<double> &dispersion) const;

### 2015-05-31 src/main.cpp

@@ -14,7 +14,7 @@ int moa_counter = 0;
 int tweaker_optional_arg[10];
 int toa_counter = 0;
 
-IS::Problem load_data(std::string file_name_str) {
+IS::Dataset load_data(std::string file_name_str) {
     const char *file_name = file_name_str.c_str();
     FILE *input = fopen(file_name, "r");
     int instances, attrs;
@@ -31,7 +31,7 @@ IS::Problem load_data(std::string file_name_str) {
         exit(1);
     }
     
-    IS::Problem problem = IS::Problem(instances);
+    IS::Dataset problem = IS::Dataset(instances);
     problem.setAttrs(attrs);
 
     for(unsigned i = 0; i < instances; ++i) {
@@ -93,7 +93,7 @@ void error_(std::string msg) {
     exit(1);
 }
 
-void print_solution(const IS::Problem &problem, const IS::Solution &sol) {
+void print_solution(const IS::Dataset &problem, const IS::Solution &sol) {
     int count = 0;
     std::bitset<MAX> bits = sol.getBits();
     for (unsigned i = 0; i < sol.getSize(); ++i) {
@@ -108,7 +108,7 @@ void print_solution(const IS::Problem &problem, const IS::Solution &sol) {
     std::cout << "Solution of size: " << count << std::endl;
 }
 
-void print_dispersions(const IS::Problem &problem) {
+void print_dispersions(const IS::Dataset &problem) {
     std::vector<double> dispersion;
     problem.getDispersions(dispersion);
     std::cout << ">>>>>> Dispersions: " << std::endl;
@@ -133,7 +133,15 @@ Metaheuristic *choose_metaheuristic(std::string metaheuristic_str) {
         int l = metaheuristic_optional_arg[0]; 
         int n = metaheuristic_optional_arg[1];
         return new Tabu(l, n);
-    } else if (metaheuristic_str == "ILS") return new ILS();
+    } else if (metaheuristic_str == "ILS") {
+        
+        // Print error and exit
+        if (moa_counter != 3) error_("ILS needs three arguments");
+        int l = metaheuristic_optional_arg[0]; 
+        int t = metaheuristic_optional_arg[1];
+        int p = metaheuristic_optional_arg[2];
+        return new ILS(l, t, p);
+    }
     error_("The metaheuristic you are providing doesn't exist");
     return NULL;   // Shut up warning
 }
@@ -201,14 +209,14 @@ int main(int argc, char *argv[]) {
[…]

### 2015-05-31 src/metaheuristic.cpp

@@ -9,10 +9,10 @@ double Metaheuristic::calc_distance(const IS::Instance &a, const IS::Instance &b
     return a.calcDistance(b);
 }
 
-double Metaheuristic::find_nearest(const IS::Problem &p, const IS::Instance &a) {
+double Metaheuristic::find_nearest(const IS::Dataset &p, const IS::Instance &a) {
     double imin = 1.0 * MAX;
     double category;
-    for (IS::Problem::const_iterator it = p.begin(); it < p.end(); ++it) {
+    for (IS::Dataset::const_iterator it = p.begin(); it < p.end(); ++it) {
         double dist = calc_distance(*it, a);
         if (dist < epsilon) {
             category = (*it).getCategory();
@@ -26,20 +26,20 @@ double Metaheuristic::find_nearest(const IS::Problem &p, const IS::Instance &a) 
     return category;
 }
 
-void Metaheuristic::oneNN(const IS::Problem &training, 
-                          const IS::Problem &result, 
+void Metaheuristic::oneNN(const IS::Dataset &training, 
+                          const IS::Dataset &result, 
                           std::vector<double> &category) {
 
     for (int i = 0; i < result.size(); i++)
         category[i] = find_nearest(training, result[i]);
 }
 
-double Metaheuristic::quality(const IS::Problem &T, 
+double Metaheuristic::quality(const IS::Dataset &T, 
                               const IS::Solution &S,
                               double alpha) {
 
     std::bitset<MAX> bits = S.getBits();
-    IS::Problem training, result;
+    IS::Dataset training, result;
     std::vector<double> category;
     int j = 0;
     for (int i = 0; i < T.size(); i++) {
@@ -74,7 +74,7 @@ double Metaheuristic::quality(const IS::Problem &T,
 
 /* Hill Climbing */
 
-void HillClimbing::optimize(const IS::Problem &T, IS::Solution &S) {
+void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) {
     std::cout << ">>>> Running Hill Climbing" << std::endl;
     S.setSize(T.size());
     S.generateRandom();
@@ -104,7 +104,7 @@ void HillClimbing::optimize(const IS::Problem &T, IS::Solution &S) {
 
 /* Simulated Annealing  */
 
-void SimulatedAnnealing::optimize(const IS::Problem &T, IS::Solution &S) {
[…]

### 2015-05-31 src/metaheuristic.h

@@ -13,20 +13,20 @@ class Metaheuristic {
   protected:
     Tweaker *tweaker;
   public:
-    virtual void optimize(const IS::Problem &T, IS::Solution &R) = 0;
+    virtual void optimize(const IS::Dataset &T, IS::Solution &R) = 0;
     void setTweaker(Tweaker *tw) { tweaker = tw; };
     Tweaker* getTweaker() { return tweaker; };
-    double quality(const IS::Problem &T, const IS::Solution &S, double alpha);
-    double find_nearest(const IS::Problem &p, const IS::Instance &a);
-    void oneNN(const IS::Problem &training, 
-               const IS::Problem &result, 
+    double quality(const IS::Dataset &T, const IS::Solution &S, double alpha);
+    double find_nearest(const IS::Dataset &p, const IS::Instance &a);
+    void oneNN(const IS::Dataset &training, 
+               const IS::Dataset &result, 
                std::vector<double> &category);
     double calc_distance(const IS::Instance &a, const IS::Instance &b); 
 };
 
 class HillClimbing : public Metaheuristic {
   public: 
-    void optimize(const IS::Problem &T, IS::Solution &R);
+    void optimize(const IS::Dataset &T, IS::Solution &R);
 };
 
 class SimulatedAnnealing : public Metaheuristic {
@@ -34,13 +34,22 @@ class SimulatedAnnealing : public Metaheuristic {
     int temperature, dec_factor;
   public: 
     SimulatedAnnealing(int t, int d): temperature(t), dec_factor(d) { };
-    void optimize(const IS::Problem &T, IS::Solution &R);
+    void optimize(const IS::Dataset &T, IS::Solution &R);
 };
 
 
 class ILS : public Metaheuristic {
+  private:
+    int li, ti, pp;
   public: 
-    void optimize(const IS::Problem &T, IS::Solution &R);
+    ILS(int l, int t, int p) : li(l), ti(t), pp(p) { };
+    int  getLocalIter() { return li; }
+    int  getTotalIter() { return ti; }
+    int  getPerturbPerc() { return pp; }
+    void setLocalIter(int l) { li = l; }
+    void setTotalIter(int t) { ti = t; }
+    void setPerturbPerc(int p) { pp = p; }
+    void optimize(const IS::Dataset &T, IS::Solution &R);
 };
 
 class Tabu : public Metaheuristic {
[…]

### 2015-05-31 src/is.cpp
*Modified main for print dispersions *Fix bug in print dispersions

@@ -24,6 +24,7 @@ void IS::Solution::copy(const Solution &a) {
 double IS::Instance::calcDistance(const IS::Instance &b) const {
     const IS::Instance &a = *this;
     double dist = 0.0;
+    cout << "a: " << a.size() << " b: " << b.size() << endl;
     assert(a.size() == b.size());
     for (int i = 0; i < a.size(); ++i)
         dist += (a[i] - b[i]) * (a[i] - b[i]);
@@ -31,9 +32,9 @@ double IS::Instance::calcDistance(const IS::Instance &b) const {
 } 
 
 void IS::Dataset::getDispersions(std::vector<double> &dispersion) const {
-    dispersion.assign(0.0, size());
+    dispersion.assign(size(), 0.0);
     IS::Instance dummy;
-    dummy.assign(0.0, size());
+    dummy.assign(getAttrs()-1, 0.0);
 
     const IS::Dataset &a = *this;
     for (int i = 0; i < a.size(); i++)

### 2015-05-31 src/main.cpp

@@ -108,12 +108,12 @@ void print_solution(const IS::Dataset &problem, const IS::Solution &sol) {
     std::cout << "Solution of size: " << count << std::endl;
 }
 
-void print_dispersions(const IS::Dataset &problem) {
+void print_dispersions(const IS::Dataset &dataset) {
     std::vector<double> dispersion;
-    problem.getDispersions(dispersion);
+    dataset.getDispersions(dispersion);
     std::cout << ">>>>>> Dispersions: " << std::endl;
     for (int i = 0; i < dispersion.size(); i++)
-        std::cout << dispersion[i];
+        std::cout << dispersion[i] << " ";
     std::cout << endl;
 }
 
@@ -210,13 +210,19 @@ int main(int argc, char *argv[]) {
     if (! flag_x) error_("You have to specify a tweaker");
 
     IS::Dataset dataset = load_data(file_name);
-    Metaheuristic *metaheuristic = choose_metaheuristic(metaheuristic_str);
-    Tweaker *tweaker = choose_tweaker(tweaker_str);
 
-    metaheuristic->setTweaker(tweaker);
+    if (flag_d) {
+        print_dispersions(dataset);
+    } else {
+        Metaheuristic *metaheuristic = choose_metaheuristic(metaheuristic_str);
+        Tweaker *tweaker = choose_tweaker(tweaker_str);
 
-    IS::Solution solution(dataset.size());
-    metaheuristic->optimize(dataset, solution);
+        metaheuristic->setTweaker(tweaker);
+
+        IS::Solution solution(dataset.size());
+        metaheuristic->optimize(dataset, solution);
+    }
+
 
 
     return 0;

### 2015-05-31 src/main.cpp
Dispersions are now printed 10 at a time

@@ -1,12 +1,13 @@
-#include <cstdlib>
+#include <algorithm>
 #include <iostream>
-#include <string>
-#include <cstring>
-#include <cstdio>
 #include <getopt.h>
-#include "is.h"
+#include <cstdlib>
+#include <cstring>
+#include <string>
+#include <cstdio>
 #include "metaheuristic.h"
 #include "tweaker.h"
+#include "is.h"
 
 int metaheuristic_optional_arg[10];
 int moa_counter = 0;
@@ -111,9 +112,16 @@ void print_solution(const IS::Dataset &problem, const IS::Solution &sol) {
 void print_dispersions(const IS::Dataset &dataset) {
     std::vector<double> dispersion;
     dataset.getDispersions(dispersion);
+
+    std::sort(dispersion.begin(), dispersion.end());
+
     std::cout << ">>>>>> Dispersions: " << std::endl;
     for (int i = 0; i < dispersion.size(); i++)
-        std::cout << dispersion[i] << " ";
+        if(i % 10 == 0 and i != 0)  {
+            std::cout << dispersion[i] << " " << std::endl;
+        } else {
+            std::cout << dispersion[i] << " ";
+        }
     std::cout << endl;
 }
 

### 2015-05-31 src/is.cpp
*Added data sets *Minor fixes

@@ -24,7 +24,6 @@ void IS::Solution::copy(const Solution &a) {
 double IS::Instance::calcDistance(const IS::Instance &b) const {
     const IS::Instance &a = *this;
     double dist = 0.0;
-    cout << "a: " << a.size() << " b: " << b.size() << endl;
     assert(a.size() == b.size());
     for (int i = 0; i < a.size(); ++i)
         dist += (a[i] - b[i]) * (a[i] - b[i]);

### 2015-05-31 src/main.cpp

@@ -1,13 +1,16 @@
-#include <algorithm>
 #include <iostream>
 #include <getopt.h>
 #include <cstdlib>
 #include <cstring>
+#include <cassert>
 #include <string>
 #include <cstdio>
+#include <ctime>
 #include "metaheuristic.h"
 #include "tweaker.h"
 #include "is.h"
+
+typedef IS::Dataset ds;
 
 int metaheuristic_optional_arg[10];
 int moa_counter = 0;
@@ -15,10 +18,10 @@ int moa_counter = 0;
 int tweaker_optional_arg[10];
 int toa_counter = 0;
 
-IS::Dataset load_data(std::string file_name_str) {
+std::pair<ds, ds> load_data(std::string file_name_str, int partition) {
     const char *file_name = file_name_str.c_str();
     FILE *input = fopen(file_name, "r");
-    int instances, attrs;
+    int insts_size, num_attrs;
 
     if (input) 
         printf("File: %s\n", file_name);
@@ -27,26 +30,56 @@ IS::Dataset load_data(std::string file_name_str) {
         exit(1);
     }
 
-    if (fscanf(input, "%d %d", &instances, &attrs) != 2) {
+    if (fscanf(input, "%d %d", &insts_size, &num_attrs) != 2) {
         printf("ERROR at data load: file malformed\n");
         exit(1);
     }
     
-    IS::Dataset problem = IS::Dataset(instances);
-    problem.setAttrs(attrs);
+    int val_size = insts_size / 10;
+    int tra_size = insts_size - val_size;
 
-    for(unsigned i = 0; i < instances; ++i) {
-        // attrs-1 because last element of line is category
-        double attr;
-        for(unsigned j = 0; j < attrs-1; ++j) { 
+    // Two datasets are built: One for validation and one for training
+    IS::Dataset validation = IS::Dataset();
+    IS::Dataset training   = IS::Dataset();
+    training.setAttrs(num_attrs);
+    validation.setAttrs(num_attrs);
+
+    std::vector<int> val_pos;
+    
+    srand(time(NULL));
+    for(unsigned i = 0; i < val_size; ++i) {
+        int j = rand() % tra_size;
+        val_pos.push_back(j);
+    }
+    assert(val_pos.size() == val_size);
+
+    for(unsigned i = 0; i < insts_size; ++i) {
+        if(std::find(val_pos.begin(), val_pos.end(), i) != val_pos.end()) {
+            /* validation instance */
[…]

### 2015-05-31 src/is.cpp
*Added bitsOn method *Hillclimbing optimize prints info every 100 iters

@@ -21,6 +21,12 @@ void IS::Solution::copy(const Solution &a) {
     bits |= a.bits;
 }
 
+int IS::Solution::bitsOn() {
+    int on = 0;
+    for(unsigned i = 0; i < size; ++i) if(bits[i]) on++;
+    return on;
+}
+
 double IS::Instance::calcDistance(const IS::Instance &b) const {
     const IS::Instance &a = *this;
     double dist = 0.0;
@@ -34,7 +40,6 @@ void IS::Dataset::getDispersions(std::vector<double> &dispersion) const {
     dispersion.assign(size(), 0.0);
     IS::Instance dummy;
     dummy.assign(getAttrs()-1, 0.0);
-
     const IS::Dataset &a = *this;
     for (int i = 0; i < a.size(); i++)
         dispersion[i] = a[i].calcDistance(dummy);

### 2015-05-31 src/is.h

@@ -36,6 +36,7 @@ namespace IS {
         std::bitset<MAX> getBits() const { return bits; };
         void setBits(const std::bitset<MAX> &a) { bits = a; }
         void generateRandom(); 
+        int bitsOn();
         std::string to_str() {
             std::string str = "";
             for (int i = 0; i < size; i++)

### 2015-05-31 src/metaheuristic.cpp

@@ -60,12 +60,12 @@ double Metaheuristic::quality(const IS::Dataset &T,
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
 
     // XXX: Print
-    cout << "Hubo un total de " << count << " aciertos" << endl;
-    cout << "Result es de tamanio: " << result.size() << endl;
-    cout << "La relación es: " << clas_rate << endl;
-    cout << "El tamaño de T es: " << T.size() << endl;
-    cout << "El porcentaje de reducción es: " << perc_redc << endl;
-    cout << "El fitness es : " << fitness << endl << endl;
+    // cout << "Hubo un total de " << count << " aciertos" << endl;
+    // cout << "Result es de tamanio: " << result.size() << endl;
+    // cout << "La relación es: " << clas_rate << endl;
+    // cout << "El tamaño de T es: " << T.size() << endl;
+    // cout << "El porcentaje de reducción es: " << perc_redc << endl;
+    // cout << "El fitness es : " << fitness << endl << endl;
 
     assert(fitness <= 1.0);
     return fitness;
@@ -79,8 +79,9 @@ void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) {
     S.setSize(T.size());
     S.generateRandom();
     double q_max, old_q = -1;
-    int iter_old = 0;
+    int iter_old = 0, iter_total = 0;
     while (1) {
+        iter_total++;
         IS::Solution R(S);
         tweaker->tweak(R); 
         assert((S.getBits() ^ R.getBits()).count() == 1);
@@ -98,8 +99,13 @@ void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) {
         }
         if (q_max > 0.95 or iter_old == 1000) break;
         old_q = q_max;
+
+        if(iter_total % 100 == 0 and iter_total != 0) {
+            cout << "Total iterations: " << iter_total << endl;
+            cout << "Best fitness so far: " << q_max << endl;
+            cout << "Solution size: " << S.bitsOn() << endl << endl;
+        }
     }
-    cout << "---> " << q_max << endl;
 }
 
 /* Simulated Annealing  */

### 2015-05-31 src/is.h
*Fixed copy constructor for Datasets *Fixed parameters requirements for main and -d flag *Comments and TODOs

@@ -54,7 +54,7 @@ namespace IS {
         int attrs;
       public:
         Dataset() { };
-        Dataset(const Dataset& a) : vector<Instance>(a.begin(), a.end()) { };
+        Dataset(const Dataset& a) : vector<Instance>(a.begin(), a.end()), attrs(a.getAttrs()) { };
         Dataset(int N) : std::vector<Instance>(N) {};
         void setAttrs(int a) { attrs = a; }
         int getAttrs() const { return attrs; }

### 2015-05-31 src/main.cpp

@@ -79,7 +79,7 @@ std::pair<ds, ds> load_data(std::string file_name_str, int partition) {
         }
     }
 
-    return pair<ds,ds>(training, validation);
+    return make_pair(training, validation);
 }
 
 void print_help() {
@@ -246,9 +246,9 @@ int main(int argc, char *argv[]) {
     if (flag_h) return 0; // Print help and exit
 
     // Checking and exit in case of error
-    if (! flag_f) error_("You have to specify a file");  
-    if (! flag_g) error_("You have to specify a metaheuristic"); 
-    if (! flag_x) error_("You have to specify a tweaker");
+    if (! flag_f and ! flag_d) error_("You have to specify a file");  
+    if (! flag_g and ! flag_d) error_("You have to specify a metaheuristic"); 
+    if (! flag_x and ! flag_d) error_("You have to specify a tweaker");
 
     // TODO: Make this a parameter
     int part = 10;
@@ -257,20 +257,18 @@ int main(int argc, char *argv[]) {
     IS::Dataset training   = dataset.first;
     IS::Dataset validation = dataset.second;
 
+    if (flag_d) {
+        print_dispersions(training);
+        return 0;
+    }
+
     Metaheuristic *metaheuristic = choose_metaheuristic(metaheuristic_str);
     Tweaker *tweaker = choose_tweaker(tweaker_str);
 
-    if (flag_d) {
-        print_dispersions(training);
-    } else {
-        Metaheuristic *metaheuristic = choose_metaheuristic(metaheuristic_str);
-        Tweaker *tweaker = choose_tweaker(tweaker_str);
+    metaheuristic->setTweaker(tweaker);
 
-        metaheuristic->setTweaker(tweaker);
-
-        IS::Solution solution(training.size());
-        metaheuristic->optimize(training, solution);
-    }
+    IS::Solution solution(training.size());
+    metaheuristic->optimize(training, solution);
 
     return 0;
 }

### 2015-05-31 src/metaheuristic.cpp

@@ -84,7 +84,7 @@ void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) {
         iter_total++;
         IS::Solution R(S);
         tweaker->tweak(R); 
-        assert((S.getBits() ^ R.getBits()).count() == 1);
+        // assert((S.getBits() ^ R.getBits()).count() == 1);
 
         // Using 0.5 because paper
         double q1 = quality(T, R, 0.5), q2 = quality(T, S, 0.5);
@@ -97,6 +97,7 @@ void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) {
         } else {
             iter_old = 0;
         }
+        // TODO: Pass these values as parameters
         if (q_max > 0.95 or iter_old == 1000) break;
         old_q = q_max;
 

### 2015-05-31 src/main.cpp
Statistics at the end of run

@@ -270,5 +270,31 @@ int main(int argc, char *argv[]) {
     IS::Solution solution(training.size());
     metaheuristic->optimize(training, solution);
 
+    // Checking quality of final solution
+    std::vector<double> categories;
+    categories.assign(validation.size(), -1);
+
+    std::bitset<MAX> bits = solution.getBits();
+    IS::Dataset final;
+    for(unsigned i = 0; i < training.size(); ++i) {
+        if(bits[i]) final.push_back(training[i]);
+    }
+
+    metaheuristic->oneNN(final, validation, categories);
+
+    int errors = 0;
+    for (int i = 0; i < validation.size(); i++)
+        if (validation[i].getCategory() != categories[i])
+            errors++;
+
+    double clas_rate = (1.0 * errors) / (1.0 * validation.size());
+    double perc_redc = (1.0 * final.size()) / (1.0 * training.size());
+    double fitness = 0.5 * clas_rate + (1 - 0.5) * perc_redc;
+
+    // XXX: Print
+    cout << "Total errors: " << errors << endl;
+    cout << "Validation error %: " << (errors * 100.0) / (validation.size() * 1.0) << endl;
+    cout << "Size %: " << (final.size() * 100.0) / (training.size() * 1.0) << endl;
+
     return 0;
 }

### 2015-05-31 src/main.cpp
New tweaker (percRandomFlips)

@@ -198,6 +198,10 @@ Tweaker *choose_tweaker(std::string tweaker_str) {
         // Print error and exit
         if (toa_counter != 1) error_("nRandomFlips needs one argument");
         return new nRandomFlips(tweaker_optional_arg[0]);
+    } else if (tweaker_str == "percRandomFlips") {
+        // Print error and exit
+        if (toa_counter != 1) error_("percRandomFlips needs one argument");
+        return new percRandomFlips(tweaker_optional_arg[0]);
     }
     return NULL;
 }

### 2015-05-31 src/tweaker.cpp

@@ -37,3 +37,18 @@ void nRandomFlips::tweak(IS::Solution &S) {
     S.setBits(bits);
 }
 
+void percRandomFlips::tweak(IS::Solution &S) {
+    std::bitset<MAX> bits = S.getBits();
+    int size = S.getSize();
+    int n = (perc * size) / 100;
+
+    srand(time(NULL));
+    for(unsigned i = 0; i < n; ++i) {
+        int j = (rand() % size) + 1;
+        bits.flip(j);
+    }
+    S.setBits(bits);
+}
+
+
+

### 2015-05-31 src/tweaker.h

@@ -29,5 +29,13 @@ class nRandomFlips : public Tweaker {
     void tweak(IS::Solution &S);
 };
 
+class percRandomFlips : public Tweaker {
+  private:
+    int perc;
+  public:
+    percRandomFlips(int x) : perc(x) { }
+    void tweak(IS::Solution &S);
+};
+
 
 #endif

### 2015-05-31 src/main.cpp
Working on results

@@ -18,6 +18,13 @@ int moa_counter = 0;
 int tweaker_optional_arg[10];
 int toa_counter = 0;
 
+typedef struct result {
+    int errors;
+    float size;
+    float val_error;
+} result;
+
+
 std::pair<ds, ds> load_data(std::string file_name_str, int partition) {
     const char *file_name = file_name_str.c_str();
     FILE *input = fopen(file_name, "r");
@@ -220,8 +227,8 @@ int main(int argc, char *argv[]) {
     };
 
     // Parser of CLI
-    int opt, option_index;
-    bool flag_h, flag_f, flag_d, flag_g, flag_m, flag_o;
+    int opt, option_index, runs;
+    bool flag_h, flag_f, flag_d, flag_g, flag_m, flag_o, flag_n;
     bool flag_x;
     memset(metaheuristic_optional_arg, 0, sizeof(metaheuristic_optional_arg));
     memset(tweaker_optional_arg, 0, sizeof(tweaker_optional_arg));
@@ -230,7 +237,7 @@ int main(int argc, char *argv[]) {
     std::string file_name, metaheuristic_str, tweaker_str;
 
     while (1) {
-        opt = getopt_long(argc, argv, "hf:dg:m:o:x:", long_options, &option_index);
+        opt = getopt_long(argc, argv, "hf:dg:m:o:x:n:", long_options, &option_index);
         if (opt == -1) break;
 
         switch (opt) {
@@ -244,6 +251,7 @@ int main(int argc, char *argv[]) {
             case 'o': tweaker_optional_arg[toa_counter++] = atoi(optarg); 
                       break;
             case 'x': flag_x = true; tweaker_str = optarg; break;
+            case 'n': flag_n = true; runs = atoi(optarg); break;
         }
     }
 
@@ -253,6 +261,10 @@ int main(int argc, char *argv[]) {
     if (! flag_f and ! flag_d) error_("You have to specify a file");  
     if (! flag_g and ! flag_d) error_("You have to specify a metaheuristic"); 
     if (! flag_x and ! flag_d) error_("You have to specify a tweaker");
+    if (! flag_n and ! flag_d) {
+        runs = 1;
+    }
+
 
     // TODO: Make this a parameter
     int part = 10;
@@ -270,35 +282,49 @@ int main(int argc, char *argv[]) {
     Tweaker *tweaker = choose_tweaker(tweaker_str);
 
     metaheuristic->setTweaker(tweaker);
[…]

### 2015-06-01 src/main.cpp
*Modified results container (now vectors of values) *Some uninitialized jumps (this was a b****) *New option for output to file *Verification of non-existant tweaker or metaheuristic

@@ -1,5 +1,6 @@
 #include <iostream>
 #include <getopt.h>
+#include <fstream>
 #include <cstdlib>
 #include <cstring>
 #include <cassert>
@@ -18,11 +19,12 @@ int moa_counter = 0;
 int tweaker_optional_arg[10];
 int toa_counter = 0;
 
-typedef struct result {
-    int errors;
-    float size;
-    float val_error;
+typedef struct results {
+    std::vector<int> errors;
+    std::vector<float> sizes;
+    std::vector<float> val_errors;
 } result;
+
 
 
 std::pair<ds, ds> load_data(std::string file_name_str, int partition) {
@@ -223,28 +225,37 @@ int main(int argc, char *argv[]) {
         {"meta-optional-arg", required_argument, 0, 'm'},
         {"tweaker-optional-arg", required_argument, 0, 'o'},
         {"tweaker", required_argument, 0, 'x'},
+        {"results", required_argument, 0, 'r'},
         {0, 0, 0, 0}
     };
 
     // Parser of CLI
-    int opt, option_index, runs;
-    bool flag_h, flag_f, flag_d, flag_g, flag_m, flag_o, flag_n;
-    bool flag_x;
+    int opt = 0, option_index, runs;
+    bool flag_h = false; 
+    bool flag_f = false; 
+    bool flag_d = false; 
+    bool flag_g = false; 
+    bool flag_m = false; 
+    bool flag_o = false; 
+    bool flag_n = false;
+    bool flag_x = false;
+    bool flag_r = false;
     memset(metaheuristic_optional_arg, 0, sizeof(metaheuristic_optional_arg));
     memset(tweaker_optional_arg, 0, sizeof(tweaker_optional_arg));
 
     // We need these three things
     std::string file_name, metaheuristic_str, tweaker_str;
+    std::string out_fn;
 
     while (1) {
-        opt = getopt_long(argc, argv, "hf:dg:m:o:x:n:", long_options, &option_index);
+        opt = getopt_long(argc, argv, "dhf:g:m:o:x:n:r:", long_options, &option_index);
         if (opt == -1) break;
+        opt = char(opt);
 
         switch (opt) {
-            case 0: break;
+            case 0  : break;
             case 'h': flag_h = true; print_help(); break;
             case 'f': flag_f = true; file_name = optarg; break;
-            case 'd': flag_d = true; break;
             case 'g': flag_g = true; metaheuristic_str = optarg; break; 
[…]

### 2015-06-01 src/main.cpp
Minor bug with output

@@ -336,9 +336,9 @@ int main(int argc, char *argv[]) {
 
 
     std::ostream *out;
+    std::ofstream rFile;
 
     if(flag_r) {
-        std::ofstream rFile;
         rFile.open(out_fn, std::ios::out);
         out = &rFile;
     } else {
@@ -355,6 +355,7 @@ int main(int argc, char *argv[]) {
     *out << "Errors: " << emean << std::endl;
     *out << "Size of solution %: " << smean << std::endl;
     *out << "Validation errors: %: " << vmean << std::endl;
+    rFile.close();
 
     return 0;
 }

### 2015-06-01 src/metaheuristic.cpp
Length in tabu is now a percentage of size of dataset

@@ -213,12 +213,13 @@ void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) {
     std::cout << ">>>> Running Tabu" << std::endl;
     IS::Solution S(T.size());
     S.generateRandom();  // First random solution
+    int max_length = (length * T.size() / 100);
 
     std::deque<IS::Solution> tabu_list;   // Tabu list
     tabu_list.push_back(S);
     best.copy(S);
     while (1) {
-        if (tabu_list.size() > length) tabu_list.pop_front();
+        if (tabu_list.size() > max_length) tabu_list.pop_front();
         IS::Solution R(S);
         tweaker->tweak(R); 
 

### 2015-06-01 src/metaheuristic.cpp
Fixed stop conditions for tabu

@@ -218,7 +218,10 @@ void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) {
     std::deque<IS::Solution> tabu_list;   // Tabu list
     tabu_list.push_back(S);
     best.copy(S);
+    double q_max, old_q = -1;
+    int iter_old = 0, iter_total = 0;
     while (1) {
+        iter_total++;
         if (tabu_list.size() > max_length) tabu_list.pop_front();
         IS::Solution R(S);
         tweaker->tweak(R); 
@@ -247,7 +250,16 @@ void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) {
             double qs = quality(T, S, 0.5), qb = quality(T, best, 0.5);
             double q_max = std::max(qs, qb);
             if (qs > qb) best.copy(S);
-            if (q_max > 0.95) break;
         }
+        // TODO: Tune iter_total and iter_old
+        if (q_max > 0.95 || iter_total == 10000 || iter_old == 1000) break;
+
+        if (old_q == q_max) {
+          iter_old ++;  
+        } else {
+            iter_old = 0;
+        }
+
+        old_q = q_max;
     }
 }

### 2015-06-01 src/tweaker.cpp

@@ -1,3 +1,4 @@
+#include <iostream>
 #include "tweaker.h"
 
 void oneRandomUnset::tweak(IS::Solution &S) {

### 2015-06-01 src/main.cpp
Change in output

@@ -344,17 +344,34 @@ int main(int argc, char *argv[]) {
     } else {
         out = &std::cout;
     }
-    double esum = std::accumulate(res.errors.begin(), res.errors.end(), 0.0);
-    double emean = esum / res.errors.size();
-    double vsum = std::accumulate(res.sizes.begin(), res.sizes.end(), 0.0);
-    double vmean = vsum / res.sizes.size();
-    double ssum = std::accumulate(res.val_errors.begin(), res.val_errors.end(), 0.0);
-    double smean = ssum / res.val_errors.size();
+    // double esum = std::accumulate(res.errors.begin(), res.errors.end(), 0.0);
+    // double emean = esum / res.errors.size();
+    // double vsum = std::accumulate(res.sizes.begin(), res.sizes.end(), 0.0);
+    // double vmean = vsum / res.sizes.size();
+    // double ssum = std::accumulate(res.val_errors.begin(), res.val_errors.end(), 0.0);
+    // double smean = ssum / res.val_errors.size();
+
+    *out << "Val_errors" << std::endl;
+    for (unsigned i = 0; i < res.val_errors.size(); ++i)  {
+        *out << res.val_errors[i] << " ";
+    }
+    *out << std::endl;
+    *out << "Sizes" << std::endl;
+    for (unsigned i = 0; i < res.sizes.size(); ++i)  {
+        *out << res.sizes[i] << " ";
+    }
+    *out << std::endl;
+    *out << "Errors" << std::endl;
+    for (unsigned i = 0; i < res.errors.size(); ++i)  {
+        *out << res.errors[i] << " ";
+    }
+    *out << std::endl;
 
 
-    *out << "Errors: " << emean << std::endl;
-    *out << "Size of solution %: " << smean << std::endl;
-    *out << "Validation errors: %: " << vmean << std::endl;
+    // *out << "Errors: " << emean << std::endl;
+    // *out << "Size of solution %: " << smean << std::endl;
+    // *out << "Validation errors: %: " << vmean << std::endl;
+
     rFile.close();
 
     return 0;

### 2015-06-01 doc/paper.tex
paper

@@
 \usepackage[utf8]{inputenc}
+\usepackage{graphicx}
+\usepackage{color}
 
@@
 
+\subsection{Comparación}
+
+Una vez decididos los parametros de cada metaheurística, se realizaron 
+hasta 15 pruebas con cada una de ellas en \emph{Glass}. Los gráficos 
+presentes sugieren que \emph{ILS} es la que mejor que se comporta para este problema, 
+tanto en promedio del tamaño de la solución final (\% del conjunto de instancias), 
+como en porcentaje de errores sobre el conjunto de validación.
+
+\begin{center}
+  \includegraphics[scale=0.4]{val_errors.jpeg}~\\[1cm]
+\end{center}
+
+\begin{center}
+  \includegraphics[scale=0.4]{sizes.jpeg}~\\[1cm]
+\end{center}
+
 % include your own bib file like this:

### 2015-06-01 doc/paper.tex
paper

@@
 
+\subsection{Entonación}
+
+Cada metaheuística se comporta de manera distinta. Entonar 
+los parametros de cada una de ellas requirió de una serie 
+de pruebas sobre distintos problemas, parametrizando
+no solo dichos valores, sino hasta operadores de vencindad. 
+Por ejemplo, \emph{Simulated Annealing}, arrojaba mejores 
+resultados a medida que mayor era la \emph{temperatura} y menor el \emph{factor de decrecimiento},
+y usando como operador de vencidad \emph{percRandomFlips} con $5\%$. Sin embargo,
+aumentar la temperatura y disminuir el factor de decrecimiento implica ejecuciones 
+de mayor tiempo (mayor
+diversificación). Es por esto que para los resultados finales, se consideró $temperatura = 15.000$ 
+y $factor = 1$, obteniendo así un equilibrio entre el tiempo de ejecución y calidad de la solución.
+
 \subsection{Comparación}

### 2015-06-01 doc/paper.tex
paper

@@
 \subsection{Operador}
-Ante la necesidad de movernos en el
-espacio de busqueda, surge la idea de un operador 
-que modifique de manera sutil (y rápida) una solución
-dada. 
-En nuestro caso, nuestro operador selecciona al
-azar un bit ``prendido'' y lo ``apaga'', que se traduce
+
+Cada metaheurística se comporta de manera distinta, por lo
+que surge la necesidad de crear operadores de vencidad
+que mejor se acoplen a ellas.
+
+Se implementaron 5 operadores de vecindad, que se explican brevemente
+a continuacion:
+
+\begin{itemize}
+  \item \emph{oneRandomUnset}:
+    Este operador selecciona al
+    azar un bit ``prendido'' y lo ``apaga''. Esto se traduce
     en ``sacar'' una instancia del conjunto de instancias solución.
-
-De esta manera, se consiguen soluciones cada vez más pequeñas y una tasa
-de aciertos $tasa\_clas$ cada vez más precisas: esto viene de si tenemos 
-$x$ aciertos y $w$ instancias que no están en la solución, aumetar el tamaño
-de una solución conllevaría a una posible $tasa\_clas = x / (w - 1) $,
-que supondría una mayor $tasa\_clas$ y por lo tanto una mejor solución. Sin embargo, 
-esto no puede estar más alejado de la realidad.
+  \item \emph{upToNRandomFlips}: Este operador tiene como parametro
+    un $N$. En cada operación hace \emph{toggle} de un número aleatorio 
+    de bits que es mayor que $1$ y menor $N$, y en donde dichos bits 
+    se escogen aleatoriamente.
+  \item \emph{nRandomFlips}: Al igual que el operador anterior, 
+    requiere de un $N$. Este operador simplemente hace toggle de $N$ bits 
+    escogidos aleatoriamente.
+  \item \emph{percRandomFlips}: Este operador requiere como parametro $X$, que 
+    representa un porcentaje. La ideas es hacer \emph{toggle} de un porcentaje de la
+    cantidad de instancias del problema original. Es decir, si un problema tiene 1000
+    instancias, y el porcentaje escogido es $5 \%$, entonces se hará \emph{toggle} 
+    sobre un total de 50 bits de la solución factible
+\end{itemize}
 
@@
 
-Cada metaheuística se comporta de manera distinta. Entonar 
-los parametros de cada una de ellas requirió de una serie 
+Cada metaheuística requiere de distintos parametros para realizar
+su trabajo. Entonar cada una de ellos requirió de una serie 
 de pruebas sobre distintos problemas, parametrizando
-no solo dichos valores, sino hasta operadores de vencindad. 
+no solo valores necesarios para una metaheurística, sino también operadores de vencindad. 
 Por ejemplo, \emph{Simulated Annealing}, arrojaba mejores 

### 2015-06-01 doc/paper.tex
Working on paper

--- 2015-06-01 doc/paper.tex
@@
 
-Se seleccionaron instancias de tamaño pequeño con menos de 1000 instancias y mediano con más de 1000 pero menos de 5000 instancias.
-Cada problema cuenta con atributos y clases en el dominio de los reales.
-
-\subsection{Resultados}
-
- Se seleccionan tres casos pequeños y se observa que el porcentaje de reducción es pequeño.
-
-\begin{table}[h]
-\begin{tabular}{ |l|l|l|l| }
-    \hline
-    Caso & Instancias & Atributos & \% red. \\ \hline
-    Haberman & 306 & 4 & 13.7 \\ \hline
-    Heart & 270 & 14 & 46 \\ \hline
-    Fertility & 100 & 10 & 34 \\ \hline
-\end{tabular}
-\caption{Datos pequeños}
-\label{tabla:1}
-\end{table}
-
-    ~\
-
-    Para los casos medianos, se observa que el porcentaje de reducción es mucho mayor
-
-    ~\
-
-\begin{table}[h]
-\begin{tabular}{ |l|l|l|l| }
-    \hline
-    Caso & Instancias & Atributos & \% red. \\ \hline
-    Phishing & 2456 & 31 & 84 \\ \hline
-    Segment & 2310 & 20 & 87 \\ \hline
-    Bank & 1372 & 5 & 90 \\ \hline
-\end{tabular}
-\caption{Datos medianos}
-\label{tabla:2}
-\end{table}
-
 \subsection{Entonación}
@@
 
+Para ILS, se probó con distintas combinaciones de las iteraciones totales (denominado tiempo en la literatura) e iteraciones locales.
+Se compara con el tanaño de la solución final.
+Se observa que respecto a esta medida, la combinación de máximo 500 iteraciones totales y máximo 50 iteraciones para la búsqueda local del problema Glass.
+Pero los valores varían para otros problemas.
+
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l| }
+    \hline
+    I. total / I. local & Glass & Iris & Pima \\ \hline
+    100 / 100 & 54.0351 & 1.53846 & 31.9249 \\ \hline
+    100 / 500 & 30.6667 & 1.11111 & -- \\ \hline
+    100 / 50 & 38.7302 & 7.14286 & 28.4018 \\ \hline
+    500 / 100 & 38.3333 & 15.7143 & -- \\ \hline
+    500 / 50 & 28.5714 & 5.33333 & 25.6621 \\ \hline
+    50 / 100 & 27.7193 & 6.66667 & 26.8657 \\ \hline
+    50 / 500 & 42.5397 & 10.303 & 34.3137 \\ \hline
+    50 / 50 & 30 & 8.20513 & 35.1852 \\ \hline
+\end{tabular}
+\caption{Entonación de parámetros ILS}
+\label{tabla:1}
+\end{table}
+
+~\
+
+Para SA, se juega con la temperatura máxima inicial como manera de modificar el comportamiento de problema.
+
+~\
+
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l| }
+    \hline
+    Temp. & Glass & Iris & Pima \\ \hline
+    5000 & 35.3333 & 14.2857 & 33.2394 \\ \hline
+    10000 & 30.1587 & 7.14286 & 33.2394 \\ \hline
+    15000 & 30.7937 & 2.85714 & 31.5423 \\ \hline
+\end{tabular}
+\caption{Entonación de parámetros SA}
+\label{tabla:2}
+\end{table}
+
+~\
+
+Con TS, se juega con valores arbitrarios para el tamaño de la lista y el número de veces que se usa el operador sobre los datos en cada iteración.
+
+~\
+
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l| }
+    \hline
+    Lista / Tweaks & Glass & Iris & Pima \\ \hline
+    10 / 15 & 52.0635 & 1.90476 & 33.2381 \\ \hline
+    10 / 5 & 50.7937 & 6.66667 & 31.0502 \\ \hline
+    15 / 15 & 22.3333 & 1.888889 & 25.1852 \\ \hline
+    15 / 5 & 19.0 & 2.90476 & 33.5185 \\ \hline
+    5 / 15 & 28.5714 & 14.8718 & 35.4338 \\ \hline
+    5 / 5 & 30.4762 & 7.14286 & 39.3519 \\ \hline
+\end{tabular}
+\caption{Entonación de parámetros TS}
+\label{tabla:2}
+\end{table}
+
 \subsection{Comparación}
 
-Una vez decididos los parametros de cada metaheurística, se realizaron 
+Con las ditintas corridas se obtienen resultados promedios para cada caso y tomando en cuenta cada metaheurística.
+Específicamente para el tamaño, se observa que para el caso de Glass, la búsqueda local obtiene en promedio mejores soluciones que las metaheurísticas de trayectoria.
+Pero en Pima, que es un caso más grande, las metaheurísticas de trayectoria son capacez de superar en desempeño a la búsqueda local.
+
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+     & LS & ILS & SA & TS  \\ \hline
+    Glass & 22.6765 & 36.3244    & 32.0952 & 33.8730 \\ \hline
+    Iris  & 6.2222 & 7.0018     & 8.0952  & 5.8966 \\ \hline
+    Pima  & 43.6754 & 30.3922 & 32.6737       & 32.9629 \\ \hline
+\end{tabular}
+\caption{Resultados para tamaño de solución}
+\label{tabla:2}
+\end{table}
+
+Se realizaron 
 hasta 15 pruebas con cada una de ellas en \emph{Glass}. Los gráficos 

### 2015-06-01 doc/paper.tex

@@
 
+\subsection{Datos}
+
+Los datos utilizados fueron
+
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l| }
+    \hline
+    Caso    & Instancias & Atributos \\ \hline
+    Iris & 150 & 5 \\ \hline
+    Glass & 214 & 10 \\ \hline
+    Pima & 768 & 9 \\ \hline
+\end{tabular}
+\caption{Casos de prueba}
+\label{tabla:1}
+\end{table}
+
 \subsection{Entonación}
 
-Cada metaheuística se comporta de manera distinta. Entonar 
-los parametros de cada una de ellas requirió de una serie 
+Cada metaheuística requiere de distintos parametros para realizar
+su trabajo. Entonar cada una de ellos requirió de una serie 
 de pruebas sobre distintos problemas, parametrizando
-no solo dichos valores, sino hasta operadores de vencindad. 
+no solo valores necesarios para una metaheurística, sino también operadores de vencindad. 
 Por ejemplo, \emph{Simulated Annealing}, arrojaba mejores 
@@
     100 / 100 & 54.0351 & 1.53846 & 31.9249 \\ \hline
-    100 / 500 & 30.6667 & 1.11111 & -- \\ \hline
+    100 / 500 & 30.6667 & 1.11111 & 29.7652 \\ \hline
     100 / 50 & 38.7302 & 7.14286 & 28.4018 \\ \hline
-    500 / 100 & 38.3333 & 15.7143 & -- \\ \hline
+    500 / 100 & 38.3333 & 15.7143 & 28.9876 \\ \hline
     500 / 50 & 28.5714 & 5.33333 & 25.6621 \\ \hline
@@
 
+Los tiempos reportados para estos experimentos son
+
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l| }
+    \hline
+     & ILS & SA & TS  \\ \hline
+    Glass & 623s & 248s & 386s \\ \hline
+\end{tabular}
+\caption{Resultados para tamaño de solución}
+\label{tabla:2}
+\end{table}
+
 \begin{center}

### 2015-06-01 doc/paper.tex
Working on paper

@@
      & LS & ILS & SA & TS  \\ \hline
-    Glass & 22.6765 & 36.3244    & 32.0952 & 33.8730 \\ \hline
+    Glass & 40.6765 & 36.3244    & 32.0952 & 33.8730 \\ \hline
     Iris  & 6.2222 & 7.0018     & 8.0952  & 5.8966 \\ \hline

### 2015-06-06 doc/paper.tex
fixed errors running time

@@
 \begin{table}[h]
+\scalebox{0.6}{
 \begin{tabular}{ |l|l|l|l| }
@@
     I. total / I. local & Glass & Iris & Pima \\ \hline
-    100 / 100 & 54.0351 & 1.53846 & 31.9249 \\ \hline
-    100 / 500 & 30.6667 & 1.11111 & 29.7652 \\ \hline
-    100 / 50 & 38.7302 & 7.14286 & 28.4018 \\ \hline
-    500 / 100 & 38.3333 & 15.7143 & 28.9876 \\ \hline
-    500 / 50 & 28.5714 & 5.33333 & 25.6621 \\ \hline
-    50 / 100 & 27.7193 & 6.66667 & 26.8657 \\ \hline
-    50 / 500 & 42.5397 & 10.303 & 34.3137 \\ \hline
-    50 / 50 & 30 & 8.20513 & 35.1852 \\ \hline
+    100 / 100 & (54.0351, 46.4615) & (1.53846, 51.1922) & (31.9249, 49.1153) \\ \hline
+    100 / 50  & (38.7302, 48.4974) & (7.14286, 60.7353) & (28.4018, 49.2278) \\ \hline
+    500 / 50  & (28.5714, 49.0846) & (5.33333, 48.9383) & (25.6621, 46.8585) \\ \hline
+    50 / 100  & (27.7193, 49.9487) & (5.77778, 58.6173) & (26.8657, 49.6148) \\ \hline
+    50 / 500  & (42.5397, 49.5682) & (10.303, 50.7914)  & (34.3137, 49.4) \\ \hline
+    50 / 50   & (30, 51.5986) & (8.20513, 69.4686)      & (35.1852, 49.5211) \\ \hline
 \end{tabular}
-\caption{Entonación de parámetros ILS}
+}
+\caption{Entonación de parámetros ILS, donde la una tupla representa (\% Error de validación, \% Tamaño al que se reduce)}
 \label{tabla:1}
@@
     Temp. & Glass & Iris & Pima \\ \hline
-    5000 & 35.3333 & 14.2857 & 33.2394 \\ \hline
-    10000 & 30.1587 & 7.14286 & 33.2394 \\ \hline
-    15000 & 30.7937 & 2.85714 & 31.5423 \\ \hline
+    5000  & (35.3333, 76.1168) & (14.2857, 85.1961) & (33.2394, 50.1578) \\ \hline
+    10000 & (30.1587, 63.0743) & (7.14286, 77.2549) & (33.2394, 50.2726) \\ \hline
+    15000 & (30.7937, 55.8895) & (2.85714, 76.1275) & (31.5423, 50.3281) \\ \hline
 \end{tabular}
-\caption{Entonación de parámetros SA}
+}
+\caption{Entonación de parámetros SA, donde la una tupla representa (\% Error de validación, \% Tamaño al que se reduce)}
 \label{tabla:2}
@@
     Lista / Tweaks & Glass & Iris & Pima \\ \hline
-    10 / 15 & 52.0635 & 1.90476 & 33.2381 \\ \hline
-    10 / 5 & 50.7937 & 6.66667 & 31.0502 \\ \hline
-    15 / 15 & 22.3333 & 1.888889 & 25.1852 \\ \hline
-    15 / 5 & 19.0 & 2.90476 & 33.5185 \\ \hline
-    5 / 15 & 28.5714 & 14.8718 & 35.4338 \\ \hline
-    5 / 5 & 30.4762 & 7.14286 & 39.3519 \\ \hline
+    10 / 15        & (52.0635, 72.7087) & (0, 68.8078)  & (33.2381, 45.9503) \\ \hline
+    10 / 5         & (50.7937, 75.1986) & (6.66667, 81.6296)  & (31.0502, 48.0671) \\ \hline
+    15 / 15        & (22.3333, 70.9966) & (0.888889, 73.3827) & (25.1852, 46.7816) \\ \hline
+    15 / 5         & (19.0, 73.8488)    & (0, 82.402)  & (33.5185,  47.749) \\ \hline
+    5 / 15         & (28.5714, 74.9568) & (8.3333, 70.5314)  & (35.4338, 47.012) \\ \hline
+    5 / 5          & (30.4762, 73.3679) & (7.14286, 83.4314)  & (39.3519, 48.841) \\ \hline
 \end{tabular}
-\caption{Entonación de parámetros TS}
+}
+\caption{Entonación de parámetros TS, donde la una tupla representa (\% Error de validación, \% Tamaño al que se reduce)}
 \label{tabla:2}
@@
 \begin{table}[h]
+\scalebox{0.6}{
 \begin{tabular}{ |l|l|l|l|l| }
@@
 \end{tabular}
-\caption{Resultados para tamaño de solución}
+}
+\caption{Resultados para error de validación promedio con parámetros entonados}
 \label{tabla:2}
@@
 
-Los tiempos reportados para estos experimentos son
+Los tiempos reportados (por corrida) para estos experimentos son
 
 \begin{table}[h]
+\scalebox{0.6}{
 \begin{tabular}{ |l|l|l|l| }
@@
      & ILS & SA & TS  \\ \hline
-    Glass & 623s & 248s & 386s \\ \hline
+    Glass & 51.25 s & 74.29s & 10.8s\\ \hline
 \end{tabular}
+}
 \caption{Resultados para tamaño de solución}

### 2015-06-21 src/main.cpp
Better output to files Metaheuristic parameters now in filename of results

@@ -50,8 +50,7 @@ int main(int argc, char *argv[]) {
     memset(tweaker_optional_arg, 0, sizeof(tweaker_optional_arg));
 
     // We need these three things
-    std::string file_name, metaheuristic_str, tweaker_str;
-    std::string out_fn;
+    std::string file_name, metaheuristic_str, tweaker_str, out_fn;
 
     while (1) {
         opt = getopt_long(argc, argv, "dhf:g:m:o:x:n:r:t:", long_options, &option_index);
@@ -82,11 +81,8 @@ int main(int argc, char *argv[]) {
     if (! flag_g and ! flag_d) error_("You have to specify a metaheuristic"); 
     if (  flag_f and   flag_t) error_("You cannot use -f and -t together");
     if (! flag_x and ! flag_d) error_("You have to specify a tweaker");
-    if ((! flag_f and ! flag_d) and (! flag_t)) error_("You have to specify a file");  
+    if ((! flag_f and ! flag_d) and (!flag_t)) error_("You have to specify a file");  
     if (! flag_n and ! flag_d) runs = 1;
-
-    // TODO: Make this a parameter
-    int part = 10;
 
     Metaheuristic *metaheuristic = choose_metaheuristic(metaheuristic_str, 
                                     metaheuristic_optional_arg, 
@@ -114,6 +110,7 @@ int main(int argc, char *argv[]) {
         vps datasets;
         load_data_tenfold(file_name, datasets);
         results res = run_tenfold(datasets, metaheuristic, runs);
+        output_results(res, metaheuristic, flag_r, out_fn);
     }
 
     if (flag_d) {

### 2015-06-21 src/metaheuristic.cpp

@@ -114,6 +114,10 @@ void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) const {
     }
 }
 
+std::string HillClimbing::output_params() const {
+    return "HillC_";
+}
+
 /* Simulated Annealing  */
 
 void SimulatedAnnealing::optimize(const IS::Dataset &T, IS::Solution &S) const {
@@ -155,6 +159,12 @@ void SimulatedAnnealing::optimize(const IS::Dataset &T, IS::Solution &S) const {
 
         if (qb > max_quality || t <= EPSILON) break;
     }
+}
+
+std::string SimulatedAnnealing::output_params() const {
+    std::string params = "SimulAnn_" + std::to_string(temperature) + "_";
+    params += std::to_string(dec_factor);
+    return params;
 }
 
 /* ILS */
@@ -214,6 +224,14 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
     }
 }
 
+std::string ILS::output_params() const {
+    std::string params = "ILS_" + std::to_string(ti) + "_";
+    params += std::to_string(li) + "_";
+    params += std::to_string(pp);
+    return params;
+}
+
+
 /* Tabu */
 
 void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) const {
@@ -270,4 +288,10 @@ void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) const {
 
         old_q = q_max;
     }
+}
+
+std::string Tabu::output_params() const {
+    std::string params = "Tabu_" + std::to_string(length) + "_";
+    params += std::to_string(number_of_tweaks);
+    return params;
 }

### 2015-06-21 src/metaheuristic.h

@@ -13,6 +13,7 @@ class Metaheuristic {
     double max_quality = 0.95; 
   public:
     virtual void optimize(const IS::Dataset &T, IS::Solution &R) const = 0;
+    virtual std::string output_params() const = 0;
     void setTweaker(Tweaker *tw) { tweaker = tw; };
     Tweaker* getTweaker() const { return tweaker; };
     double quality(const IS::Dataset &T, const IS::Solution &S, double alpha) const;
@@ -28,6 +29,7 @@ class Metaheuristic {
 class HillClimbing : public Metaheuristic {
   public: 
     void optimize(const IS::Dataset &T, IS::Solution &R) const;
+    std::string output_params() const;
 };
 
 class SimulatedAnnealing : public Metaheuristic {
@@ -36,6 +38,7 @@ class SimulatedAnnealing : public Metaheuristic {
   public: 
     SimulatedAnnealing(int t, int d): temperature(t), dec_factor(d) { };
     void optimize(const IS::Dataset &T, IS::Solution &R) const;
+    std::string output_params() const;
 };
 
 class ILS : public Metaheuristic {
@@ -50,6 +53,7 @@ class ILS : public Metaheuristic {
     void setTotalIter(int t) { ti = t; }
     void setPerturbPerc(int p) { pp = p; }
     void optimize(const IS::Dataset &T, IS::Solution &R) const;
+    std::string output_params() const;
 };
 
 class Tabu : public Metaheuristic {
@@ -58,6 +62,7 @@ class Tabu : public Metaheuristic {
   public:
     Tabu(int l, int n) : length(l), number_of_tweaks(n) { };
     void optimize(const IS::Dataset &T, IS::Solution &R) const;
+    std::string output_params() const;
 };
 
 

### 2015-06-21 src/tweaker.cpp

@@ -43,7 +43,6 @@ void percRandomFlips::tweak(IS::Solution &S) {
 
     for (int i = 0; i < n; ++i) {
         int j = (rand() % size);
-        cout << "j = " << j << endl;
         bits.flip(j);
     }
     S.setBits(bits);

### 2015-06-21 src/utils.cpp

@@ -1,5 +1,6 @@
 #include <algorithm>
 #include <iostream>
+#include <fstream>
 #include <cstdio>
 #include "metaheuristic.h"
 #include "utils.h"
@@ -233,6 +234,7 @@ results run_tenfold(const vps &datasets, const Metaheuristic *metaheuristic, int
 
     return res;
 
+
     // std::ostream *out;
    // std::ofstream rFile;
 
@@ -297,4 +299,37 @@ IS::Dataset load_dataset(const char *filename) {
     }
 
     return dataset;
+}
+
+void output_results(const results &res, 
+                    const Metaheuristic *metaheuristic,
+                    bool flag_r,
+                    std::string out_fn) {
+
+    std::ostream *out;
+    std::ofstream out_file;
+    
+    if(flag_r) {
+        out_fn += "_" + metaheuristic->output_params();
+        out_file.open(out_fn, std::ios::out);
+        out = &out_file;
+    } else {
+       out = &std::cout;
+    }
+
+    *out << "Val_errors" << std::endl;
+    for (unsigned i = 0; i < res.val_errors.size(); ++i)  {
+       *out << res.val_errors[i] << " ";
+    }
+    *out << std::endl;
+    *out << "Sizes" << std::endl;
+    for (unsigned i = 0; i < res.sizes.size(); ++i)  {
+       *out << res.sizes[i] << " ";
+    }
+    *out << std::endl;
+    *out << "Errors" << std::endl;
+    for (unsigned i = 0; i < res.errors.size(); ++i)  {
+       *out << res.errors[i] << " ";
+    }
+    *out << std::endl;
 }

### 2015-06-21 src/utils.h

@@ -36,3 +36,9 @@ results run_tenfold(const vps &datasets,
                  int runs);
 
 IS::Dataset load_dataset(const char *filename);
+
+void output_results(const results &res, 
+                    const Metaheuristic *metaheuristic,
+                    bool flag_r,
+                    std::string out_fn);
+

### 2015-06-21 src/main.cpp
Fix en glass Other stuff

@@ -105,7 +105,7 @@ int main(int argc, char *argv[]) {
         IS::Dataset dataset = load_data_basic(file_name);
         // Initial random solution
         IS::Solution solution(dataset.size());
-        solution.generateRandom(); 
+        // solution.generateRandom(); 
         metaheuristic->optimize(dataset, solution);
     } else if (flag_t) {
         vps datasets;

### 2015-06-21 src/utils.cpp

@@ -197,7 +197,11 @@ results run_tenfold(const vps &datasets, const Metaheuristic *metaheuristic, int
     results res;
     IS::Dataset training, validation;
 
+    cout << "Datasets: " << datasets.size() << endl;
+    cout << "Runs: " << runs << endl;
+
     for(int i = 0; i < 10; ++i) {
+        cout << BOLDGREEN << "Running fold" << RESET << endl;
         training = datasets[i].first;
         validation = datasets[i].second;
 
@@ -222,9 +226,9 @@ results run_tenfold(const vps &datasets, const Metaheuristic *metaheuristic, int
                 if (validation[k].getCategory() != categories[k])
                     errors++;
 
-            double clas_rate = (1.0 * errors) / (1.0 * validation.size());
-            double perc_redc = (1.0 * final.size()) / (1.0 * training.size());
-            double fitness = 0.5 * clas_rate + (1 - 0.5) * perc_redc;
+            // double clas_rate = (1.0 * errors) / (1.0 * validation.size());
+            // double perc_redc = (1.0 * final.size()) / (1.0 * training.size());
+            // double fitness = 0.5 * clas_rate + (1 - 0.5) * perc_redc;
 
             res.errors.push_back(errors);
             res.val_errors.push_back((errors * 100.0) / (validation.size() * 1.0));
@@ -310,7 +314,7 @@ void output_results(const results &res,
     std::ofstream out_file;
     
     if(flag_r) {
-        out_fn += "_" + metaheuristic->output_params();
+        out_fn += "_" + metaheuristic->output_params() + ".result";
         out_file.open(out_fn, std::ios::out);
         out = &out_file;
     } else {

### 2015-06-21 src/utils.h

@@ -1,4 +1,21 @@
 #include "is.h"
+#define RESET   "\033[0m"
+#define BLACK   "\033[30m"      /* Black */
+#define RED     "\033[31m"      /* Red */
+#define GREEN   "\033[32m"      /* Green */
+#define YELLOW  "\033[33m"      /* Yellow */
+#define BLUE    "\033[34m"      /* Blue */
+#define MAGENTA "\033[35m"      /* Magenta */
+#define CYAN    "\033[36m"      /* Cyan */
+#define WHITE   "\033[37m"      /* White */
+#define BOLDBLACK   "\033[1m\033[30m"      /* Bold Black */
+#define BOLDRED     "\033[1m\033[31m"      /* Bold Red */
+#define BOLDGREEN   "\033[1m\033[32m"      /* Bold Green */
+#define BOLDYELLOW  "\033[1m\033[33m"      /* Bold Yellow */
+#define BOLDBLUE    "\033[1m\033[34m"      /* Bold Blue */
+#define BOLDMAGENTA "\033[1m\033[35m"      /* Bold Magenta */
+#define BOLDCYAN    "\033[1m\033[36m"      /* Bold Cyan */
+#define BOLDWHITE   "\033[1m\033[37m"      /* Bold White */
 
 typedef IS::Dataset ds;
 typedef std::pair<ds, ds> pds;

### 2015-06-21 src/main.cpp
added debugging flag and method

@@ -31,6 +31,7 @@ int main(int argc, char *argv[]) {
         {"tweaker-optional-arg", required_argument, 0, 'o'},
         {"tweaker", required_argument, 0, 'x'},
         {"results", required_argument, 0, 'r'},
+        {"debugging", no_argument, 0, 'a'},
         {0, 0, 0, 0}
     };
 
@@ -46,6 +47,7 @@ int main(int argc, char *argv[]) {
     bool flag_x = false;
     bool flag_r = false;
     bool flag_t = false;
+    bool flag_a = false;
     memset(metaheuristic_optional_arg, 0, sizeof(metaheuristic_optional_arg));
     memset(tweaker_optional_arg, 0, sizeof(tweaker_optional_arg));
 
@@ -71,6 +73,7 @@ int main(int argc, char *argv[]) {
             case 'r': flag_r = true; out_fn = optarg; break;
             case 'd': flag_d = true; break;
             case 't': flag_t = true; file_name = optarg; break;
+            case 'a': flag_a = true;  break;
             case '?': break;
         }
     }
@@ -107,6 +110,7 @@ int main(int argc, char *argv[]) {
         IS::Solution solution(dataset.size());
         solution.generateRandom(); 
         metaheuristic->optimize(dataset, solution);
+        if (flag_a) debug_print(dataset, solution);
     } else if (flag_t) {
         vps datasets;
         load_data_tenfold(file_name, datasets);

### 2015-06-21 src/utils.cpp

@@ -70,6 +70,7 @@ void print_help() {
     std::cout << "-h, --help\t\t\t\tPrint this help" << std::endl;  
     std::cout << "-f, --file\t-Needed-\t\tExpect a file name" << std::endl; 
     std::cout << "-d, --disp\t\t\t\tPrint dispersion vector" << std::endl; 
+    std::cout << "-a, --debugging\t\t\t\tPrint for debugging" << std::endl;  
 
     // Metaheuristic
     std::cout << "-g, --metaheuristic [NAME]     -Needed-\tSpecify metaheuristic"; 
@@ -332,4 +333,34 @@ void output_results(const results &res,
        *out << res.errors[i] << " ";
     }
     *out << std::endl;
+}
+
+void debug_print(const IS::Dataset &dataset, const IS::Solution &solution) {
+   HillClimbing meta;
+   std::cout << "----------------------------------------- " << std::endl;
+   meta.quality(dataset, solution, 0.5);
+   std::cout << "----------------------------------------- " << std::endl;
+   std::cout << "Imprimiendo en archivos respectivos" << std::endl;
+   assert(dataset.size() == solution.getSize());
+   std::string sol_name = "scripts/solution.txt", comp_name = "scripts/complement.txt";
+
+   std::ofstream sol_out, comp_out; 
+
+   // Opening files
+   sol_out.open(sol_name, std::ios::out);
+   comp_out.open(comp_name, std::ios::out);
+
+   // Printing to files
+   std::bitset<MAX> bits = solution.getBits();
+   for (int i = 0; i < dataset.size(); i++) {
+      if (bits.test(i)) sol_out << dataset[i].toString() << std::endl;
+      else comp_out << dataset[i].toString() << std::endl;
+   }
+   std::cout << "Se imprimieron " << to_string(bits.count()) << " lineas en ";
+   std::cout << "solution.txt" << std::endl;
+
+   std::cout << "Se imprimieron " << to_string(dataset.size() - bits.count()); 
+   std::cout << " lineas en complement.txt" << std::endl;;
+   sol_out.close();
+   comp_out.close();
 }

### 2015-06-21 src/utils.h

@@ -42,3 +42,4 @@ void output_results(const results &res,
                     bool flag_r,
                     std::string out_fn);
 
+void debug_print(const IS::Dataset &dataset, const IS::Solution &solution);

### 2015-06-21 src/scripts/validator.hs
validator

@@ -1,0 +1,35 @@
+{-# LANGUAGE ScopedTypeVariables #-}
+module Main (main) where
+
+import Control.Monad
+import System.IO
+import System.Environment
+import Data.Functor
+import Data.Function
+import Data.List
+
+type Instance = ([Double], Int)
+
+findNearest :: [Instance] -> Instance -> Int
+findNearest set inst = snd . head $ sort $ map (\x -> (euclidean inst x, snd x)) set
+
+euclidean :: Instance -> Instance -> Double
+euclidean (a,_) (b,_) = sqrt . sum $ zipWith (*) z z
+    where z = zipWith (-) a b
+
+knn :: [Instance] -> [Instance] -> [Int]
+knn solution complement = map (findNearest solution) complement
+
+corrects :: [Instance] -> [Int] -> Int
+corrects set cat = foldl' count 0 $ zipWith (-) cat (map snd set)
+    where count = (\acc x -> if (x == 0) then acc + 1 else acc)
+
+loadData :: String -> IO [Instance]
+loadData fileName = readFile fileName >>= return . lines >>= return . (map words) >>= return . toDouble >>= return . (map toInstance)
+    where toDouble       = (map (map (\x -> read x :: Double)))
+          toInstance xs  = (init xs, (fromInteger $ floor $ last xs) :: Int)
+
+main = do
+    solution   <- loadData "solution.txt"
+    complement <- loadData "complement.txt"
+    putStrLn $ show $ corrects complement $ knn solution complement 

### 2015-06-22 src/scripts/validator.hs
validator

@@ -7,6 +7,7 @@ import System.Environment
 import Data.Functor
 import Data.Function
 import Data.List
+import Numeric
 
 type Instance = ([Double], Int)
 
@@ -29,7 +30,15 @@ loadData fileName = readFile fileName >>= return . lines >>= return . (map words
     where toDouble       = (map (map (\x -> read x :: Double)))
           toInstance xs  = (init xs, (fromInteger $ floor $ last xs) :: Int)
 
+percReduc :: [a] -> [a] -> Double
+percReduc a b = (fromIntegral lb) / (fromIntegral $ la + lb)
+    where la = length a
+          lb = length b
+
 main = do
     solution   <- loadData "solution.txt"
     complement <- loadData "complement.txt"
-    putStrLn $ show $ corrects complement $ knn solution complement 
+    putStrLn $ ">>>> Aciertos = " ++  (show $ corrects complement $ knn solution complement)
+    putStrLn $ "Tamaño de solucion = " ++ (show $ length solution)
+    putStrLn $ "Tamaño de T = " ++ (show $ (length solution) + (length complement))
+    putStrLn $ "Porcentaje de reduccion = " ++  (show $ percReduc solution complement)

### 2015-06-22 src/metaheuristic.cpp
ILS stop condition is fixed

@@ -63,13 +63,13 @@ double Metaheuristic::quality(const IS::Dataset &T,
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
 
     // // XXX: Print
-     // cout << "Hubo un total de " << count << " aciertos" << endl;
-     // cout << "Result es de tamanio: " << result.size() << endl;
-     // cout << "La relación es: " << clas_rate << endl;
-     // cout << "El tamaño de T es: " << T.size() << endl;
-     // cout << "El tamaño de training es: " << training.size() << endl;
-     // cout << "El porcentaje de reducción es: " << perc_redc << endl;
-     // cout << "El fitness es : " << fitness << endl << endl;
+      cout << "Hubo un total de " << count << " aciertos" << endl;
+      cout << "Result es de tamanio: " << result.size() << endl;
+      cout << "La relación es: " << clas_rate << endl;
+      cout << "El tamaño de T es: " << T.size() << endl;
+      cout << "El tamaño de training es: " << training.size() << endl;
+      cout << "El porcentaje de reducción es: " << perc_redc << endl;
+      cout << "El fitness es : " << fitness << endl << endl;
 
     assert(fitness <= 1.0);
     return fitness;
@@ -172,19 +172,21 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
     IS::Solution best(S);
     IS::Solution home_base(S);
 
-    double q_max, old_q = -1;
-    int iter_old = 0, iter = 0;
+    double q_max, q_BEST = -1.0;
+    int global_iter = 0;
     
     // TODO: Tune this
     int max_out_iter = getTotalIter();
+    int max_iter = getLocalIter(); 
+    std::cout << "local iter = " << max_iter << endl;
 
     while(1) {
         // Local search
 
         // TODO: Ideally some tuneable number of iterations
         // 500 to 1000 random comes to mind
-        int max_iter = getLocalIter(); 
         cout << ">>> Running local search" << endl;
+        int local_iter = 0;
         while(1) {
             IS::Solution R(S);
             tweaker->tweak(R); 
@@ -193,13 +195,13 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
                 S.setBits(R.getBits());
 
             q_max = std::max(q1, q2);
-            if (old_q == q_max) {
-              iter_old ++;  
-            } else {
-                iter_old = 0;
[…]

### 2015-06-22 src/metaheuristic.cpp
tabu stop condition

@@ -63,13 +63,13 @@ double Metaheuristic::quality(const IS::Dataset &T,
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
 
     // // XXX: Print
-      cout << "Hubo un total de " << count << " aciertos" << endl;
-      cout << "Result es de tamanio: " << result.size() << endl;
-      cout << "La relación es: " << clas_rate << endl;
-      cout << "El tamaño de T es: " << T.size() << endl;
-      cout << "El tamaño de training es: " << training.size() << endl;
-      cout << "El porcentaje de reducción es: " << perc_redc << endl;
-      cout << "El fitness es : " << fitness << endl << endl;
+      //cout << "Hubo un total de " << count << " aciertos" << endl;
+      //cout << "Result es de tamanio: " << result.size() << endl;
+      //cout << "La relación es: " << clas_rate << endl;
+      //cout << "El tamaño de T es: " << T.size() << endl;
+      //cout << "El tamaño de training es: " << training.size() << endl;
+      //cout << "El porcentaje de reducción es: " << perc_redc << endl;
+      //cout << "El fitness es : " << fitness << endl << endl;
 
     assert(fitness <= 1.0);
     return fitness;
@@ -238,16 +238,14 @@ void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) const {
     std::cout << ">>>> Running Tabu" << std::endl;
     std::cout << ">>>> quality = " << max_quality << std::endl;
     IS::Solution S(best);
-    S.generateRandom();  // First random solution
     int max_length = (length * T.size() / 100);
 
-    std::deque<IS::Solution> tabu_list;   // Tabu list
-    tabu_list.push_back(S);
-    double q_max, old_q = -1;
-    int iter_old = 0, iter_total = 0;
+    std::deque<IS::Solution> tl;   // Tabu list
+    tl.push_back(S);
+    double q_max, q_BEST = -1;
+    int iter_old = 0;
     while (1) {
-        iter_total++;
-        if (tabu_list.size() > max_length) tabu_list.pop_front();
+        while (tl.size() > max_length) tl.pop_front();
         IS::Solution R(S);
         tweaker->tweak(R); 
 
@@ -255,37 +253,33 @@ void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) const {
             IS::Solution W(S); 
             tweaker->tweak(W);
 
-            // Checking if W is in tabu list
-            std::deque<IS::Solution>::iterator it;
[…]

### 2015-06-22 src/metaheuristic.h

@@ -58,9 +58,9 @@ class ILS : public Metaheuristic {
 
 class Tabu : public Metaheuristic {
   private:
-    int length, number_of_tweaks;
+    int length, number_of_tweaks, iter_limit;
   public:
-    Tabu(int l, int n) : length(l), number_of_tweaks(n) { };
+    Tabu(int l, int n, int i) : length(l), number_of_tweaks(n), iter_limit(i) { };
     void optimize(const IS::Dataset &T, IS::Solution &R) const;
     std::string output_params() const;
 };

### 2015-06-22 src/utils.cpp

@@ -146,10 +146,11 @@ Metaheuristic *choose_metaheuristic(std::string metaheuristic_str,
         return new SimulatedAnnealing(t, d);
     } else if (metaheuristic_str == "Tabu") {
         // Print error and exit
-        if (moa_counter != 2) error_("Tabu needs two arguments");
+        if (moa_counter != 3) error_("Tabu needs three arguments");
         int l = metaheuristic_optional_arg[0]; 
         int n = metaheuristic_optional_arg[1];
-        return new Tabu(l, n);
+        int i = metaheuristic_optional_arg[2];
+        return new Tabu(l, n, i);
     } else if (metaheuristic_str == "ILS") {
         
         // Print error and exit

### 2015-06-22 src/utils.cpp
Added time calculation

@@ -2,6 +2,7 @@
 #include <iostream>
 #include <fstream>
 #include <cstdio>
+#include <ctime>
 #include "metaheuristic.h"
 #include "utils.h"
 
@@ -208,6 +209,8 @@ results run_tenfold(const vps &datasets, const Metaheuristic *metaheuristic, int
         validation = datasets[i].second;
 
         for(int j = 0; j < runs; ++j) {
+            clock_t start = clock();
+            double duration;
             IS::Solution solution(training.size());
             metaheuristic->optimize(training, solution);
 
@@ -220,6 +223,8 @@ results run_tenfold(const vps &datasets, const Metaheuristic *metaheuristic, int
             for(int k = 0; k < training.size(); ++k) {
                 if(bits[k]) final.push_back(training[k]);
             }
+
+            duration = ( std::clock() - start ) / (double) CLOCKS_PER_SEC;
 
             metaheuristic->oneNN(final, validation, categories);
 
@@ -235,6 +240,7 @@ results run_tenfold(const vps &datasets, const Metaheuristic *metaheuristic, int
             res.errors.push_back(errors);
             res.val_errors.push_back((errors * 100.0) / (validation.size() * 1.0));
             res.sizes.push_back((final.size() * 100.0) / (training.size() * 1.0));
+            res.times.push_back(duration);
         }
     }
 
@@ -338,6 +344,12 @@ void output_results(const results &res,
        *out << res.errors[i] << " ";
     }
     *out << std::endl;
+    *out << "Times" << std::endl;
+    for (unsigned i = 0; i < res.times.size(); ++i)  {
+       *out << res.times[i] << " ";
+    }
+    *out << std::endl;
+
 }
 
 void debug_print(const IS::Dataset &dataset, const IS::Solution &solution) {

### 2015-06-22 src/utils.h

@@ -25,6 +25,7 @@ typedef struct results {
     std::vector<int> errors;
     std::vector<float> sizes;
     std::vector<float> val_errors;
+    std::vector<double> times;
 } result;
 
 

### 2015-06-22 src/is.h
*Population based added *Parametrized max iterations in meteheuristic *Fitness as a member of Solution *< operator on Solution

@@ -28,12 +28,15 @@ namespace IS {
       private:
         std::bitset<MAX> bits;
         int size;
+        double fitness;
       public:
         Solution(int N) : size(N) { generateRandom(); };
         Solution(std::bitset<MAX> a, int N) : bits(a), size(N) { };
         Solution(const Solution &a) : bits(a.getBits()), size(a.getSize()) { };
         int getSize() const { return size; };
         void setSize(int n) { size = n; };
+        void setFitness(double f) { fitness = f; };
+        double getFitness() const { return fitness; };
         std::bitset<MAX> getBits() const { return bits; };
         void setBits(const std::bitset<MAX> &a) { bits = a; }
         void generateRandom(); 
@@ -46,6 +49,15 @@ namespace IS {
         void copy(const Solution &a);
         bool operator==(const Solution &b) const {
             return (size == b.size && bits == b.bits);
+        }
+        bool operator<(const Solution& y) const {
+            const IS::Solution &x = *this;
+            std::bitset<MAX> bx = x.getBits();
+            std::bitset<MAX> by = y.getBits();
+            for (int i = size; i >= 0; i--) {
+                if (bx[i] ^ by[i]) return by[i];
+            }
+            return false;
         }
     };
 

### 2015-06-22 src/metaheuristic.cpp

@@ -200,7 +200,7 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
                 local_iter = 0;
             } else local_iter++;
 
-            std::cout << "ITER ---XXXXXX " << local_iter << std::endl;
+            // std::cout << "ITER ---XXXXXX " << local_iter << std::endl;
             if (q_max > max_quality or local_iter == max_iter) break;
         }
 
@@ -211,10 +211,10 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
         if (qs > qb) best.setBits(S.getBits());
         // Flipping 25% of bits for perturb
         // TODO: This CAN be an argument
-        cout << ">>> Perturbing <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<" << endl;
+        // cout << ">>> Perturbing <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<" << endl;
         //weightedRandomPlus tweaker = weightedRandomPlus(perc, 50);
         int perc = S.getSize() / getPerturbPerc();
-        cout << "perc ---> " << perc << endl;
+        // cout << "perc ---> " << perc << endl;
         nRandomFlips tweaker = nRandomFlips(perc);
         tweaker.tweak(S);
 

### 2015-06-22 src/populationbased.h

@@ -1,0 +1,36 @@
+#ifndef POPULATIONBASED_H
+#define POPULATIONBASED_H
+
+#include <set>
+#include "metaheuristic.h"
+#include "is.h"
+
+typedef pair<IS::Solution, IS::Solution> Children;
+typedef set<IS::Solution> Population;
+using namespace std;
+
+class PopulationBased : public Metaheuristic {
+  protected:
+    int popsize;
+    Metaheuristic *local_search;
+    Tweaker *mutation;
+  public:
+    PopulationBased(int p, Metaheuristic *ls) : popsize(p), local_search(ls) { };
+    set<IS::Solution> breed(const set<IS::Solution> &p) const;
+    pair<IS::Solution, IS::Solution> recombination(const IS::Solution &sa, 
+                                                   const IS::Solution &sb) 
+                                                   const;
+    void selection(set<IS::Solution> &sa, const set<IS::Solution> &sb);
+    set<IS::Solution> generatePopulation(int size) const;
+    IS::Solution tourneySelection(const set<IS::Solution> &p, int s) const;
+};
+
+class Hybrid : public PopulationBased {
+  public:
+    void optimize(const IS::Dataset &ds, IS::Solution &sol) const;
+    std::string output_params() const { return ""; };
+    Hybrid(int p, Metaheuristic *ls): PopulationBased(p, ls) { };
+    // set<IS::Solution> breed(set<IS::Solution> &p);
+};
+
+#endif

### 2015-06-22 doc/paper.tex
paper

@@
 
-\title{Selección de Instancias}
+\title{Comparación de Metaheurísticas y Algoritmos Poblacionales
+       para Selección de Instancias}
 
@@
 \maketitle
+\section{Introducción}
+  En este artículo se considera el problema de Selección de Instancias, como
+  estrategia de reducción de datos para Minería de Datos. Dado un conjunto de
+  instancias de datos, el problema de 
+  \textit{Selección de Instancias} (IS por sus siglas en inglés), busca
+  escoger un subconjunto de instancias de menor cardinalidad que el conjunto original
+  que mantenga o mejore la clasificación de las muestras a ser usadas como un conjunto
+  de entrenamiento.
+
+  En el presente artículo se implementan 3 Metaheurísticas de trayectorias 
+  (Hill CLimb, ILS, Tabú) y 2 Metaheurísticas poblacionales (Algoritmo Memético y 
+  Algoritmo Genético), mostrando para cada una de ellas los resultados 
+  y tiempos de corrida al aplicarlas sobre diferentes conjuntos de datos.
+
+  Este artículo se organiza en 6 secciones. Se comienza con una definición
+  formal del problema, en donde se explica la terminología relevante a SI. 
+  Posteriormente, se describe la representación de las soluciones, así 
+  como también el proceso de generación de soluciones iniciales. Se continúa
+  con los algoritmos y metaheurísticas usadas, mostrando el pseudocódigo de cada
+  una de ellas. Seguido a esto, se describen cada uno de los conjuntos de datos usados
+  en los experimentos. Finalmente, se presentan los resultados experimentales y se 
+  concluye en base a dicho resultados.
+
 \begin{abstract}
-  El problema de \textit{Selección de Instancias} (IS por sus siglas en inglés), busca
+  El problema de \textit{Selección de Instancias}  busca
   escoger un subconjunto de instancias de menor cardinalidad del conjunto original

### 2015-06-22 src/is.cpp
Memetic

@@ -8,7 +8,6 @@
 #include "is.h"
 
 void IS::Solution::generateRandom() {
-    srand(time(NULL));
     for (int i = 0; i < size; i++) {
         int p = rand() % 100; 
         this->bits[i] = (p > 10) ? 1 : 0;

### 2015-06-22 src/main.cpp

@@ -21,6 +21,9 @@ int moa_counter = 0;
 int tweaker_optional_arg[10];
 int toa_counter = 0;
 
+int population_optional_arg[10];
+int poa_counter = 0;
+
 int main(int argc, char *argv[]) {
 
     static struct option long_options[] = {
@@ -28,6 +31,7 @@ int main(int argc, char *argv[]) {
         {"file", required_argument, 0, 'f'},
         {"disp", no_argument, 0, 'd'},
         {"metaheuristic", required_argument, 0, 'g'},
+        {"population-base", required_argument, 0, 'p'},
         {"meta-optional-arg", required_argument, 0, 'm'},
         {"tweaker-optional-arg", required_argument, 0, 'o'},
         {"tweaker", required_argument, 0, 'x'},
@@ -49,14 +53,16 @@ int main(int argc, char *argv[]) {
     bool flag_r = false;
     bool flag_t = false;
     bool flag_a = false;
+    bool flag_P = false;
     memset(metaheuristic_optional_arg, 0, sizeof(metaheuristic_optional_arg));
     memset(tweaker_optional_arg, 0, sizeof(tweaker_optional_arg));
+    memset(population_optional_arg, 0, sizeof(population_optional_arg));
 
     // We need these three things
-    std::string file_name, metaheuristic_str, tweaker_str, out_fn;
+    std::string file_name, metaheuristic_str, tweaker_str, out_fn, population_str;
 
     while (1) {
-        opt = getopt_long(argc, argv, "dhf:g:m:o:x:n:r:t:", long_options, &option_index);
+        opt = getopt_long(argc, argv, "dhf:g:m:o:x:n:r:t:p:P:", long_options, &option_index);
         if (opt == -1) break;
         opt = char(opt);
 
@@ -75,6 +81,9 @@ int main(int argc, char *argv[]) {
             case 'd': flag_d = true; break;
             case 't': flag_t = true; file_name = optarg; break;
             case 'a': flag_a = true;  break;
+            case 'P': flag_P = true; population_str = optarg; break;
+            case 'p': population_optional_arg[poa_counter++] = atoi(optarg); 
+                      break;
             case '?': break;
         }
     }
@@ -104,6 +113,15 @@ int main(int argc, char *argv[]) {
     }
 
     metaheuristic->setTweaker(tweaker); // Setting tweaker
+
+    if (flag_P) {
+        Metaheuristic *ls = metaheuristic;
+        metaheuristic = choose_population(population_str, ls, 
[…]

### 2015-06-22 src/metaheuristic.cpp

@@ -63,13 +63,13 @@ double Metaheuristic::quality(const IS::Dataset &T,
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
 
     // // XXX: Print
-      //cout << "Hubo un total de " << count << " aciertos" << endl;
-      //cout << "Result es de tamanio: " << result.size() << endl;
-      //cout << "La relación es: " << clas_rate << endl;
-      //cout << "El tamaño de T es: " << T.size() << endl;
-      //cout << "El tamaño de training es: " << training.size() << endl;
-      //cout << "El porcentaje de reducción es: " << perc_redc << endl;
-      //cout << "El fitness es : " << fitness << endl << endl;
+      cout << "Hubo un total de " << count << " aciertos" << endl;
+      cout << "Result es de tamanio: " << result.size() << endl;
+      cout << "La relación es: " << clas_rate << endl;
+      cout << "El tamaño de T es: " << T.size() << endl;
+      cout << "El tamaño de training es: " << training.size() << endl;
+      cout << "El porcentaje de reducción es: " << perc_redc << endl;
+      cout << "El fitness es : " << fitness << endl << endl;
 
     assert(fitness <= 1.0);
     return fitness;
@@ -176,7 +176,7 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
     int global_iter = 0;
     
     // TODO: Tune this
-    int max_out_iter = getTotalIter();
+    int max_out_iter = iterations;
     int max_iter = getLocalIter(); 
     std::cout << "local iter = " << max_iter << endl;
 
@@ -225,7 +225,7 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
 }
 
 std::string ILS::output_params() const {
-    std::string params = "ILS_" + std::to_string(ti) + "_";
+    std::string params = "ILS_" + std::to_string(iterations) + "_";
     params += std::to_string(li) + "_";
     params += std::to_string(pp);
     return params;
@@ -279,7 +279,7 @@ void Tabu::optimize(const IS::Dataset &T, IS::Solution &best) const {
             iter_old = 0;
         } else iter_old++;  
 
-        if (q_max > max_quality || iter_old == iter_limit) break;
+        if (q_max > max_quality || iter_old == iterations) break;
     }
 }
 

### 2015-06-22 src/populationbased.cpp

@@ -1,55 +1,77 @@
 #include "metaheuristic.h"
 #include "populationbased.h"
+#include "utils.h"
 
 
-Population PopulationBased::generatePopulation(int size) const {
-    Population population;
-    for(int i = 0; i < popsize; ++i) population.insert(IS::Solution(size));
-    return population;
+void PopulationBased::generatePopulation(Population &population, int size) const {
+    while (population.empty() || population.size() < popsize)
+        population.insert(IS::Solution(size));
 }
 
 
 void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
-    Population population = generatePopulation(sol.getSize());
+    std::cout << ">>>> Running Hybrid algorithm" << std::endl;
+    Population population;
+    generatePopulation(population, sol.getSize());
     IS::Solution best = IS::Solution(*(population.begin()));
-    best.setFitness(quality(ds, best, 0.5));
+    double best_fitness = quality(ds, best, 0.5);
+    best.setFitness(best_fitness);
     int iter = 0, iter_old = 0;
-    double old_q = -1.0;
+    double old_best = -1.0;
 
     while(1) {
         iter++;
         Population::const_iterator sit;
+        local_search->set_max_quality(std::min(best.getFitness() + 0.1, 0.95));
+        Population new_pop;
         for (sit = population.begin(); sit != population.end(); sit++) {
             IS::Solution p = IS::Solution(*sit);
             local_search->optimize(ds, p);
-            p.setFitness(quality(ds, p, 0.5));
-            if (p.getFitness() > best.getFitness()) {
+            double p_fitness = quality(ds, p, 0.5);
+            p.setFitness(p_fitness);
+            new_pop.insert(p);
+            if (p_fitness > best_fitness) {
                 best.copy(p);
+                best_fitness = p_fitness;
             }
         }
-        if (best.getFitness() - old_q < EPSILON) iter_old++;
+        if (best_fitness - old_best < EPSILON) {
+            for (int i = 0; i < 10; i++) {
+                std::cout << BOLDGREEN << "HHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHH";
+                std::cout  << RESET << std::endl;
+            }
+            iter_old++;
+        }
         else {
-            old_q = best.getFitness();
+            std::cout << "BEST = " << best_fitness << " , old_best = " << old_best << endl;
[…]

### 2015-06-22 src/populationbased.h

@@ -13,15 +13,14 @@ class PopulationBased : public Metaheuristic {
   protected:
     int popsize;
     Metaheuristic *local_search;
-    Tweaker *mutation;
   public:
     PopulationBased(int p, Metaheuristic *ls) : popsize(p), local_search(ls) { };
     set<IS::Solution> breed(const set<IS::Solution> &p) const;
     pair<IS::Solution, IS::Solution> recombination(const IS::Solution &sa, 
                                                    const IS::Solution &sb) 
                                                    const;
-    void selection(set<IS::Solution> &sa, const set<IS::Solution> &sb);
-    set<IS::Solution> generatePopulation(int size) const;
+    Population selection(const Population &sa, const Population &sb) const;
+    void generatePopulation(Population &p, int size) const;
     IS::Solution tourneySelection(const set<IS::Solution> &p, int s) const;
 };
 

### 2015-06-22 src/utils.cpp

@@ -4,6 +4,7 @@
 #include <cstdio>
 #include <ctime>
 #include "metaheuristic.h"
+#include "populationbased.h"
 #include "utils.h"
 
 IS::Dataset load_data_basic(std::string file_name_str) {
@@ -133,6 +134,21 @@ void print_dispersions(const IS::Dataset &dataset) {
     std::cout << endl;
 }
 
+Metaheuristic *choose_population(std::string population_str, 
+                                   Metaheuristic *local_search,
+                                   int (&population_optional_arg)[10],
+                                   int poa_counter) {
+
+   local_search->set_no_change_best(50);
+   PopulationBased *pb; 
+   if (population_str == "Hybrid") {
+      if (poa_counter != 1) error_("Hybrid needs Population size as argument");
+      return new Hybrid(population_optional_arg[0], local_search);
+   }
+   error_("The Population based algorithm you are providing doesn't exist");
+   return NULL;   // Shut up warning
+}
+
 Metaheuristic *choose_metaheuristic(std::string metaheuristic_str, 
                                     int (&metaheuristic_optional_arg)[10],
                                     int moa_counter) {
@@ -245,46 +261,6 @@ results run_tenfold(const vps &datasets, const Metaheuristic *metaheuristic, int
     }
 
     return res;
-
-
-    // std::ostream *out;
-   // std::ofstream rFile;
-
-   // if(flag_r) {
-   //     rFile.open(out_fn, std::ios::out);
-   //     out = &rFile;
-   // } else {
-   //     out = &std::cout;
-   // }
-    // double esum = std::accumulate(res.errors.begin(), res.errors.end(), 0.0);
-    // double emean = esum / res.errors.size();
-    // double vsum = std::accumulate(res.sizes.begin(), res.sizes.end(), 0.0);
-    // double vmean = vsum / res.sizes.size();
-    // double ssum = std::accumulate(res.val_errors.begin(), res.val_errors.end(), 0.0);
-    // double smean = ssum / res.val_errors.size();
-
-    //*out << "Val_errors" << std::endl;
-    //for (unsigned i = 0; i < res.val_errors.size(); ++i)  {
-    //    *out << res.val_errors[i] << " ";
-    //}
-    //*out << std::endl;
-    //*out << "Sizes" << std::endl;
-    //for (unsigned i = 0; i < res.sizes.size(); ++i)  {
-    //    *out << res.sizes[i] << " ";
[…]

### 2015-06-23 src/populationbased.cpp
working

@@ -90,3 +90,49 @@ Children PopulationBased::recombination(const IS::Solution &sa,
     }
     return make_pair(IS::Solution(ca, sol_size), IS::Solution(cb, sol_size));
 }
+
+
+void SGA::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
+    Population population = generatePopulation(sol.getSize());
+    IS::Solution best = IS::Solution(*(population.begin()));
+    best.setFitness(quality(ds, best, 0.5));
+    int iter = 0, iter_old = 0;
+    double old_q = -1.0;
+
+    while(1) {
+        iter++;
+        IS::Solution p1 = tourneySelection(population, 3);
+        IS::Solution p2 = tourneySelection(population, 3);
+        IS::Solution p1_cand = IS::Solution(p1);
+        IS::Solution p2_cand = IS::Solution(p2);
+        Children recombination(p1_cand, p2_cand);
+        mutation->tweak(p1_cand);
+        mutation->tweak(p2_cand);
+        p1_cand.setFitness(quality(ds, p1_cand, 0.5));
+        p2_cand.setFitness(quality(ds, p2_cand, 0.5));
+        if(p1.getFitness() < p1_cand.getFitness() ) {
+             
+        }
+
+        Population::const_iterator sit;
+        for (sit = population.begin(); sit != population.end(); sit++) {
+            IS::Solution p = IS::Solution(*sit);
+            local_search->optimize(ds, p);
+            p.setFitness(quality(ds, p, 0.5));
+            if (p.getFitness() > best.getFitness()) {
+                best.copy(p);
+            }
+        }
+        if (best.getFitness() - old_q < EPSILON) iter_old++;
+        else {
+            old_q = best.getFitness();
+            iter_old = 0;
+        }
+        bool stop = best.getFitness() >= max_quality or iter == iterations;
+        stop = stop or iter_old == no_change_best;
+        if (stop) break;
+        Population breeded = breed(population);
+        population = breeded;
+    }
+    sol.copy(best);
+}

### 2015-06-23 src/populationbased.h

@@ -17,7 +17,7 @@ class PopulationBased : public Metaheuristic {
   public:
     PopulationBased(int p, Metaheuristic *ls) : popsize(p), local_search(ls) { };
     set<IS::Solution> breed(const set<IS::Solution> &p) const;
-    pair<IS::Solution, IS::Solution> recombination(const IS::Solution &sa, 
+    Children recombination(const IS::Solution &sa, 
                                                    const IS::Solution &sb) 
                                                    const;
     void selection(set<IS::Solution> &sa, const set<IS::Solution> &sb);
@@ -30,7 +30,17 @@ class Hybrid : public PopulationBased {
     void optimize(const IS::Dataset &ds, IS::Solution &sol) const;
     std::string output_params() const { return ""; };
     Hybrid(int p, Metaheuristic *ls): PopulationBased(p, ls) { };
-    // set<IS::Solution> breed(set<IS::Solution> &p);
+};
+
+class SGA : public PopulationBased {
+  protected:
+    double cross_p;
+    double mutation_p;
+  public:
+    void optimize(const IS::Dataset &ds, IS::Solution &sol) const;
+    std::string output_params() const { return ""; };
+    SGA(int p, Metaheuristic *ls, double cp, double mp): 
+        PopulationBased(p, ls), cross_p(cp), mutation_p(mp) { };
 };
 
 #endif

### 2015-06-23 src/metaheuristic.cpp
working

@@ -80,11 +80,15 @@ double Metaheuristic::quality(const IS::Dataset &T,
 
 void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) const {
     std::cout << ">>>> Running Hill Climbing" << std::endl;
-    std::cout << ">>>> quality = " << max_quality << std::endl;
-    double q_max, old_q = -1;
-    int iter_old = 0, iter_total = 0;
+    std::cout << "quality = " << max_quality << std::endl;
+    std::cout << "ite_limit = " << ite_limit << std::endl;
+    std::cout << "no_change_best = " << no_change_best << std::endl;
+    std::cout << ">>>>>>>>>>>>>>>>>>>>" << std::endl;
+
+    double q_max, q_best = -1;
+    int no_change = 0, iter = 0;
     while (1) {
-        iter_total++;
+        iter++;
         IS::Solution R(S);
         tweaker->tweak(R); 
         // assert((S.getBits() ^ R.getBits()).count() == 1);
@@ -95,20 +99,16 @@ void HillClimbing::optimize(const IS::Dataset &T, IS::Solution &S) const {
             S.setBits(R.getBits());
         
         q_max = std::max(q1, q2);
-        if (old_q == q_max) {
-          iter_old ++;  
+        if (q_best > q_max) {
+            no_change++;  
         } else {
-            iter_old = 0;
+            q_best = q_max;
+            no_change = 0;
         }
         // TODO: Pass these values as parameters
-        if (q_max > max_quality or iter_old == 1000) break;
-        old_q = q_max;
-
-        if(iter_total % 100 == 0 and iter_total != 0) {
-            cout << "Total iterations: " << iter_total << endl;
-            cout << "Best fitness so far: " << q_max << endl;
-            cout << "Solution size: " << S.getBits().count() << endl << endl;
-        }
+        bool stop = q_max > max_quality or no_change == no_change_best;
+        stop = stop || iter == ite_limit;
+        if (stop) break;
     }
 }
 
@@ -167,27 +167,27 @@ std::string SimulatedAnnealing::output_params() const {
 
 void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
     std::cout << ">>>> Running ILS" << std::endl;
-    std::cout << ">>>> quality = " << max_quality << std::endl;
+    std::cout << "quality = " << max_quality << std::endl;
+    std::cout << "ite_limit = " << ite_limit << std::endl;
[…]

### 2015-06-23 src/populationbased.cpp

@@ -51,7 +51,7 @@ void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
         generatePopulation(population, sol.getSize());
         assert(population.size() == popsize);
         
-        bool stop = best_fitness > max_quality or iter == iterations;
+        bool stop = best_fitness > max_quality or iter == ite_limit;
         stop = stop or iter_old == no_change_best;
         for (int i = 0; i < 10; i++) {
             std::cout << CYAN << "KKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKKK";

### 2015-06-23 src/utils.cpp

@@ -154,28 +154,26 @@ Metaheuristic *choose_metaheuristic(std::string metaheuristic_str,
                                     int moa_counter) {
     Metaheuristic *metaheuristic;
     if (metaheuristic_str == "HillClimbing") {
-        return new HillClimbing();
-    } else if (metaheuristic_str == "SimulatedAnnealing") {
-        // Print error and exit
-        if (moa_counter != 2) error_("SimulatedAnnealing needs two arguments");
-        int t = metaheuristic_optional_arg[0]; 
-        int d = metaheuristic_optional_arg[1];
-        return new SimulatedAnnealing(t, d);
+        if (moa_counter != 2) error_("HillClimbing needs two arguments");
+        int i = metaheuristic_optional_arg[0];
+        int nc = metaheuristic_optional_arg[1];
+        return new HillClimbing(i, nc);
     } else if (metaheuristic_str == "Tabu") {
         // Print error and exit
-        if (moa_counter != 3) error_("Tabu needs three arguments");
+        if (moa_counter != 4) error_("Tabu needs four arguments");
         int l = metaheuristic_optional_arg[0]; 
         int n = metaheuristic_optional_arg[1];
         int i = metaheuristic_optional_arg[2];
-        return new Tabu(l, n, i);
+        int nc = metaheuristic_optional_arg[3];
+        return new Tabu(l, n, i, nc);
     } else if (metaheuristic_str == "ILS") {
-        
         // Print error and exit
-        if (moa_counter != 3) error_("ILS needs three arguments");
+        if (moa_counter != 4) error_("ILS needs four arguments");
         int l = metaheuristic_optional_arg[0]; 
         int t = metaheuristic_optional_arg[1];
-        int p = metaheuristic_optional_arg[2];
-        return new ILS(l, t, p);
+        int i = metaheuristic_optional_arg[2];
+        int nc = metaheuristic_optional_arg[3];
+        return new ILS(l, t, i, nc);
     }
     error_("The metaheuristic you are providing doesn't exist");
     return NULL;   // Shut up warning
@@ -329,7 +327,7 @@ void output_results(const results &res,
 }
 
 void debug_print(const IS::Dataset &dataset, const IS::Solution &solution) {
-   HillClimbing meta;
+   HillClimbing meta = HillClimbing(0,0);
    std::cout << "----------------------------------------- " << std::endl;
    meta.quality(dataset, solution, 0.5);
    std::cout << "----------------------------------------- " << std::endl;

### 2015-06-23 src/populationbased.cpp
working

@@ -3,16 +3,19 @@
 #include "utils.h"
 
 
-void PopulationBased::generatePopulation(Population &population, int size) const {
-    while (population.empty() || population.size() < popsize)
-        population.insert(IS::Solution(size));
+void PopulationBased::generatePopulation(Population &population, int size, const IS::Dataset &ds) const {
+    while (population.empty() || population.size() < popsize) {
+        IS::Solution sol = IS::Solution(size);
+        sol.setFitness(quality(ds, sol, 0.5));
+        population.insert(sol);
+    }
 }
 
 
 void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
     std::cout << ">>>> Running Hybrid algorithm" << std::endl;
     Population population;
-    generatePopulation(population, sol.getSize());
+    generatePopulation(population, sol.getSize(), ds);
     IS::Solution best = IS::Solution(*(population.begin()));
     double best_fitness = quality(ds, best, 0.5);
     best.setFitness(best_fitness);
@@ -48,7 +51,7 @@ void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
             iter_old = 0;
         }
         population = new_pop;
-        generatePopulation(population, sol.getSize());
+        generatePopulation(population, sol.getSize(), ds);
         assert(population.size() == popsize);
         
         bool stop = best_fitness > max_quality or iter == iterations;
@@ -115,8 +118,10 @@ Children PopulationBased::recombination(const IS::Solution &sa,
 
 
 void SGA::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
-    Population population = generatePopulation(sol.getSize());
-    IS::Solution best = IS::Solution(*(population.begin()));
+    Population population;
+    generatePopulation(population, sol.getSize(), ds);
+    IS::Solution best = findBest(population);
+    double best_fitness = quality(ds, best, 0.5);
     best.setFitness(quality(ds, best, 0.5));
     int iter = 0, iter_old = 0;
     double old_q = -1.0;
@@ -128,22 +133,25 @@ void SGA::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
         IS::Solution p1_cand = IS::Solution(p1);
         IS::Solution p2_cand = IS::Solution(p2);
         Children recombination(p1_cand, p2_cand);
-        mutation->tweak(p1_cand);
-        mutation->tweak(p2_cand);
[…]

### 2015-06-23 src/populationbased.h

@@ -20,8 +20,9 @@ class PopulationBased : public Metaheuristic {
                                                    const IS::Solution &sb) 
                                                    const;
     Population selection(const Population &sa, const Population &sb) const;
-    void generatePopulation(Population &p, int size) const;
+    void generatePopulation(Population &population, int size, const IS::Dataset &ds) const;
     IS::Solution tourneySelection(const set<IS::Solution> &p, int s) const;
+    IS::Solution findBest(const Population &p) const;
 };
 
 class Hybrid : public PopulationBased {
@@ -40,6 +41,7 @@ class SGA : public PopulationBased {
     std::string output_params() const { return ""; };
     SGA(int p, Metaheuristic *ls, double cp, double mp): 
         PopulationBased(p, ls), cross_p(cp), mutation_p(mp) { };
+    SGA(int p, Metaheuristic *ls) : PopulationBased(p, ls) { };
 };
 
 #endif

### 2015-06-23 src/utils.cpp

@@ -144,6 +144,9 @@ Metaheuristic *choose_population(std::string population_str,
    if (population_str == "Hybrid") {
       if (poa_counter != 1) error_("Hybrid needs Population size as argument");
       return new Hybrid(population_optional_arg[0], local_search);
+   } else if (population_str == "SGA") {
+      if (poa_counter != 1) error_("SGA needs Population size as argument");
+      return new SGA(population_optional_arg[0], local_search);
    }
    error_("The Population based algorithm you are providing doesn't exist");
    return NULL;   // Shut up warning

### 2015-06-23 src/main.cpp
working

@@ -119,6 +119,8 @@ int main(int argc, char *argv[]) {
         metaheuristic = choose_population(population_str, ls, 
                                           population_optional_arg, poa_counter);
         metaheuristic->setTweaker(tweaker);
+        metaheuristic->set_no_change_best(5);
+        metaheuristic->set_ite_limit(50);
     }
 
 

### 2015-06-23 src/metaheuristic.cpp

@@ -63,13 +63,13 @@ double Metaheuristic::quality(const IS::Dataset &T,
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
 
     // // XXX: Print
-      cout << "Hubo un total de " << count << " aciertos" << endl;
-      cout << "Result es de tamanio: " << result.size() << endl;
-      cout << "La relación es: " << clas_rate << endl;
-      cout << "El tamaño de T es: " << T.size() << endl;
-      cout << "El tamaño de training es: " << training.size() << endl;
-      cout << "El porcentaje de reducción es: " << perc_redc << endl;
-      cout << "El fitness es : " << fitness << endl << endl;
+      //cout << "Hubo un total de " << count << " aciertos" << endl;
+      //cout << "Result es de tamanio: " << result.size() << endl;
+      //cout << "La relación es: " << clas_rate << endl;
+      //cout << "El tamaño de T es: " << T.size() << endl;
+      //cout << "El tamaño de training es: " << training.size() << endl;
+      //cout << "El porcentaje de reducción es: " << perc_redc << endl;
+      //cout << "El fitness es : " << fitness << endl << endl;
 
     assert(fitness <= 1.0);
     return fitness;
@@ -173,7 +173,6 @@ void ILS::optimize(const IS::Dataset &T, IS::Solution &S) const {
     std::cout << ">>>>>>>>>>>>>>>>>>>>" << std::endl;
 
     IS::Solution best(S);
-    IS::Solution home_base(S);
 
     double q_max, q_BEST = -1.0;
     int global_iter = 0;

### 2015-06-23 src/populationbased.cpp

@@ -14,66 +14,71 @@ void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
     Population population;
     generatePopulation(population, sol.getSize());
     IS::Solution best = IS::Solution(*(population.begin()));
-    double best_fitness = quality(ds, best, 0.5);
-    best.setFitness(best_fitness);
-    int iter = 0, iter_old = 0;
-    double old_best = -1.0;
+    
+    best.setFitness(quality(ds, best, 0.5));
 
-    while(1) {
+    int iter = 0, no_change = 0;
+    while(1) {  // Main loop
         iter++;
         Population::const_iterator sit;
-        local_search->set_max_quality(std::min(best.getFitness() + 0.1, 0.95));
-        Population new_pop;
+        Population new_pop;  // New population 
+        bool best_changed = false;
         for (sit = population.begin(); sit != population.end(); sit++) {
+            local_search->set_max_quality(std::min(best.getFitness() + 0.1, 0.95));
             IS::Solution p = IS::Solution(*sit);
             local_search->optimize(ds, p);
-            double p_fitness = quality(ds, p, 0.5);
-            p.setFitness(p_fitness);
+            p.setFitness(quality(ds, p, 0.5)); 
             new_pop.insert(p);
-            if (p_fitness > best_fitness) {
+            if (p.getFitness() > best.getFitness()) {
                 best.copy(p);
-                best_fitness = p_fitness;
+                best.setFitness(p.getFitness());
+                best_changed = true;
             }
         }
-        if (best_fitness - old_best < EPSILON) {
+        if (! best_changed) {
             for (int i = 0; i < 10; i++) {
                 std::cout << BOLDGREEN << "HHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHH";
                 std::cout  << RESET << std::endl;
             }
-            iter_old++;
-        }
-        else {
-            std::cout << "BEST = " << best_fitness << " , old_best = " << old_best << endl;
-            old_best = best_fitness;
-            iter_old = 0;
+            no_change++;
+        } else {
+            no_change = 0;
         }
         population = new_pop;
         generatePopulation(population, sol.getSize());
         assert(population.size() == popsize);
         
-        bool stop = best_fitness > max_quality or iter == ite_limit;
-        stop = stop or iter_old == no_change_best;
[…]

### 2015-06-23 src/populationbased.h

@@ -15,7 +15,7 @@ class PopulationBased : public Metaheuristic {
     Metaheuristic *local_search;
   public:
     PopulationBased(int p, Metaheuristic *ls) : popsize(p), local_search(ls) { };
-    set<IS::Solution> breed(const set<IS::Solution> &p) const;
+    set<IS::Solution> breed(const IS::Dataset &ds, const set<IS::Solution> &p) const;
     pair<IS::Solution, IS::Solution> recombination(const IS::Solution &sa, 
                                                    const IS::Solution &sb) 
                                                    const;

### 2015-06-23 src/utils.cpp
fixed

@@ -139,7 +139,6 @@ Metaheuristic *choose_population(std::string population_str,
                                    int (&population_optional_arg)[10],
                                    int poa_counter) {
 
-   local_search->set_no_change_best(50);
    PopulationBased *pb; 
    if (population_str == "Hybrid") {
       if (poa_counter != 1) error_("Hybrid needs Population size as argument");

### 2015-06-23 src/main.cpp
working

@@ -118,7 +118,8 @@ int main(int argc, char *argv[]) {
         Metaheuristic *ls = metaheuristic;
         metaheuristic = choose_population(population_str, ls, 
                                           population_optional_arg, poa_counter);
-        metaheuristic->setTweaker(tweaker);
+
+        metaheuristic->setTweaker(new weightedRandomPlus(5,40));
     }
 
 

### 2015-06-23 src/metaheuristic.cpp

@@ -63,13 +63,13 @@ double Metaheuristic::quality(const IS::Dataset &T,
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
 
     // // XXX: Print
-      cout << "Hubo un total de " << count << " aciertos" << endl;
-      cout << "Result es de tamanio: " << result.size() << endl;
-      cout << "La relación es: " << clas_rate << endl;
-      cout << "El tamaño de T es: " << T.size() << endl;
-      cout << "El tamaño de training es: " << training.size() << endl;
-      cout << "El porcentaje de reducción es: " << perc_redc << endl;
-      cout << "El fitness es : " << fitness << endl << endl;
+      // cout << "Hubo un total de " << count << " aciertos" << endl;
+      // cout << "Result es de tamanio: " << result.size() << endl;
+      // cout << "La relación es: " << clas_rate << endl;
+      // cout << "El tamaño de T es: " << T.size() << endl;
+      // cout << "El tamaño de training es: " << training.size() << endl;
+      // cout << "El porcentaje de reducción es: " << perc_redc << endl;
+      // cout << "El fitness es : " << fitness << endl << endl;
 
     assert(fitness <= 1.0);
     return fitness;

### 2015-06-23 src/populationbased.cpp

@@ -8,6 +8,7 @@ void PopulationBased::generatePopulation(Population &population, int size, const
         IS::Solution sol = IS::Solution(size);
         sol.setFitness(quality(ds, sol, 0.5));
         population.insert(sol);
+        cout << BOLDBLUE << sol.getFitness() << RESET << endl;
     }
 }
 
@@ -120,6 +121,13 @@ Children PopulationBased::recombination(const IS::Solution &sa,
 void SGA::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
     Population population;
     generatePopulation(population, sol.getSize(), ds);
+
+    Population::const_iterator sit = population.begin();
+    for(; sit != population.end(); ++sit) {
+        cout << BOLDRED << sit->getFitness() << RESET << endl;
+    }
+
+
     IS::Solution best = findBest(population);
     double best_fitness = quality(ds, best, 0.5);
     best.setFitness(quality(ds, best, 0.5));
@@ -158,7 +166,7 @@ void SGA::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
             old_q = best.getFitness();
             iter_old = 0;
         }
-        bool stop = best.getFitness() >= max_quality or iter == iterations;
+        bool stop = best.getFitness() >= max_quality or iter == ite_limit;
         stop = stop or iter_old == no_change_best;
         if (stop) break;
         Population breeded = breed(population);

### 2015-06-23 src/populationbased.cpp
elitist

@@ -24,7 +24,7 @@ void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
         Population new_pop;  // New population 
         bool best_changed = false;
         for (sit = population.begin(); sit != population.end(); sit++) {
-            local_search->set_max_quality(std::min(best.getFitness() + 0.1, 0.95));
+            local_search->set_max_quality(std::min(best.getFitness() + 0.95, 0.95));
             IS::Solution p = IS::Solution(*sit);
             local_search->optimize(ds, p);
             p.setFitness(quality(ds, p, 0.5)); 
@@ -70,15 +70,15 @@ Population PopulationBased::breed(const IS::Dataset &ds, const Population &p) co
         IS::Solution c1(children.first), c2(children.second);
         tweaker->tweak(c1);
         tweaker->tweak(c2);
-        //double q_children = quality(ds, c1, 0.5) + quality(ds, c2, 0.5);
-        //double q_parents = quality(ds, p1, 0.5) + quality(ds, p2, 0.5);
-        //if (q_children > q_parents) {  // Elitist
-        breeded.insert(c1);
-        breeded.insert(c2);
-        //} else {
-        //    breeded.insert(p1);
-        //    breeded.insert(p2);
-        //}
+        double q_children = quality(ds, c1, 0.5) + quality(ds, c2, 0.5);
+        double q_parents = quality(ds, p1, 0.5) + quality(ds, p2, 0.5);
+        if (q_children > q_parents) {  // Elitist
+            breeded.insert(c1);
+            breeded.insert(c2);
+        } else {
+            breeded.insert(p1);
+            breeded.insert(p2);
+        }
     }
     return breeded;
 }

### 2015-06-23 src/main.cpp
working

@@ -120,7 +120,7 @@ int main(int argc, char *argv[]) {
                                           population_optional_arg, poa_counter);
         metaheuristic->setTweaker(tweaker);
         metaheuristic->set_no_change_best(5);
-        metaheuristic->set_ite_limit(50);
+        metaheuristic->set_ite_limit(10);
     }
 
 

### 2015-06-23 src/metaheuristic.cpp

@@ -63,13 +63,13 @@ double Metaheuristic::quality(const IS::Dataset &T,
     double fitness = alpha * clas_rate + (1 - alpha) * perc_redc;
 
     // // XXX: Print
-      cout << "Hubo un total de " << count << " aciertos" << endl;
-      cout << "Result es de tamanio: " << result.size() << endl;
-      cout << "La relación es: " << clas_rate << endl;
-      cout << "El tamaño de T es: " << T.size() << endl;
-      cout << "El tamaño de training es: " << training.size() << endl;
-      cout << "El porcentaje de reducción es: " << perc_redc << endl;
-      cout << "El fitness es : " << fitness << endl << endl;
+      //cout << "Hubo un total de " << count << " aciertos" << endl;
+      //cout << "Result es de tamanio: " << result.size() << endl;
+      //cout << "La relación es: " << clas_rate << endl;
+      //cout << "El tamaño de T es: " << T.size() << endl;
+      //cout << "El tamaño de training es: " << training.size() << endl;
+      //cout << "El porcentaje de reducción es: " << perc_redc << endl;
+      //cout << "El fitness es : " << fitness << endl << endl;
 
     assert(fitness <= 1.0);
     return fitness;

### 2015-06-23 src/populationbased.cpp

@@ -25,6 +25,7 @@ void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
         bool best_changed = false;
         for (sit = population.begin(); sit != population.end(); sit++) {
             local_search->set_max_quality(std::min(best.getFitness() + 0.95, 0.95));
+            std::cout << "Best -----> " << best.getFitness() << std::endl;
             IS::Solution p = IS::Solution(*sit);
             local_search->optimize(ds, p);
             p.setFitness(quality(ds, p, 0.5)); 

### 2015-06-23 doc/paper.tex
resultados

@@
 
+\subsection{Resultados}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     & 78.8211     & 0.09902      &  329    &  62.30\\ \hline
+    ILS      & 77.60396    & 0.11925      &  323.2  &  336.89\\ \hline
+    Memético & 78.35113    & 0.1378       &  327.04 &  160.96\\ \hline
+    SGA      & 80.1339     & 0.52334      &  334.48 &  44.45\\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Abalone}
+\label{tabla:3}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  33.3913    & 0.42512      &  23.04  & 2.6786           \\ \hline
+    ILS      &  31.5072    & 0.412238     &  21.74  & 15.253           \\ \hline
+    Memético &  33.42028   & 0.412238     &  23.06  & 8.322            \\ \hline
+    SGA      &  37.2463    & 1.07246      &  25.7   & 2.018            \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Australian}
+\label{tabla:4}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  16.2951    &  0.5155      &  10.18  & 1.3566           \\ \hline
+    ILS      &  12.8966    &  0.444467    &  8.06   & 10.0512          \\ \hline
+    Memético &  14.2838    &  0.462071    &  8.92   & 4.0198          \\ \hline
+    SGA      &  20.70254   &  2.179019    &  12.94  & 0.7512          \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Balance}
+\label{tabla:5}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  35.13766   & 0.97264      &  12.12  & 0.5884           \\ \hline
+    ILS      &  31.28007   & 0.889045     &  10.8   & 4.1052           \\ \hline
+    Memético &  33.99377   & 0.92767      &  11.72  & 1.7144           \\ \hline
+    SGA      &  39.48908   & 1.320954     &  13.62  & 0.3142           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Bupa}
+\label{tabla:6}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  30.85025   & 2.55893      &  10.38  & 0.6244          \\ \hline
+    ILS      &  26.52584   & 4.028173     &  8.92   & 7.0984          \\ \hline
+    Memético &  29.12677   & 3.643914     &  9.8    & 1.9566          \\ \hline
+    SGA      &  28.8253    & 5.3836       &  9.7    & 0.3122         \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Ecoli}
+\label{tabla:7}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  40.26272   & 4.080946     &  8.66   & 0.3336          \\ \hline
+    ILS      &  37.98743   & 4.327704     &  8.22   & 3.7046          \\ \hline
+    Memético &  38.22237   & 4.659532     &  8.24   & 1.2344          \\ \hline
+    SGA      &  38.49303   & 8.02254      &  8.3    & 0.18           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Glass}
+\label{tabla:8}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  25.86667   & 0.994915     &  7.9    & 0.429            \\ \hline
+    ILS      &  25.80216   & 0.893281     &  7.88   & 2.9763           \\ \hline
+    Memético &  26.38925   & 0.944243     &  8.06   & 1.0734           \\ \hline
+    SGA      &  26.41724   & 0.58108      &  8.08   & 0.1846           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Haberman}
+\label{tabla:9}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  16.07302   & 2.272815     &  5.64   & 1.5468           \\ \hline
+    ILS      &  13.40159   & 2.279123     &  4.7    & 15.304           \\ \hline
+    Memético &  15.16667   & 2.044946     &  5.32   & 5.7892           \\ \hline
+    SGA      &  17.27461   & 4.868776     &  6.06   & 1.0562           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Ionosphere}
+\label{tabla:8}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  5.733331   & 3.970368     &  0.86   & 0.0348           \\ \hline
+    ILS      &  5.73333    & 3.659257     &  0.86   & 0.0152           \\ \hline
+    Memético &  6.39999    & 3.99999      &  0.96   & 0.1462           \\ \hline
+    SGA      &  5.86666    & 4.992591     &  0.88   & 0.452            \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Iris}
+\label{tabla:9}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  19.3827    & 1.419281     &  8.32   & 0.8218          \\ \hline
+    ILS      &  11.77317   & 1.389261     &  5.1    & 7.9858          \\ \hline
+    Memético &  15.78001   & 1.394267     &  6.82   & 2.5364          \\ \hline
+    SGA      &  18.3673    & 4.473317     &  7.96   & 0.4278          \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Monk-2}
+\label{tabla:10}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  11.64502   & 2.53245      &  2.5    & 0.2808          \\ \hline
+    ILS      &  9.748918   & 2.191604     &  2.1    & 2.1954          \\ \hline
+    Memético &  10.49351   & 2.811653     &  2.6    & 0.7974          \\ \hline
+    SGA      &  11.70563   & 5.405375     &  2.52   & 0.1154         \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Newthyroid}
+\label{tabla:10}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  26.13381   & 0.355893     &  20.08  & 2.2424          \\ \hline
+    ILS      &  25.74077   & 0.32985      &  18.78  & 12.0942         \\ \hline
+    Memético &  26.20867   & 0.3240552    &  20.14  & 6.8738          \\ \hline
+    SGA      &  30.4033    & 0.8534       &  23.36  & 16198      \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Pima}
+\label{tabla:11}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  33.50879   & 0.683068     &  15.48  & 1.0738           \\ \hline
+    ILS      &  32.72987   & 0.586793     &  15.12  & 5.7216          \\ \hline
+    Memético &  33.91212   & 0.69262      &  15.66  & 2.9838          \\ \hline
+    SGA      &  34.41073   & 0.471327     &  15.9   & 0.6406          \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Saheart}
+\label{tabla:12}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  39.0095    & 1.891115     &  8.1    & 1.067            \\ \hline
+    ILS      &  33.8666    & 1.698715     &  7.04   & 8.053            \\ \hline
+    Memético &  32.32379   & 2.275174     &  6.72   & 4.1412          \\ \hline
+    SGA      &  35.09998   & 6.571566     &  7.3    & 0.719           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Sonar}
+\label{tabla:13}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  34.52851   & 0.4338048    &  33.08  & 3.063            \\ \hline
+    ILS      &  33         & 1.16216      &  31.62  & 16.518           \\ \hline
+    Memético &  34.09035   & 0.9440143    &  32.66  & 10.2386          \\ \hline
+    SGA      &  34.75614   & 0.3804321    &  33.3   & 2.5356           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Tic-Tac-Toe}
+\label{tabla:14}
+\end{table}
+
+\clearpage
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  49.09916   & 3.123149     & 41.54   & 41.54            \\ \hline
+    ILS      &  46.66246   & 3.070576     & 39.48   & 56.56            \\ \hline
+    Memético &  47.49272   & 3.054687     & 40.18   & 18.8208          \\ \hline
+    SGA      &  45.90784   & 4.85396      & 38.84   & 3.509            \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Vehicle}
+\label{tabla:15}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  8.681703   & 0.7069326    & 4.94    & 2.3308           \\ \hline
+    ILS      &  9.489349   & 0.601433     & 5.4     & 0.5102           \\ \hline
+    Memético &  8.365288   & 0.7850499    & 4.76    & 7.4894           \\ \hline
+    SGA      &  9.243108   & 1.241914     & 5.25    & 2.1942           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para WDBC}
+\label{tabla:15}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  28.10458   &  2.32212     &  5      & 0.3314           \\ \hline
+    ILS      &  28.9085    &  2.322049    &  5.14   & 2.5746           \\ \hline
+    Memético &  28.69935   &  2.11015     &  5.1    & 0.9834           \\ \hline
+    SGA      &  30.24836   &  3.258773    &  5.38   & 0.1628           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Wine}
+\label{tabla:15}
+\end{table}
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
+    Tabu     &  13.28302   & 3.937526     &  70.4   & 53.3222          \\ \hline
+    ILS      &  13.31699   & 1.305241     &  70.58  & 591.7742         \\ \hline
+    Memético &  14.02265   & 1.47631      &  74.32  & 172.797          \\ \hline
+    SGA      &  14.49056   & 2.111949     &  76.8   & 35.232           \\ \hline
+\end{tabular}
+}
+\caption{Resultados promedios para Wine}
+\label{tabla:15}
+\end{table}
+
 \begin{center}

### 2015-06-23 doc/paper.tex
paper

@@
 
-Cada metaheurística se comporta de manera distinta, por lo
-que surge la necesidad de crear operadores de vencidad
-que mejor se acoplen a ellas.
+Un operador de vencidad es una función que recibe una solución $S$,
+retorna una solución $S'$,  \emph{Operador} : $S$ -> $S'$, donde $S'$ es una solución \emph{un poco} distinta
+a $S$, y que tiende a ser mejor.
 
-Se implementaron 5 operadores de vecindad, que se explican brevemente
-a continuacion:
+Se implementó un operador de vencidad, el cual llamamos \emph{weightedRandomPlus}.
+Este operador tiene parametrizado una \emph{probabilidad de reducción} y
+un \emph{porcentaje de operación}.
 
-\begin{itemize}
-  \item \emph{oneRandomUnset}:
-    Este operador selecciona al
-    azar un bit ``prendido'' y lo ``apaga''. Esto se traduce
-    en ``sacar'' una instancia del conjunto de instancias solución.
-  \item \emph{upToNRandomFlips}: Este operador tiene como parametro
-    un $N$. En cada operación hace \emph{toggle} de un número aleatorio 
-    de bits que es mayor que $1$ y menor $N$, y en donde dichos bits 
-    se escogen aleatoriamente.
-  \item \emph{nRandomFlips}: Al igual que el operador anterior, 
-    requiere de un $N$. Este operador simplemente hace toggle de $N$ bits 
-    escogidos aleatoriamente.
-  \item \emph{percRandomFlips}: Este operador requiere como parametro $X$, que 
-    representa un porcentaje. La ideas es hacer \emph{toggle} de un porcentaje de la
-    cantidad de instancias del problema original. Es decir, si un problema tiene 1000
-    instancias, y el porcentaje escogido es $5 \%$, entonces se hará \emph{toggle} 
-    sobre un total de 50 bits de la solución factible
-\end{itemize}
+El \emph{porcentaje de operación} (que llamaremos $Po$ es básicamente
+indica sobre cuantos bits de la solución se operarán. Así si $Po = 10 \%$,
+y el \emph{data set} es de tamaño $1000$, entonces se operará sobre
+$100$ bits de la solución.
 
+Una vez que se sabe la cantidad de bits a operar $c = Po * 1000 = 100$,
+la \emph{probabilidad de reducción} (que llamaremos $Pr$)
+decidirá qué hacer con un proximo bit.
+Si $Pr = 60 \% = 0.6$, y un número escogido aleatoremente $0 \le n_{r} \le 1.0$
+es también $n_{r} \le Pr$, se escogerá un bit aleatoremanete y que esté
+\emph{prendido}, con el fin de apagarlo. En caso contrario que $n_{r} > Pr$,
+se escogerá aleatoreamente un bit \emph{apagdo} para prenderlo. A continuación
+se presenta el pseudo-código del operador. \\
+
+    {\fontsize{10}{10}\selectfont
+    \begin{algorithmic}
+        \Function{weightedRandomPlus}{$S: Solucion$, $sz: Entero$, 
+                                      $Pr: Entero$, $Po: Entero$}
+        \State $c \gets Po * sz$  \Comment{sz es el tamaño de dataset} 
+          \For{$i \gets 1 \textbf{ to } c$}
+                \State $rand \gets getRand()$    \Comment{Entre 0 y 1} 
+                \If{$rand < Pr$}
+                \State $\Call{apagarBitPrendido}$ 
+                \Else
+                \State $\Call{prenderBitApagado}$ 
+                \EndIf
+            \EndFor
+            \State return $S$
+        \EndFunction
+    \end{algorithmic}
+    }
+
+    ~\
+
+  Recordando que \emph{apagarBitPrendido} escoge un aleatoreamente un bit prendido,
+  y  que \emph{prenderBitApagado} escoge un aleatoreamente un bit apagado.
+
+\clearpage
 \section{Experimentos}
@@
 
+\clearpage
+
 \begin{table}[h]

### 2015-06-23 doc/paper.tex
working on paper

--- 2015-06-23 doc/paper.tex
@@
 
-\section{Experimentos}
+\section{Evaluación Experimental}
 
@@
 
-Los datos utilizados fueron
+La tarea principal de la evaluación experimental en este trabajo es evaluar el desempeño de distintas metaheurísticas sobre varios conjuntos de datos.
+Dicho desempeño se mide mediante tres indicadores:
+\begin{itemize}
+\item Porcentaje reducción 
+    Porcentaje al cual se reduce el conjunto de entrenamiento
+\item Tasa de errores de validación
+    Tasa de errores de clasificación del conjunto de validación usando el conjunto de entrenamiento reducido.
+\item tiempo de ejecución
+    Tiempo de ejecución de una corrida de la metaheurística
+\end{itemize}
+
+\subsection{Conjuntos de datos}
+Los conjuntos de datos se escogieron intentando alcanzar un balance entre tener conjuntos similares entre sí en función de número de instancias y número de atributos, y conjuntos de datos que fueran diversos entre sí, tratando de abarcar el máximo espacio posible del universo de conjuntos de datos.
+
+Los datos son obtenidos de KEEL 
+Data-Mining Software Tool [hay que poner el link en referencias].
+En la siguientes tablas se muestran los conjuntos de datos considerados pequeños, medianos y grandes respectivamente.
 
 \begin{table}[h]
-\begin{tabular}{ |l|l|l|l| }
+\begin{tabular}{ |l|l|l|l|l| }
     \hline
-    Caso    & Instancias & Atributos \\ \hline
-    Iris & 150 & 5 \\ \hline
-    Glass & 214 & 10 \\ \hline
-    Pima & 768 & 9 \\ \hline
+    Caso    & Instancias & Atributos & Clases \\ \hline
+    bupa & 345  & 6 &  2 \\ \hline
+    ecoli & 336 & 7 & 8 \\ \hline
+    glass & 214 & 9 & 10 \\ \hline
+    haberman & 306 & 3 & 2 \\ \hline
+    hayes-roth & 160 & 4 & 3 \\ \hline
+    ionosphere & 351 & 33 & 2 \\ \hline
+    iris & 150 & 4 & 5 \\ \hline
+    monk-2 & 432 & 6 & 2 \\ \hline
+    newthyroid & 215  & 5 & 3 \\ \hline
+    saheart & 462 & 9 & 2 \\ \hline
+    sonar & 208 & 60 & 2 \\ \hline
+    wine & 178 & 13 & 3 \\ \hline
 \end{tabular}
-\caption{Casos de prueba}
+\caption{Conjuntos de datos pequeños}
 \label{tabla:1}
@@
 
-\subsection{Entonación}
-
-Cada metaheuística requiere de distintos parametros para realizar
-su trabajo. Entonar cada una de ellos requirió de una serie 
-de pruebas sobre distintos problemas, parametrizando
-no solo valores necesarios para una metaheurística, sino también operadores de vencindad. 
-Por ejemplo, \emph{Simulated Annealing}, arrojaba mejores 
-resultados a medida que mayor era la \emph{temperatura} y menor el \emph{factor de decrecimiento},
-y usando como operador de vencidad \emph{percRandomFlips} con $5\%$. Sin embargo,
-aumentar la temperatura y disminuir el factor de decrecimiento implica ejecuciones 
-de mayor tiempo (mayor
-diversificación). Es por esto que para los resultados finales, se consideró $temperatura = 15.000$ 
-y $factor = 1$, obteniendo así un equilibrio entre el tiempo de ejecución y calidad de la solución.
-
-Para ILS, se probó con distintas combinaciones de las iteraciones totales (denominado tiempo en la literatura) e iteraciones locales.
-Se compara con el tanaño de la solución final.
-Se observa que respecto a esta medida, la combinación de máximo 500 iteraciones totales y máximo 50 iteraciones para la búsqueda local del problema Glass.
-Pero los valores varían para otros problemas.
-
 \begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l| }
+\begin{tabular}{ |l|l|l|l|l| }
     \hline
-    I. total / I. local & Glass & Iris & Pima \\ \hline
-    100 / 100 & (54.0351, 46.4615) & (1.53846, 51.1922) & (31.9249, 49.1153) \\ \hline
-    100 / 50  & (38.7302, 48.4974) & (7.14286, 60.7353) & (28.4018, 49.2278) \\ \hline
-    500 / 50  & (28.5714, 49.0846) & (5.33333, 48.9383) & (25.6621, 46.8585) \\ \hline
-    50 / 100  & (27.7193, 49.9487) & (5.77778, 58.6173) & (26.8657, 49.6148) \\ \hline
-    50 / 500  & (42.5397, 49.5682) & (10.303, 50.7914)  & (34.3137, 49.4) \\ \hline
-    50 / 50   & (30, 51.5986) & (8.20513, 69.4686)      & (35.1852, 49.5211) \\ \hline
+    Caso    & Instancias & Atributos & Clases \\ \hline
+    balance & 625  & 4 &  3 \\ \hline
+    pima & 768  & 8 &  2 \\ \hline
+    tic-tac-toe & 958  & 9 &  2 \\ \hline
+    vehicle & 846  & 18 &  4 \\ \hline
 \end{tabular}
-}
-\caption{Entonación de parámetros ILS, donde la una tupla representa (\% Error de validación, \% Tamaño al que se reduce)}
+\caption{Conjuntos de datos medianos}
 \label{tabla:1}
@@
 
-~\
+\begin{table}[h]
+\begin{tabular}{ |l|l|l|l|l| }
+    \hline
+    Caso    & Instancias & Atributos & Clases \\ \hline
+    banana & 5300  & 2 &  2 \\ \hline
+    abalone & 4174  & 8 &  28 \\ \hline
+\end{tabular}
+\caption{Conjuntos de datos grandes}
+\label{tabla:1}
+\end{table}
 
-Para SA, se juega con la temperatura máxima inicial como manera de modificar el comportamiento de problema.
+\subsection{Preparación de datos}
 
-~\
+Cada experimento se ejecuta particionando el conjunto de datos en dos partes: entrenamiento y validación.
+La manera de hacer esto es que cada instancia tiene 90\% de probabilidad de estar en entrenamiento y 10\% de estar en validación.
+Para garantizar que no exista un sesgo en la escogencia de esta partición, se usa la técnica de validación cruzada en 10 grupos.
+Esto consiste en dividir el conjunto de datos D en 10 partes iguales $D_1, D_2, ..., D_{10}$.
+En función de estas particiones, se definen los pares $(D \setminus D_i, D_i)$.
+Cada metaheurística es evaluada usando estos pares y sobre cada par haciendo 5 corridas distintas.
+
+\subsection{Parámetros}
+
+Cada metaheurística se entonó hasta encontrar los parámetros que mejoraban el desempeño de la misma.
 
 \begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l| }
+\scalebox{0.8}{
+\begin{tabular}{ |l|l|l|l|l| }
     \hline
-    Temp. & Glass & Iris & Pima \\ \hline
-    5000  & (35.3333, 76.1168) & (14.2857, 85.1961) & (33.2394, 50.1578) \\ \hline
-    10000 & (30.1587, 63.0743) & (7.14286, 77.2549) & (33.2394, 50.2726) \\ \hline
-    15000 & (30.7937, 55.8895) & (2.85714, 76.1275) & (31.5423, 50.3281) \\ \hline
+    Parámetro & ILS & Tabu & Hibrído & SGA  \\ \hline
+    Iter. totales & 1000 & 1000 & 50 & 1000 \\ \hline
+    Iter. locales & 100 & 100     & 10  & 100 \\ \hline
+    Iter. sin cambio & 20 & 50     & 5  & 50 \\ \hline
+    Tamaño de lista & - & 10    & -  & - \\ \hline
+    \# de Tweaks & - & 5     & -  & - \\ \hline
+    Población & - & -     & 20  & 30 \\ \hline
 \end{tabular}
 }
-\caption{Entonación de parámetros SA, donde la una tupla representa (\% Error de validación, \% Tamaño al que se reduce)}
+\caption{Parámetros usados en metaheurísticas}
 \label{tabla:2}
@@
 
-~\
-
-Con TS, se juega con valores arbitrarios para el tamaño de la lista y el número de veces que se usa el operador sobre los datos en cada iteración.
-
-~\
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l| }
-    \hline
-    Lista / Tweaks & Glass & Iris & Pima \\ \hline
-    10 / 15        & (52.0635, 72.7087) & (0, 68.8078)  & (33.2381, 45.9503) \\ \hline
-    10 / 5         & (50.7937, 75.1986) & (6.66667, 81.6296)  & (31.0502, 48.0671) \\ \hline
-    15 / 15        & (22.3333, 70.9966) & (0.888889, 73.3827) & (25.1852, 46.7816) \\ \hline
-    15 / 5         & (19.0, 73.8488)    & (0, 82.402)  & (33.5185,  47.749) \\ \hline
-    5 / 15         & (28.5714, 74.9568) & (8.3333, 70.5314)  & (35.4338, 47.012) \\ \hline
-    5 / 5          & (30.4762, 73.3679) & (7.14286, 83.4314)  & (39.3519, 48.841) \\ \hline
-\end{tabular}
-}
-\caption{Entonación de parámetros TS, donde la una tupla representa (\% Error de validación, \% Tamaño al que se reduce)}
-\label{tabla:2}
-\end{table}
+\subsection{Ejecuciones}
+Los experimentos se corrieron sobre computadoras de procesador Intel(R) Core(TM) i7-3770 CPU @ 3.40GHz, con 8 GB de RAM sobre un sistema Debian Linux.
 

### 2015-06-23 doc/paper.tex

@@
 
-Cada metaheurística se comporta de manera distinta, por lo
-que surge la necesidad de crear operadores de vencidad
-que mejor se acoplen a ellas.
+Un operador de vencidad es una función que recibe una solución $S$,
+retorna una solución $S'$,  \emph{Operador} : $S$ -> $S'$, donde $S'$ es una solución \emph{un poco} distinta
+a $S$, y que tiende a ser mejor.
 
-Se implementaron 5 operadores de vecindad, que se explican brevemente
-a continuacion:
+Se implementó un operador de vencidad, el cual llamamos \emph{weightedRandomPlus}.
+Este operador tiene parametrizado una \emph{probabilidad de reducción} y
+un \emph{porcentaje de operación}.
 
-\begin{itemize}
-  \item \emph{oneRandomUnset}:
-    Este operador selecciona al
-    azar un bit ``prendido'' y lo ``apaga''. Esto se traduce
-    en ``sacar'' una instancia del conjunto de instancias solución.
-  \item \emph{upToNRandomFlips}: Este operador tiene como parametro
-    un $N$. En cada operación hace \emph{toggle} de un número aleatorio 
-    de bits que es mayor que $1$ y menor $N$, y en donde dichos bits 
-    se escogen aleatoriamente.
-  \item \emph{nRandomFlips}: Al igual que el operador anterior, 
-    requiere de un $N$. Este operador simplemente hace toggle de $N$ bits 
-    escogidos aleatoriamente.
-  \item \emph{percRandomFlips}: Este operador requiere como parametro $X$, que 
-    representa un porcentaje. La ideas es hacer \emph{toggle} de un porcentaje de la
-    cantidad de instancias del problema original. Es decir, si un problema tiene 1000
-    instancias, y el porcentaje escogido es $5 \%$, entonces se hará \emph{toggle} 
-    sobre un total de 50 bits de la solución factible
-\end{itemize}
+El \emph{porcentaje de operación} (que llamaremos $Po$ es básicamente
+indica sobre cuantos bits de la solución se operarán. Así si $Po = 10 \%$,
+y el \emph{data set} es de tamaño $1000$, entonces se operará sobre
+$100$ bits de la solución.
+
+Una vez que se sabe la cantidad de bits a operar $c = Po * 1000 = 100$,
+la \emph{probabilidad de reducción} (que llamaremos $Pr$)
+decidirá qué hacer con un proximo bit.
+Si $Pr = 60 \% = 0.6$, y un número escogido aleatoremente $0 \le n_{r} \le 1.0$
+es también $n_{r} \le Pr$, se escogerá un bit aleatoremanete y que esté
+\emph{prendido}, con el fin de apagarlo. En caso contrario que $n_{r} > Pr$,
+se escogerá aleatoreamente un bit \emph{apagdo} para prenderlo. A continuación
+se presenta el pseudo-código del operador. \\
+
+    {\fontsize{10}{10}\selectfont
+    \begin{algorithmic}
+        \Function{weightedRandomPlus}{$S: Solucion$, $sz: Entero$, 
+                                      $Pr: Entero$, $Po: Entero$}
+        \State $c \gets Po * sz$  \Comment{sz es el tamaño de dataset} 
+          \For{$i \gets 1 \textbf{ to } c$}
+                \State $rand \gets getRand()$    \Comment{Entre 0 y 1} 
+                \If{$rand < Pr$}
+                \State $\Call{apagarBitPrendido}$ 
+                \Else
+                \State $\Call{prenderBitApagado}$ 
+                \EndIf
+            \EndFor
+            \State return $S$
+        \EndFunction
+    \end{algorithmic}
+    }
+
+    ~\
+
+  Recordando que \emph{apagarBitPrendido} escoge un aleatoreamente un bit prendido,
+  y  que \emph{prenderBitApagado} escoge un aleatoreamente un bit apagado.
+
+\clearpage
 
 \section{Evaluación Experimental}
-
-\subsection{Datos}
 

### 2015-06-23 doc/paper.tex
paper

--- 2015-06-23 doc/paper.tex
@@
 \section{Introducción}
-  En este artículo se considera el problema de Selección de Instancias, como
-  estrategia de reducción de datos para Minería de Datos. Dado un conjunto de
+  En este trabajo se considera el problema de Selección de Instancias, como
+  estrategia de reducción de datos para aplicación de procesos de KDD. Dado un conjunto de
   instancias de datos, el problema de 
@@
 
-  En el presente artículo se implementan 3 Metaheurísticas de trayectorias 
+  En el presente trabajo se implementan 3 Metaheurísticas de trayectorias 
   (Hill CLimb, ILS, Tabú) y 2 Metaheurísticas poblacionales (Algoritmo Memético y 
-  Algoritmo Genético), mostrando para cada una de ellas los resultados 
+  SGA (Stationary Genetic Algorithm por sus siglas en inglés), 
+  mostrando para cada una de ellas los resultados 
   y tiempos de corrida al aplicarlas sobre diferentes conjuntos de datos.
 
-  Este artículo se organiza en 6 secciones. Se comienza con una definición
+  Este trabajo se organiza en 6 secciones. Se comienza con una definición
   formal del problema, en donde se explica la terminología relevante a SI. 

### 2015-06-23 doc/paper.tex
working

--- 2015-06-23 doc/paper.tex
@@
 \begin{table}[h]
-\scalebox{0.8}{
+\scalebox{1}{
 \begin{tabular}{ |l|l|l|l|l| }
@@
 \subsection{Ejecuciones}
-Los experimentos se corrieron sobre computadoras de procesador Intel(R) Core(TM) i7-3770 CPU @ 3.40GHz, con 8 GB de RAM sobre un sistema Debian Linux.
-
-\subsection{Comparación}
-
-Con las ditintas corridas se obtienen resultados promedios para cada caso y tomando en cuenta cada metaheurística.
-Específicamente para el tamaño, se observa que para el caso de Glass, la búsqueda local obtiene en promedio mejores soluciones que las metaheurísticas de trayectoria.
-Pero en Pima, que es un caso más grande, las metaheurísticas de trayectoria son capacez de superar en desempeño a la búsqueda local.
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-     & LS & ILS & SA & TS  \\ \hline
-    Glass & 40.6765 & 36.3244    & 32.0952 & 33.8730 \\ \hline
-    Iris  & 6.2222 & 7.0018     & 8.0952  & 5.8966 \\ \hline
-    Pima  & 43.6754 & 30.3922 & 32.6737       & 32.9629 \\ \hline
-\end{tabular}
-}
-\caption{Resultados para error de validación promedio con parámetros entonados}
-\label{tabla:2}
-\end{table}
-
-Se realizaron 
-hasta 15 pruebas con cada una de ellas en \emph{Glass}. Los gráficos 
-presentes sugieren que \emph{ILS} es la que mejor que se comporta para este problema, 
-tanto en promedio del tamaño de la solución final (\% del conjunto de instancias), 
-como en porcentaje de errores sobre el conjunto de validación.
-
-Los tiempos reportados (por corrida) para estos experimentos son
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l| }
-    \hline
-     & ILS & SA & TS  \\ \hline
-    Glass & 51.25 s & 74.29s & 10.8s\\ \hline
-\end{tabular}
-}
-\caption{Resultados para tamaño de solución}
-\label{tabla:2}
-\end{table}
+Los experimentos se corrieron sobre computadores de procesador Intel(R) Core(TM) i7-3770 CPU @ 3.40GHz, con 8 GB de RAM sobre un sistema Debian Linux.
 

### 2015-06-23 doc/paper.tex
paper

--- 2015-06-23 doc/paper.tex
@@
     Memético &  26.20867   & 0.3240552    &  20.14  & 6.8738          \\ \hline
-    SGA      &  30.4033    & 0.8534       &  23.36  & 16198      \\ \hline
+    SGA      &  30.4033    & 0.8534       &  23.36  & 1.6198      \\ \hline
 \end{tabular}
@@
 }
-\caption{Resultados promedios para Wine}
+\caption{Resultados promedios para Banana}
 \label{tabla:15}

### 2015-06-23 src/main.cpp
doc

@@ -26,12 +26,13 @@ int poa_counter = 0;
 
 int main(int argc, char *argv[]) {
 
+    // CLI options
     static struct option long_options[] = {
         {"help", no_argument, 0, 'h'},
         {"file", required_argument, 0, 'f'},
         {"disp", no_argument, 0, 'd'},
         {"metaheuristic", required_argument, 0, 'g'},
-        {"population-base", required_argument, 0, 'p'},
+        {"population-based", required_argument, 0, 'P'},
         {"meta-optional-arg", required_argument, 0, 'm'},
         {"tweaker-optional-arg", required_argument, 0, 'o'},
         {"tweaker", required_argument, 0, 'x'},
@@ -119,6 +120,7 @@ int main(int argc, char *argv[]) {
         metaheuristic = choose_population(population_str, ls, 
                                           population_optional_arg, poa_counter);
         metaheuristic->setTweaker(tweaker);
+        // Hard coded parameters since others since these seem to work fine
         metaheuristic->set_no_change_best(30);
         metaheuristic->set_ite_limit(1000);
     }

### 2015-06-23 src/metaheuristic.cpp

@@ -1,15 +1,36 @@
+/**
+  * @file
+  * @author Gabriel Formica <gabrielformica93@gmail.com>
+  * @author Melecio Ponte <pontezambrano@gmail.com>
+  *
+  * @section Description
+  *
+  * Metaheuristics' classes
+  */
+
 #include "metaheuristic.h"
 #include "is.h"
 
 using namespace std;
 
-/* Metaheuristics methods */
-
+/**
+  * Calculate euclidean distance. Simple call IS::Instance Method
+  * @param 'a'  :  Instance
+  * @param 'b'  :  Instance
+  * @return Euclidean distance between a and b
+  */
 double Metaheuristic::calc_distance(const IS::Instance &a, 
                                     const IS::Instance &b) const {
     return a.calcDistance(b);
 }
 
+/**
+  * Find instance nearest to an instance
+  * @param 'p'  :  Dataset
+  * @param 'a'  :  Instance
+  * @return Euclidean distance to the nearest instance in p to a 
+  *         (and different to a)
+  */
 double Metaheuristic::find_nearest(const IS::Dataset &p, const IS::Instance &a) const {
     double imin = 10000000;
     double category;
@@ -27,6 +48,13 @@ double Metaheuristic::find_nearest(const IS::Dataset &p, const IS::Instance &a) 
     return category;
 }
 
+/**
+  * One-NN method. Just calculate the nearest distances of elements
+  * in result to p. Save these distances in category
+  * @param 'training'  :  Dataset
+  * @param 'result'    :  Dataset 
+  * @param 'Category'  :  vector<double> (vector of distances)
+  */
 void Metaheuristic::oneNN(const IS::Dataset &training, 
                           const IS::Dataset &result, 
                           std::vector<double> &category) const {
@@ -35,6 +63,17 @@ void Metaheuristic::oneNN(const IS::Dataset &training,
         category[i] = find_nearest(training, result[i]);
 }
 
+/**
+  * Fitness function.
+  * @param 'T'      :  Dataset
+  * @param 'S'      :  Solution 
+  * @param 'alpha'  :  double
+  * 
+  * This function calculates:
+  * alpha * (instances_well_clasified) + (1 - alpha) * (percentage_of_reduction)
+  *
+  * @return Fitness of S. 
+  */
 double Metaheuristic::quality(const IS::Dataset &T, 
                               const IS::Solution &S,
[…]

### 2015-06-23 src/metaheuristic.h

@@ -1,3 +1,14 @@
+/**
+  * @file
+  * @author Gabriel Formica <gabrielformica93@gmail.com>
+  * @author Melecio Ponte <pontezambrano@gmail.com>
+  *
+  * @section Description
+  *
+  * Metaheuristics' classes
+  */
+
+
 #ifndef METAHEURISTIC_H
 #define METAHEURISTIC_H
 
@@ -7,6 +18,7 @@
 #include "is.h"
 #include "tweaker.h"
 
+/* Abstract class */
 class Metaheuristic {
   protected:
     Tweaker *tweaker;
@@ -14,22 +26,30 @@ class Metaheuristic {
     int ite_limit = 1000;
     int no_change_best = 500;
   public:
+    /* Optimizes is basically apply the metaheuristic */
     virtual void optimize(const IS::Dataset &T, IS::Solution &R) const = 0;
-    virtual std::string output_params() const = 0;
+
+    /* Getters */
     Tweaker* getTweaker() const { return tweaker; };
+    double get_max_quality() const { return max_quality; };
+
+    /* Basic functions */
+    virtual std::string output_params() const = 0;
     double quality(const IS::Dataset &T, const IS::Solution &S, double alpha) const;
     double find_nearest(const IS::Dataset &p, const IS::Instance &a) const ;
     void oneNN(const IS::Dataset &training, 
                const IS::Dataset &result, 
                std::vector<double> &category) const;
     double calc_distance(const IS::Instance &a, const IS::Instance &b) const; 
-    double get_max_quality() const { return max_quality; };
+
+    /* Setters */
     void setTweaker(Tweaker *tw) { tweaker = tw; };
     void set_max_quality(double x) { max_quality = x; }
     void set_no_change_best(int x) { no_change_best = x; }
     void set_ite_limit(int x) {ite_limit = x; }
 };
 
+/* Hill Climbing */
 class HillClimbing : public Metaheuristic {
   public: 
     HillClimbing(int i, int nc) { 
@@ -40,33 +60,33 @@ class HillClimbing : public Metaheuristic {
     std::string output_params() const;
 };
 
-class SimulatedAnnealing : public Metaheuristic {
-  private:
-    int temperature, dec_factor;
-  public: 
-    SimulatedAnnealing(int t, int d): temperature(t), dec_factor(d) { };
-    void optimize(const IS::Dataset &T, IS::Solution &R) const;
-    std::string output_params() const;
-};
+/* Iterated Local Search */
 
[…]

### 2015-06-23 src/populationbased.cpp

@@ -1,8 +1,23 @@
+/**
+  * @file
+  * @author Gabriel Formica <gabrielformica93@gmail.com>
+  * @author Melecio Ponte <pontezambrano@gmail.com>
+  *
+  * @section Description
+  *
+  * Population-based algorithm classes
+  */
 #include "metaheuristic.h"
 #include "populationbased.h"
 #include "utils.h"
 
-
+/**
+  * Generate random population until popsize
+  * @param 'population'  : Population 
+  * @param 'ds'          : Dataset 
+  *
+  * This function fill population until population.size reach popsize
+  */
 void PopulationBased::generatePopulation(Population &population, int size, const IS::Dataset &ds) const {
     while (population.empty() || population.size() < popsize) {
         IS::Solution sol = IS::Solution(size);
@@ -11,7 +26,13 @@ void PopulationBased::generatePopulation(Population &population, int size, const
     }
 }
 
-
+/**
+  * Optimize of Hybrid
+  * @param 'T'      :  Dataset
+  * @param 'S'      :  Solution 
+  *
+  * This function save in S the best found solution 
+  */
 void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
     std::cout << ">>>> Running Hybrid algorithm" << std::endl;
     Population population;
@@ -64,6 +85,13 @@ void Hybrid::optimize(const IS::Dataset &ds, IS::Solution &sol) const {
     sol.copy(best);
 }
 
+/**
+  * Breed Method
+  * @param 'ds'      :  Dataset
+  * @param 'p'       :  Population 
+  *
+  * Apply recombination and mutation, and then selects by elitist criteria
+  */
 Population PopulationBased::breed(const IS::Dataset &ds, const Population &p) const {
     Population breeded;
     for (int i = 0; i < popsize/2; ++i) {
@@ -87,6 +115,12 @@ Population PopulationBased::breed(const IS::Dataset &ds, const Population &p) co
     return breeded;
 }
 
+/**
+  * Tourney selection (as in literature)
+  * @param 'p'       :  Population 
+  * @param 's '      :  int
+  * @return A solution choose by tourney selection
+  */
 IS::Solution PopulationBased::tourneySelection(const Population &p, int s) const {
     Population::iterator sit = p.begin();
     advance(sit, (rand() % popsize));
@@ -104,6 +138,14 @@ IS::Solution PopulationBased::tourneySelection(const Population &p, int s) const
     return best;
[…]

### 2015-06-23 src/tweaker.cpp

@@ -1,6 +1,20 @@
+/**
+  * @file
+  * @author Gabriel Formica <gabrielformica93@gmail.com>
+  * @author Melecio Ponte <pontezambrano@gmail.com>
+  *
+  * @section Description
+  *
+  * Tweaker class
+  */
+
 #include <iostream>
 #include "tweaker.h"
 
+/**
+  * Just turn off one random bit (but a bit that is on)
+  * @param 'S'  :  Solution 
+  */
 void oneRandomUnset::tweak(IS::Solution &S) {
     std::bitset<MAX> bits = S.getBits();
     int size = S.getSize();
@@ -12,6 +26,10 @@ void oneRandomUnset::tweak(IS::Solution &S) {
     S.setBits(bits);
 }
 
+/**
+  * Apply upto Random flips on random bits of solution
+  * @param 'S'  :  Solution 
+  */
 void upToNRandomFlips::tweak(IS::Solution &S) {
     std::bitset<MAX> bits = S.getBits();
     int size = S.getSize();
@@ -25,6 +43,10 @@ void upToNRandomFlips::tweak(IS::Solution &S) {
     S.setBits(bits);
 }
 
+/**
+  * Apply n Random flips on random bits of solution
+  * @param 'S'  :  Solution 
+  */
 void nRandomFlips::tweak(IS::Solution &S) {
     std::bitset<MAX> bits = S.getBits();
     int size = S.getSize();
@@ -36,6 +58,11 @@ void nRandomFlips::tweak(IS::Solution &S) {
     S.setBits(bits);
 }
 
+/**
+  * Apply flip over an amount of bits (percetange of size of dataset)
+  * and choosen randomly
+  * @param 'S'  :  Solution 
+  */
 void percRandomFlips::tweak(IS::Solution &S) {
     std::bitset<MAX> bits = S.getBits();
     int size = S.getSize();
@@ -48,6 +75,11 @@ void percRandomFlips::tweak(IS::Solution &S) {
     S.setBits(bits);
 }
 
+/**
+  * Choose an amount of bits (percentage of size of dataset) 
+  * of bits and turn them off or on with a probability
+  * @param 'S'  :  Solution 
+  */
 void weightedRandom::tweak(IS::Solution &S) {
     std::bitset<MAX> bits = S.getBits();
     int size = S.getSize();
@@ -62,6 +94,13 @@ void weightedRandom::tweak(IS::Solution &S) {
     S.setBits(bits);
 }
 
+/**
[…]

### 2015-06-23 src/utils.cpp

@@ -1,3 +1,13 @@
+/**
+  * @file
+  * @author Gabriel Formica <gabrielformica93@gmail.com>
+  * @author Melecio Ponte <pontezambrano@gmail.com>
+  *
+  * @section Description
+  *
+  * Utils classes
+  */
+
 #include <algorithm>
 #include <iostream>
 #include <fstream>
@@ -88,12 +98,29 @@ void print_help() {
     std::cout << "-o, --tweaker-optional-arg [VALUE]\tValue for tweaker"; 
     std::cout << std::endl << std::endl;
 
+    // Population
+    std::cout << "-P, --population-based [VALUE]\tSpecify population based algorithm"; 
+    std::cout << std::endl << std::endl;
+
+    std::cout << "-p [VALUE]\tValue for population based"; 
+    std::cout << std::endl << std::endl;
+
     // Example
-    std::cout << "Example" << std::endl;
-    std::cout << "   ./instance_selection -g \"Tabu\" -m 100 -m 30 ";
-    std::cout << "-x \"nRandomFlips\" -o 15" << std::endl << std::endl;
-    std::cout << "   Wich will use Tabu with a tabu list of 100 elements and";
-    std::cout << std::endl <<  "   number of tweaks equals to 30" << std::endl;
+    std::cout << "Running an simple example: " << std::endl;
+    std::cout << "   ./instance_selection -g \"Tabu\" -m 30 -m 10 -m 1000 -m 100";
+    std::cout << " -x \"weightedRandomPlus\" -o 10 -o 60 -f file_path";
+    std::cout << std::endl << std::endl;
+    std::cout << "Wich will use Tabu with a tabu list of 30 elements, ";
+    std::cout << std::endl <<  "10  number of tweaks equals to 30"; 
+    std::cout << std::endl <<  "1000 as a limit of global iterations and 100 ";
+    std::cout << std::endl <<  "as a limit of iterations without change"; 
+    std::cout << std::endl << std::endl;
+    std::cout << "Running a 10Fold validation:" << std::endl;
+    std::cout << "   ./instance_selection -g \"Tabu\" -m 30 -m 10 -m 1000 -m 100";
+    std::cout << " -x \"weightedRandomPlus\" -o 10 -o 60 -t file_DIR -n 5"; 
+    std::cout << std::endl << std::endl;
+    std::cout << "Same as above, but you have to specify the dir. -n 5 runs 5 times every validation";
+    std::cout << std::endl;
 
 }
 

### 2015-06-23 doc/paper.tex
working on paper

--- 2015-06-23 doc/paper.tex
@@
 
-    \subsection{Búsqueda local}
-    Para la búsqueda local se hace uso de un algoritmo inspirado en \emph{hill-climbing} el cual toma una solución inicial y tiene como objetivo mejorarla haciendo uso de un operador $op: S \to S$. \\
+    Para los algoritmos basados en poblaciones, se utilizó como operador de mutación el cruce en un punto (\emph{one-point crossover}), mientras que para la selección se usó la técnica de torneo (\emph{tourney selection}) con un torneo de tamaño 3.
 
-    {\fontsize{10}{10}\selectfont
+    \subsection{Algoritmo poblacional híbrido (memético)}
+    Para la busqueda poblacional híbrida se considera una combinación de la noción de población y mutación con la búsqueda local clásica.
+
+    {\fontsize{9}{12}\selectfont
     \begin{algorithmic}
-        \Function{BusquedaLocal}{$T: Problema$}
-            \State $S \gets solucionAleatoria()$
+        \Function{Híbrido}{$T: Problema$}
+            \State $Mejor \gets \Call{solucionAleatoria}$
+            \State $P \gets \Call{poblacionAleatoria}$
             \While{$\lnot condicionDeParada$}
-                \State $R \gets S$
-                \State $R \gets operador(R)$
-                \If{\Call{calidad}{R,T} $<$ \Call{calidad}{S,T}}
-                    \State $S \gets R$
+                \For{Cada $P_i$ en $P$}
+                  \State $P_i \gets \Call{busquedaLocal}{T}$
+                    \If{\Call{calidad}{$P_i$} $>$ \Call{calidad}{$Mejor$}}
+                        \State $Mejor \gets P_i$
+                    \EndIf
+                \EndFor
+            \State $P \gets \Call{reproducir}{P}$
+            \EndWhile
+            \State return $Mejor$
+        \EndFunction
+    \end{algorithmic}
+    }
+
+    ~\
+
+    Como condición de parada se establece que el mínimo de calidad de una solución sea $0.95$ y una cantidad determinada de iteraciones en las cuales la solución no cambie.
+Adicionalmente, siempre se favorece la descendencia al reproducir la población.
+
+    \subsection{SGA}
+    Para el SGA se considera utiliza un criterio de selección elitista, es decir, los hijos sólo reemplazan a los padres si son mejores. \\
+
+    {\fontsize{10}{12}\selectfont
+    \begin{algorithmic}
+        \Function{SGA}{$T: Problema$}
+            \State $P \gets \Call{poblacionAleatoria}$
+            \State $Mejor \gets $ mejor individuo en P
+            \While{$\lnot condicionDeParada$}
+                \State $P_1 \gets $ Seleccion de indv. de P  
+                \State $P_2 \gets $ Seleccion de indv. de P 
+                \State $c_1,c_2 \gets $ recombinar $P_1 y P_2$
+                \State Mutar $c_1$ y $c_2$ 
+                \If{$c_1$ y $c_2$ son mejores que $P_1$ y $P_2$}
+                    \State Reemplazar
+                \EndIf
+                \State $Mejor' \gets $ mejor individuo en P 
+                \If{$Mejor'$ $>$ $Mejor$}
+                    \State $Mejor \gets Mejor'$
                 \EndIf
             \EndWhile
-            \State return $S$
+            \State return $Mejor$
         \EndFunction
@@
 
+\clearpage
+
 \subsection{Resultados}
@@
 \scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
+\begin{tabular}{ |l|l|l|l|l|l| }
     \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     & 78.8211     & 0.09902      &  329    &  62.30\\ \hline
-    ILS      & 77.60396    & 0.11925      &  323.2  &  336.89\\ \hline
-    Memético & 78.35113    & 0.1378       &  327.04 &  160.96\\ \hline
-    SGA      & 80.1339     & 0.52334      &  334.48 &  44.45\\ \hline
+        \multicolumn{6}{ |c| }{Casos de prueba pequeños} \\
+    \hline
+    Caso & Heuristica & Err val & Tam Reduc & Error & Tiempo \\ \hline
+    \multirow{4}{*}{Abalone}   & Tabu     & 78.8211     & 0.09902 & 329    & 62.30 \\
+                               & ILS      & 77.60396    & 0.11925 & 323.2  & 336.89 \\
+                               & Memetico & 78.35113    & 0.1378  & 327.04 & 160.96 \\
+                               & SGA      & 80.1339     & 0.52334 & 334.48 & 44.45\\ \hline
+    \multirow{4}{*}{Bupa} & Tabu     & 35.13766   & 0.97264      &  12.12  & 0.5884 \\
+                             & ILS      & 31.28007   & 0.889045     &  10.8   & 4.1052 \\
+                             & Memetico & 33.99377   & 0.92767      &  11.72  & 1.7144 \\
+                             & SGA      & 39.48908   & 1.320954     &  13.62  & 0.3142\\ \hline
+    \multirow{4}{*}{Ecoli}   & Tabu     & 30.85025   & 2.55893      &  10.38  & 0.6244 \\
+                             & ILS      & 26.52584   & 4.028173     &  8.92   & 7.0984 \\
+                             & Memetico & 29.12677   & 3.643914     &  9.8    & 1.9566 \\
+                             & SGA      & 28.8253    & 5.3836       &  9.7    & 0.3122\\ \hline
+    \multirow{4}{*}{Glass}   & Tabu     & 40.26272   & 4.080946     &  8.66   & 0.3336 \\
+                             & ILS      & 37.98743   & 4.327704     &  8.22   & 3.7046 \\
+                             & Memetico & 38.22237   & 4.659532     &  8.24   & 1.2344 \\
+                             & SGA      & 38.49303   & 8.02254      &  8.3    & 0.18\\ \hline
+    \multirow{4}{*}{Haberman}   & Tabu     & 25.86667   & 0.994915     &  7.9    & 0.429 \\
+                                & ILS      & 25.80216   & 0.893281     &  7.88   & 2.9763 \\
+                                & Memetico & 26.38925   & 0.944243     &  8.06   & 1.0734 \\
+                                & SGA      & 26.41724   & 0.58108      &  8.08   & 0.1846\\ \hline
+    \multirow{4}{*}{Ionosphere}   & Tabu     & 16.07302   & 2.272815     &  5.64   & 1.5468 \\
+                                  & ILS      & 13.40159   & 2.279123     &  4.7    & 15.304 \\
+                                  & Memetico & 15.16667   & 2.044946     &  5.32   & 5.7892 \\
+                                  & SGA      & 17.27461   & 4.868776     &  6.06   & 1.0562\\ \hline
+    \multirow{4}{*}{Iris}   & Tabu     & 5.733331   & 3.970368     &  0.86   & 0.0348 \\
+                            & ILS      & 5.73333    & 3.659257     &  0.86   & 0.0152 \\
+                            & Memetico & 6.39999    & 3.99999      &  0.96   & 0.1462 \\
+                            & SGA      & 5.86666    & 4.992591     &  0.88   & 0.452\\ \hline
+    \multirow{4}{*}{Monk-2}   & Tabu     & 19.3827    & 1.419281     &  8.32   & 0.8218 \\
+                              & ILS      & 11.77317   & 1.389261     &  5.1    & 7.9858 \\
+                              & Memetico & 15.78001   & 1.394267     &  6.82   & 2.5364 \\
+                              & SGA      & 18.3673    & 4.473317     &  7.96   & 0.4278\\ \hline
+    \multirow{4}{*}{Newthyroid}   & Tabu     & 11.64502   & 2.53245      &  2.5    & 0.2808 \\
+                                  & ILS      & 9.748918   & 2.191604     &  2.1    & 2.1954 \\
+                                  & Memetico & 10.49351   & 2.811653     &  2.6    & 0.7974 \\
+                                  & SGA      & 11.70563   & 5.405375     &  2.52   & 0.1154\\ \hline
+    \multirow{4}{*}{Saheart}  & Tabu     & 33.50879   & 0.683068     &  15.48  & 1.0738 \\
+                              & ILS      & 32.72987   & 0.586793     &  15.12  & 5.7216 \\
+                              & Memetico & 33.91212   & 0.69262      &  15.66  & 2.9838 \\
+                              & SGA      & 34.41073   & 0.471327     &  15.9   & 0.6406\\ \hline
+    \multirow{4}{*}{Sonar}  & Tabu     & 39.0095    & 1.891115     &  8.1    & 1.067 \\
+                            & ILS      & 33.8666    & 1.698715     &  7.04   & 8.053 \\
+                            & Memetico & 32.32379   & 2.275174     &  6.72   & 4.1412 \\
+                            & SGA      & 35.09998   & 6.571566     &  7.3    & 0.719\\ \hline
+    \multirow{4}{*}{Wine} & Tabu     & 28.10458   &  2.32212     &  5.0      & 0.3314 \\
+                          & ILS      & 28.9085    &  2.322049    &  5.14   & 2.5746 \\
+                          & Memetico & 28.69935   &  2.11015     &  5.1    & 0.9834 \\
+                          & SGA      & 30.24836   &  3.258773    &  5.38   & 0.1628\\ \hline
+    \multirow{4}{*}{Saheart}  & Tabu     & 33.50879   & 0.683068     &  15.48  & 1.0738 \\
+                              & ILS      & 32.72987   & 0.586793     &  15.12  & 5.7216 \\
+                              & Memetico & 33.91212   & 0.69262      &  15.66  & 2.9838 \\
+                              & SGA      & 34.41073   & 0.471327     &  15.9   & 0.6406\\ \hline
 \end{tabular}
 }
-\caption{Resultados promedios para Abalone}
-\label{tabla:3}
+\end{table}
+
+\newpage
+
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l|l| }
+    \hline
+        \multicolumn{6}{ |c| }{Casos de prueba medianos} \\
+    \hline
+    Caso & Heuristica & Err val & Tam Reduc & Error & Tiempo \\ \hline
+    \multirow{4}{*}{Australian}   & Tabu     & 33.3913    & 0.42512  & 23.04 & 2.6786 \\
+                                  & ILS      & 31.5072    & 0.412238 & 21.74 & 15.253 \\
+                                  & Memetico & 33.42028   & 0.412238 & 23.06 & 8.322 \\
+                                  & SGA      & 37.2463    & 1.07246  & 25.7  & 2.018\\ \hline
+    \multirow{4}{*}{Balance} & Tabu     & 16.2951    &  0.5155      &  10.18  & 1.3566 \\
+                             & ILS      & 12.8966    &  0.444467    &  8.06   & 10.0512 \\
+                             & Memetico & 14.2838    &  0.462071    &  8.92   & 4.0198 \\
+                             & SGA      & 20.70254   &  2.179019    &  12.94  & 0.7512\\ \hline
+    \multirow{4}{*}{Pima}   & Tabu     & 26.13381   & 0.355893     &  20.08  & 2.2424 \\
+                            & ILS      & 25.74077   & 0.32985      &  18.78  & 12.0942 \\
+                            & Memetico & 26.20867   & 0.3240552    &  20.14  & 6.8738 \\
+                            & SGA      & 30.4033    & 0.8534       &  23.36  & 16198.0\\ \hline
+    \multirow{4}{*}{Tic-Tac-Toe}   & Tabu     & 34.52851   & 0.4338048    &  33.08  & 3.063 \\
+                                   & ILS      & 33.0         & 1.16216      &  31.62  & 16.518 \\
+                                   & Memetico & 34.09035   & 0.9440143    &  32.66  & 10.2386 \\
+                                   & SGA      & 34.75614   & 0.3804321    &  33.3   & 2.5356\\ \hline
+    \multirow{4}{*}{Vehicle}  & Tabu     & 49.09916   & 3.123149     & 41.54   & 41.54 \\
+                              & ILS      & 46.66246   & 3.070576     & 39.48   & 56.56 \\
+                              & Memetico & 47.49272   & 3.054687     & 40.18   & 18.8208 \\
+                              & SGA      & 45.90784   & 4.85396      & 38.84   & 3.509\\ \hline
+    \multirow{4}{*}{WDBC} & Tabu     & 8.681703   & 0.7069326    & 4.94    & 2.3308 \\
+                          & ILS      & 9.489349   & 0.601433     & 5.4     & 0.5102 \\
+                          & Memetico & 8.365288   & 0.7850499    & 4.76    & 7.4894 \\
+                          & SGA      & 9.243108   & 1.241914     & 5.25    & 2.1942\\ \hline
+\end{tabular}
+}
 \end{table}
@@
 \scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
+\begin{tabular}{ |l|l|l|l|l|l| }
     \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  33.3913    & 0.42512      &  23.04  & 2.6786           \\ \hline
-    ILS      &  31.5072    & 0.412238     &  21.74  & 15.253           \\ \hline
-    Memético &  33.42028   & 0.412238     &  23.06  & 8.322            \\ \hline
-    SGA      &  37.2463    & 1.07246      &  25.7   & 2.018            \\ \hline
+        \multicolumn{6}{ |c| }{Casos de prueba grandes} \\
+    \hline
+    Caso & Heuristica & Err val & Tam Reduc & Error & Tiempo \\ \hline
+    \multirow{4}{*}{Banana} & Tabu     & 13.28302   & 3.937526     &  70.4   & 53.3222 \\
+                            & ILS      & 13.31699   & 1.305241     &  70.58  & 591.7742 \\
+                            & Memetico & 14.02265   & 1.47631      &  74.32  & 172.797 \\
+                            & SGA      & 14.49056   & 2.111949     &  76.8   & 35.232\\ \hline
 \end{tabular}
 }
-\caption{Resultados promedios para Australian}
-\label{tabla:4}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  16.2951    &  0.5155      &  10.18  & 1.3566           \\ \hline
-    ILS      &  12.8966    &  0.444467    &  8.06   & 10.0512          \\ \hline
-    Memético &  14.2838    &  0.462071    &  8.92   & 4.0198          \\ \hline
-    SGA      &  20.70254   &  2.179019    &  12.94  & 0.7512          \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Balance}
-\label{tabla:5}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  35.13766   & 0.97264      &  12.12  & 0.5884           \\ \hline
-    ILS      &  31.28007   & 0.889045     &  10.8   & 4.1052           \\ \hline
-    Memético &  33.99377   & 0.92767      &  11.72  & 1.7144           \\ \hline
-    SGA      &  39.48908   & 1.320954     &  13.62  & 0.3142           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Bupa}
-\label{tabla:6}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  30.85025   & 2.55893      &  10.38  & 0.6244          \\ \hline
-    ILS      &  26.52584   & 4.028173     &  8.92   & 7.0984          \\ \hline
-    Memético &  29.12677   & 3.643914     &  9.8    & 1.9566          \\ \hline
-    SGA      &  28.8253    & 5.3836       &  9.7    & 0.3122         \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Ecoli}
-\label{tabla:7}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  40.26272   & 4.080946     &  8.66   & 0.3336          \\ \hline
-    ILS      &  37.98743   & 4.327704     &  8.22   & 3.7046          \\ \hline
-    Memético &  38.22237   & 4.659532     &  8.24   & 1.2344          \\ \hline
-    SGA      &  38.49303   & 8.02254      &  8.3    & 0.18           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Glass}
-\label{tabla:8}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  25.86667   & 0.994915     &  7.9    & 0.429            \\ \hline
-    ILS      &  25.80216   & 0.893281     &  7.88   & 2.9763           \\ \hline
-    Memético &  26.38925   & 0.944243     &  8.06   & 1.0734           \\ \hline
-    SGA      &  26.41724   & 0.58108      &  8.08   & 0.1846           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Haberman}
-\label{tabla:9}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  16.07302   & 2.272815     &  5.64   & 1.5468           \\ \hline
-    ILS      &  13.40159   & 2.279123     &  4.7    & 15.304           \\ \hline
-    Memético &  15.16667   & 2.044946     &  5.32   & 5.7892           \\ \hline
-    SGA      &  17.27461   & 4.868776     &  6.06   & 1.0562           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Ionosphere}
-\label{tabla:8}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  5.733331   & 3.970368     &  0.86   & 0.0348           \\ \hline
-    ILS      &  5.73333    & 3.659257     &  0.86   & 0.0152           \\ \hline
-    Memético &  6.39999    & 3.99999      &  0.96   & 0.1462           \\ \hline
-    SGA      &  5.86666    & 4.992591     &  0.88   & 0.452            \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Iris}
-\label{tabla:9}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  19.3827    & 1.419281     &  8.32   & 0.8218          \\ \hline
-    ILS      &  11.77317   & 1.389261     &  5.1    & 7.9858          \\ \hline
-    Memético &  15.78001   & 1.394267     &  6.82   & 2.5364          \\ \hline
-    SGA      &  18.3673    & 4.473317     &  7.96   & 0.4278          \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Monk-2}
-\label{tabla:10}
-\end{table}
-
-\clearpage
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  11.64502   & 2.53245      &  2.5    & 0.2808          \\ \hline
-    ILS      &  9.748918   & 2.191604     &  2.1    & 2.1954          \\ \hline
-    Memético &  10.49351   & 2.811653     &  2.6    & 0.7974          \\ \hline
-    SGA      &  11.70563   & 5.405375     &  2.52   & 0.1154         \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Newthyroid}
-\label{tabla:10}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  26.13381   & 0.355893     &  20.08  & 2.2424          \\ \hline
-    ILS      &  25.74077   & 0.32985      &  18.78  & 12.0942         \\ \hline
-    Memético &  26.20867   & 0.3240552    &  20.14  & 6.8738          \\ \hline
-    SGA      &  30.4033    & 0.8534       &  23.36  & 16198      \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Pima}
-\label{tabla:11}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  33.50879   & 0.683068     &  15.48  & 1.0738           \\ \hline
-    ILS      &  32.72987   & 0.586793     &  15.12  & 5.7216          \\ \hline
-    Memético &  33.91212   & 0.69262      &  15.66  & 2.9838          \\ \hline
-    SGA      &  34.41073   & 0.471327     &  15.9   & 0.6406          \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Saheart}
-\label{tabla:12}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  39.0095    & 1.891115     &  8.1    & 1.067            \\ \hline
-    ILS      &  33.8666    & 1.698715     &  7.04   & 8.053            \\ \hline
-    Memético &  32.32379   & 2.275174     &  6.72   & 4.1412          \\ \hline
-    SGA      &  35.09998   & 6.571566     &  7.3    & 0.719           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Sonar}
-\label{tabla:13}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  34.52851   & 0.4338048    &  33.08  & 3.063            \\ \hline
-    ILS      &  33         & 1.16216      &  31.62  & 16.518           \\ \hline
-    Memético &  34.09035   & 0.9440143    &  32.66  & 10.2386          \\ \hline
-    SGA      &  34.75614   & 0.3804321    &  33.3   & 2.5356           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Tic-Tac-Toe}
-\label{tabla:14}
-\end{table}
-
-\clearpage
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  49.09916   & 3.123149     & 41.54   & 41.54            \\ \hline
-    ILS      &  46.66246   & 3.070576     & 39.48   & 56.56            \\ \hline
-    Memético &  47.49272   & 3.054687     & 40.18   & 18.8208          \\ \hline
-    SGA      &  45.90784   & 4.85396      & 38.84   & 3.509            \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Vehicle}
-\label{tabla:15}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  8.681703   & 0.7069326    & 4.94    & 2.3308           \\ \hline
-    ILS      &  9.489349   & 0.601433     & 5.4     & 0.5102           \\ \hline
-    Memético &  8.365288   & 0.7850499    & 4.76    & 7.4894           \\ \hline
-    SGA      &  9.243108   & 1.241914     & 5.25    & 2.1942           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para WDBC}
-\label{tabla:15}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  28.10458   &  2.32212     &  5      & 0.3314           \\ \hline
-    ILS      &  28.9085    &  2.322049    &  5.14   & 2.5746           \\ \hline
-    Memético &  28.69935   &  2.11015     &  5.1    & 0.9834           \\ \hline
-    SGA      &  30.24836   &  3.258773    &  5.38   & 0.1628           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Wine}
-\label{tabla:15}
-\end{table}
-
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l| }
-    \hline
-             &  Err Val    & Tamaño Reduc &  Err    & Tiempo (segundos)\\ \hline
-    Tabu     &  13.28302   & 3.937526     &  70.4   & 53.3222          \\ \hline
-    ILS      &  13.31699   & 1.305241     &  70.58  & 591.7742         \\ \hline
-    Memético &  14.02265   & 1.47631      &  74.32  & 172.797          \\ \hline
-    SGA      &  14.49056   & 2.111949     &  76.8   & 35.232           \\ \hline
-\end{tabular}
-}
-\caption{Resultados promedios para Wine}
-\label{tabla:15}
 \end{table}

### 2015-06-23 doc/paper.tex
more paper what ladilla

@@
 \maketitle
+
+\begin{abstract}
+  El problema de \textit{Selección de Instancias}  busca
+  escoger un subconjunto de instancias de menor cardinalidad del conjunto original
+  que mantenga la clasificación de las muestras a ser usadas como un conjunto
+  de entrenamiento.
+En este articulo se considera la resolución de este problema mediante varios algoritmos metaheurísticos de trayectoria y poblacionales.
+Particularmente se consideran las técnicas Tabu, ILS, Memético y SGA.
+Se comparan estas implementaciones con un conjunto extenso de datos y se extraen conclusiones acerca de su desempeño en tiempo y calidad de solución.
+\end{abstract}
+
 \section{Introducción}
@@
 
-  En el presente trabajo se implementan 3 Metaheurísticas de trayectorias 
-  (Hill CLimb, ILS, Tabú) y 2 Metaheurísticas poblacionales (Algoritmo Memético y 
+  En el presente trabajo se implementan 2 Metaheurísticas de trayectorias 
+  (ILS, Tabú) y 2 Metaheurísticas poblacionales (Algoritmo Memético y 
   SGA (Stationary Genetic Algorithm por sus siglas en inglés), 
@@
   concluye en base a dicho resultados.
-
-\begin{abstract}
-  El problema de \textit{Selección de Instancias}  busca
-  escoger un subconjunto de instancias de menor cardinalidad del conjunto original
-  que mantenga o mejore la clasificación de las muestras a ser usadas como un conjunto
-  de entrenamiento. En este articulo se considera la resolución de este problema
-  mediante varios algoritmos metaheurísticos, finalizando con los resultados 
-  experimentales y comparaciones entre ellos.
-\end{abstract}
 
@@
     Caso & Heuristica & Err val & Tam Reduc & Error & Tiempo \\ \hline
-    \multirow{4}{*}{Abalone}   & Tabu     & 78.8211     & 0.09902 & 329    & 62.30 \\
-                               & ILS      & 77.60396    & 0.11925 & 323.2  & 336.89 \\
-                               & Memetico & 78.35113    & 0.1378  & 327.04 & 160.96 \\
-                               & SGA      & 80.1339     & 0.52334 & 334.48 & 44.45\\ \hline
     \multirow{4}{*}{Bupa} & Tabu     & 35.13766         & 0.97264      &  12.12          & 0.5884 \\
-                             & ILS      & 31.28007   & 0.889045     &  10.8   & 4.1052 \\
+                             & ILS      & \bf{31.28007} & \bf{0.889045}     &  \bf{10.8} & 4.1052 \\
                              & Memetico & 33.99377      & 0.92767      &  11.72          & 1.7144 \\
-                             & SGA      & 39.48908   & 1.320954     &  13.62  & 0.3142\\ \hline
-    \multirow{4}{*}{Ecoli}   & Tabu     & 30.85025   & 2.55893      &  10.38  & 0.6244 \\
-                             & ILS      & 26.52584   & 4.028173     &  8.92   & 7.0984 \\
+                             & SGA      & 39.48908      & 1.320954     &  13.62          & \bf{0.3142}\\ \hline
+    \multirow{4}{*}{Ecoli}   & Tabu     & 30.85025      & \bf{2.55893}      &  10.38  & 0.6244 \\
+                             & ILS      & \bf{26.52584} & 4.028173     &  \bf{8.92}   & 7.0984 \\
                              & Memetico & 29.12677      & 3.643914     &  9.8    & 1.9566 \\
-                             & SGA      & 28.8253    & 5.3836       &  9.7    & 0.3122\\ \hline
-    \multirow{4}{*}{Glass}   & Tabu     & 40.26272   & 4.080946     &  8.66   & 0.3336 \\
-                             & ILS      & 37.98743   & 4.327704     &  8.22   & 3.7046 \\
+                             & SGA      & 28.8253       & 5.3836       &  9.7    & \bf{0.3122}\\ \hline
+    \multirow{4}{*}{Glass}   & Tabu     & 40.26272   & \bf{4.080946}     &  8.66   & 0.3336 \\
+                             & ILS      & \bf{37.98743}   & 4.327704     &  \bf{8.22}   & 3.7046 \\
                              & Memetico & 38.22237   & 4.659532     &  8.24   & 1.2344 \\
-                             & SGA      & 38.49303   & 8.02254      &  8.3    & 0.18\\ \hline
+                             & SGA      & 38.49303   & 8.02254      &  8.3    & \bf{0.18}\\ \hline
     \multirow{4}{*}{Haberman}   & Tabu     & 25.86667   & 0.994915        &  7.9    & 0.429 \\
-                                & ILS      & 25.80216   & 0.893281     &  7.88   & 2.9763 \\
+                                & ILS      & \bf{25.80216}   & 0.893281   &  \bf{7.88}   & 2.9763 \\
                                 & Memetico & 26.38925   & 0.944243        &  8.06   & 1.0734 \\
-                                & SGA      & 26.41724   & 0.58108      &  8.08   & 0.1846\\ \hline
+                                & SGA      & 26.41724   & \bf{0.58108}    &  8.08   & \bf{0.1846}\\ \hline
     \multirow{4}{*}{Ionosphere}   & Tabu     & 16.07302   & 2.272815     &  5.64   & 1.5468 \\
-                                  & ILS      & 13.40159   & 2.279123     &  4.7    & 15.304 \\
-                                  & Memetico & 15.16667   & 2.044946     &  5.32   & 5.7892 \\
-                                  & SGA      & 17.27461   & 4.868776     &  6.06   & 1.0562\\ \hline
-    \multirow{4}{*}{Iris}   & Tabu     & 5.733331   & 3.970368     &  0.86   & 0.0348 \\
-                            & ILS      & 5.73333    & 3.659257     &  0.86   & 0.0152 \\
+                                  & ILS      & \bf{13.40159}   & 2.279123     &  \bf{4.7}    & 15.304 \\
+                                  & Memetico & 15.16667   & \bf{2.044946}     &  5.32   & 5.7892 \\
+                                  & SGA      & 17.27461   & 4.868776     &  6.06   & \bf{1.0562}\\ \hline
+    \multirow{4}{*}{Iris}   & Tabu     & \bf{5.733331}   & 3.970368     &  \bf{0.86}   & 0.0348 \\
+                            & ILS      & \bf{5.73333}    & \bf{3.659257}     &  \bf{0.86}   & \bf{0.0152} \\
                             & Memetico & 6.39999         & 3.99999      &  0.96   & 0.1462 \\
@@
     \multirow{4}{*}{Monk-2}   & Tabu     & 19.3827    & 1.419281     &  8.32   & 0.8218 \\
-                              & ILS      & 11.77317   & 1.389261     &  5.1    & 7.9858 \\
+                              & ILS      & \bf{11.77317}   & \bf{1.389261}     &  \bf{5.1}    & 7.9858 \\
                               & Memetico & 15.78001   & 1.394267     &  6.82   & 2.5364 \\
-                              & SGA      & 18.3673    & 4.473317     &  7.96   & 0.4278\\ \hline
+                              & SGA      & 18.3673    & 4.473317     &  7.96   & \bf{0.4278}\\ \hline
     \multirow{4}{*}{Newthyroid}   & Tabu     & 11.64502   & 2.53245      &  2.5    & 0.2808 \\
-                                  & ILS      & 9.748918   & 2.191604     &  2.1    & 2.1954 \\
+                                  & ILS      & \bf{9.748918}   & \bf{2.191604}     &  \bf{2.1}    & 2.1954 \\
                                   & Memetico & 10.49351   & 2.811653     &  2.6    & 0.7974 \\
-                                  & SGA      & 11.70563   & 5.405375     &  2.52   & 0.1154\\ \hline
+                                  & SGA      & 11.70563   & 5.405375     &  2.52   & \bf{0.1154}\\ \hline
     \multirow{4}{*}{Saheart}  & Tabu     & 33.50879   & 0.683068     &  15.48  & 1.0738 \\
-                              & ILS      & 32.72987   & 0.586793     &  15.12  & 5.7216 \\
+                              & ILS      & \bf{32.72987}   & 0.586793     &  \bf{15.12}  & 5.7216 \\
                               & Memetico & 33.91212   & 0.69262      &  15.66  & 2.9838 \\
-                              & SGA      & 34.41073   & 0.471327     &  15.9   & 0.6406\\ \hline
+                              & SGA      & 34.41073   & \bf{0.471327}     &  15.9   & \bf{0.6406}\\ \hline
     \multirow{4}{*}{Sonar}  & Tabu     & 39.0095    & 1.891115     &  8.1    & 1.067 \\
-                            & ILS      & 33.8666    & 1.698715     &  7.04   & 8.053 \\
-                            & Memetico & 32.32379   & 2.275174     &  6.72   & 4.1412 \\
-                            & SGA      & 35.09998   & 6.571566     &  7.3    & 0.719\\ \hline
-    \multirow{4}{*}{Wine} & Tabu     & 28.10458   &  2.32212     &  5.0      & 0.3314 \\
+                            & ILS      & 33.8666    & \bf{1.698715}     &  7.04   & 8.053 \\
+                            & Memetico & \bf{32.32379}   & 2.275174     &  \bf{6.72}   & 4.1412 \\
+                            & SGA      & 35.09998   & 6.571566     &  7.3    & \bf{0.719}\\ \hline
+    \multirow{4}{*}{Wine} & Tabu     & \bf{28.10458}   &  2.32212     &  \bf{5.0}      & 0.3314 \\
                           & ILS      & 28.9085    &  2.322049    &  5.14   & 2.5746 \\
-                          & Memetico & 28.69935   &  2.11015     &  5.1    & 0.9834 \\
-                          & SGA      & 30.24836   &  3.258773    &  5.38   & 0.1628\\ \hline
+                          & Memetico & 28.69935   &  \bf{2.11015}     &  5.1    & 0.9834 \\
+                          & SGA      & 30.24836   &  3.258773    &  5.38   & \bf{0.1628}\\ \hline
     \multirow{4}{*}{Saheart}  & Tabu     & 33.50879   & 0.683068     &  15.48  & 1.0738 \\
-                              & ILS      & 32.72987   & 0.586793     &  15.12  & 5.7216 \\
+                              & ILS      & \bf{32.72987}   & 0.586793     &  \bf{15.12}  & 5.7216 \\
                               & Memetico & 33.91212   & 0.69262      &  15.66  & 2.9838 \\
-                              & SGA      & 34.41073   & 0.471327     &  15.9   & 0.6406\\ \hline
+                              & SGA      & 34.41073   & \bf{0.471327}     &  15.9   & \bf{0.6406}\\ \hline
 \end{tabular}
@@
     \multirow{4}{*}{Australian}   & Tabu     & 33.3913     & 0.42512  & 23.04 & 2.6786 \\
-                                  & ILS      & 31.5072    & 0.412238 & 21.74 & 15.253 \\
-                                  & Memetico & 33.42028   & 0.412238 & 23.06 & 8.322 \\
-                                  & SGA      & 37.2463    & 1.07246  & 25.7  & 2.018\\ \hline
+                                  & ILS      & \bf{31.5072} & \bf{0.412238} & \bf{21.74} & 15.253 \\
+                                  & Memetico & 33.42028   & \bf{0.412238} & 23.06 & 8.322 \\
+                                  & SGA      & 37.2463    & 1.07246  & 25.7  & \bf{2.018}\\ \hline
     \multirow{4}{*}{Balance} & Tabu     & 16.2951    &  0.5155      &  10.18  & 1.3566 \\
-                             & ILS      & 12.8966    &  0.444467    &  8.06   & 10.0512 \\
+                             & ILS      & \bf{12.8966}    &  \bf{0.444467}    &  \bf{8.06}   & 10.0512 \\
                              & Memetico & 14.2838    &  0.462071    &  8.92   & 4.0198 \\
-                             & SGA      & 20.70254   &  2.179019    &  12.94  & 0.7512\\ \hline
-    \multirow{4}{*}{Pima}   & Tabu     & 26.13381   & 0.355893     &  20.08  & 2.2424 \\
-                            & ILS      & 25.74077   & 0.32985      &  18.78  & 12.0942 \\
-                            & Memetico & 26.20867   & 0.3240552    &  20.14  & 6.8738 \\
+                             & SGA      & 20.70254   &  2.179019    &  12.94  & \bf{0.7512}\\ \hline
+    \multirow{4}{*}{Pima}   & Tabu     & 26.13381   & 0.355893     &  20.08  & \bf{2.2424} \\
+                            & ILS      & \bf{25.74077}   & 0.32985      &  \bf{18.78}  & 12.0942 \\
+                            & Memetico & 26.20867   & \bf{0.3240552}    &  20.14  & 6.8738 \\
                             & SGA      & 30.4033    & 0.8534       &  23.36  & 16198.0\\ \hline
     \multirow{4}{*}{Tic-Tac-Toe}   & Tabu     & 34.52851   & 0.4338048    &  33.08  & 3.063 \\
-                                   & ILS      & 33.0         & 1.16216      &  31.62  & 16.518 \\
+                                   & ILS      & \bf{33.0}         & 1.16216      &  \bf{31.62}  & 16.518 \\
                                    & Memetico & 34.09035   & 0.9440143    &  32.66  & 10.2386 \\
-                                   & SGA      & 34.75614   & 0.3804321    &  33.3   & 2.5356\\ \hline
+                                   & SGA      & 34.75614   & \bf{0.3804321}    &  33.3   & \bf{2.5356}\\ \hline
     \multirow{4}{*}{Vehicle}  & Tabu     & 49.09916   & 3.123149     & 41.54   & 41.54 \\
-                              & ILS      & 46.66246   & 3.070576     & 39.48   & 56.56 \\
-                              & Memetico & 47.49272   & 3.054687     & 40.18   & 18.8208 \\
-                              & SGA      & 45.90784   & 4.85396      & 38.84   & 3.509\\ \hline
+                              & ILS      & \bf{46.66246}   & 3.070576     & \bf{39.48}   & 56.56 \\
+                              & Memetico & 47.49272   & \bf{3.054687}     & 40.18   & 18.8208 \\
+                              & SGA      & 45.90784   & 4.85396      & 38.84   & \bf{3.509}\\ \hline
     \multirow{4}{*}{WDBC} & Tabu     & 8.681703   & 0.7069326    & 4.94    & 2.3308 \\
-                          & ILS      & 9.489349   & 0.601433     & 5.4     & 0.5102 \\
-                          & Memetico & 8.365288   & 0.7850499    & 4.76    & 7.4894 \\
+                          & ILS      & 9.489349   & \bf{0.601433}     & 5.4     & \bf{0.5102} \\
+                          & Memetico & \bf{8.365288}   & 0.7850499    & \bf{4.76}    & 7.4894 \\
                           & SGA      & 9.243108   & 1.241914     & 5.25    & 2.1942\\ \hline
@@
     Caso & Heuristica & Err val & Tam Reduc & Error & Tiempo \\ \hline
-    \multirow{4}{*}{Banana} & Tabu     & 13.28302   & 3.937526     &  70.4   & 53.3222 \\
-                            & ILS      & 13.31699   & 1.305241     &  70.58  & 591.7742 \\
+    \multirow{4}{*}{Banana} & Tabu     & \bf{13.28302}   & 3.937526     &  \bf{70.4}   & 53.3222 \\
+                            & ILS      & 13.31699   & \bf{1.305241}     &  70.58  & 591.7742 \\
                             & Memetico & 14.02265   & 1.47631      &  74.32  & 172.797 \\
-                            & SGA      & 14.49056   & 2.111949     &  76.8   & 35.232\\ \hline
+                            & SGA      & 14.49056   & 2.111949     &  76.8   & \bf{35.232}\\ \hline
 \end{tabular}
@@
 \end{center}
+
+\section{Conclusiones y recomendaciones}
+
+Las metaheurísticas lograron resolver en gran medida el problema propuesto.
+Considerando los porcentajes promedio de reducción y de errores de validación se observan resultados prometedores para los casos evaluados.
+El mayor espacio a mejora se encuentra en los errores de validación, para lo cual se debe considerar evaluar entonaciones adicionales sobre los operadores usados por las metaheurísticas.
+
+Se hace evidente además que las metaheurísticas funcionan de una manera por casos, por lo cual los parámetros para un conjunto de datos no necesariamente son buenos para otro.
+Por lo cual se requiere un estudio más avanzado de cada uno de los conjuntos de datos que se quieran tratar.
+
+En general, si se está buscando rapidez, las metaheurísticas de Tabu y SGA tienen un tiempo menor de corrida, sin embargo, para este trabajo se necesita mayor cuidado en la entonación de ésta última ya que fue común que sus resultados no fueran tan buenos como los de otros.
+Análogamente, la técnica con mejor desempeño en detrimento del tiempo fue el algoritmo memético.
 

### 2015-06-24 doc/paper.tex
paper

--- 2015-06-23 doc/paper.tex
@@
     Para los algoritmos basados en poblaciones, se utilizó como operador de mutación el cruce en un punto (\emph{one-point crossover}), mientras que para la selección se usó la técnica de torneo (\emph{tourney selection}) con un torneo de tamaño 3.
+
+    \subsection{Tabú}
+    Tabú mantiene una \emph{lista tabú} en la cual mantiene un historial de soluciones candidatas recientemente consideras 
+    y rechaza retornar dichos candidatos hasta que estén lo suficientemente en el pasado. De esta manera,
+
+    {\fontsize{9}{12}\selectfont
+    \begin{algorithmic}
+        \Function{Tabú}{$T: Problema, S: Solution, l: Entero, N: Entero$}
+            \State $S$ \Comment{Solución candidata}
+            \State $N$ \Comment{Numero de tweaks a aplicar}
+            \State $l$ \Comment{Largo máximo de la lista tabu}
+            \State $Mejor \gets S$ 
+            \State $Tabu \gets {}$  \Comment{Lista tabu inicialmente vacía}
+            \State $Tabu \gets S \cup Tabu$ 
+            \While{$\lnot condicionDeParada$}
+                \If{\Call{length}{$Tabu$} $>$ $l$}
+                    \State \Call{removeOldestElement}{$Tabu$}
+                \EndIf
+                \State $R \gets \Call{tweak}{S}$
+                \For{$N - 1 \ veces$}
+                \State $S' \gets \Call{copy}{S}$
+                \State $W \gets \Call{tweak}{S'}$
+                \If{W no peternece a Tabu Y ($\Call{quality}{W} > \Call{quality}{R}$
+                    O R pertenece a Tabu)}
+                    \State $R \gets W$
+                \EndIf
+            \EndFor
+            \If{R no peternece a Tabu Y ($\Call{quality}{R} > \Call{quality}{S}$}
+                \State $S \gets R$
+                \State $Tabu \gets R \cup Tabu$ 
+            \EndIf
+            \If{$\Call{quality}{S} > \Call{quality}{Mejor}$}
+                \State $Mejor \gets S$
+            \EndIf
+            \EndWhile
+            \State return $Mejor$
+        \EndFunction
+    \end{algorithmic}
+    }
+
+    ~\
+
+    Como condición de parada se establece que el mínimo de calidad de una solución sea $0.95$ y una cantidad determinada de iteraciones en las cuales la solución no cambie.
+
+    \subsection{ILS}
+    ILS es basicamente una versión mejorada de Hill-Climbing con reinicios aleatorios. A continuación, se presenta al lector el pseudo-código:
+
+    {\fontsize{9}{12}\selectfont
+    \begin{algorithmic}
+        \Function{ILS}{$T: Problema, S: Solution,  t: Entero$}
+            \State $t$ \Comment{Tiempo para reinicio}
+            \State $S$ \Comment{Solución inicial}
+            \State $Mejor \gets S$
+            \While{$\lnot condicionDeParada$}
+                \State $time \gets numero aleatorio en base a t$
+
+                \While{$\lnot condicionDeParadaInterna$}
+                    \State $S' \gets \Call{copy}{S}$
+                    \State $R \gets \Call{tweak}{S'}$
+                    \If{$\Call{quality}{R} > \Call{quality}{S}$}
+                        \State $S \gets R$
+                    \EndIf
+                \EndWhile
+            \If{$\Call{quality}{S} > \Call{quality}{Mejor}$}
+                \State $Mejor \gets S$
+            \EndIf
+            \State $S \gets \Call{perturb}{S}$
+            \EndWhile
+            \State return $Mejor$
+        \EndFunction
+    \end{algorithmic}
+    }
+
+    ~\
+
+    Como condición de parada se establece que el mínimo de calidad de una solución sea $0.95$ y una cantidad determinada de iteraciones en las cuales la solución no cambie.
 

### 2015-06-24 doc/paper.tex
working on paper

--- 2015-06-23 doc/paper.tex
@@
 
-\clearpage
-
 \section{Evaluación Experimental}
@@
 
-Los datos son obtenidos de KEEL 
-Data-Mining Software Tool [hay que poner el link en referencias].
-En la siguientes tablas se muestran los conjuntos de datos considerados pequeños, medianos y grandes respectivamente.
+Los datos son obtenidos de \emph{KEEL Data-Mining Software Tool}. En la siguientes tablas se muestran los conjuntos de datos considerados pequeños, medianos y grandes respectivamente.
 
@@
 
-\clearpage
+\subsection{Resultados}
 
-\subsection{Resultados}
+\begin{table}[h]
+\scalebox{0.6}{
+\begin{tabular}{ |l|l|l|l|l|l| }
+    \hline
+        \multicolumn{6}{ |c| }{Casos de prueba grandes} \\
+    \hline
+    Caso & Heuristica & Err val & Tam Reduc & Error & Tiempo \\ \hline
+    \multirow{4}{*}{Banana} & Tabu     & \bf{13.28302}   & 3.937526     &  \bf{70.4}   & 53.3222 \\
+                            & ILS      & 13.31699   & \bf{1.305241}     &  70.58  & 591.7742 \\
+                            & Memetico & 14.02265   & 1.47631      &  74.32  & 172.797 \\
+                            & SGA      & 14.49056   & 2.111949     &  76.8   & \bf{35.232}\\ \hline
+\end{tabular}
+}
+\end{table}
 
@@
 \newpage
+\vfill
 
@@
 
-\begin{table}[h]
-\scalebox{0.6}{
-\begin{tabular}{ |l|l|l|l|l|l| }
-    \hline
-        \multicolumn{6}{ |c| }{Casos de prueba grandes} \\
-    \hline
-    Caso & Heuristica & Err val & Tam Reduc & Error & Tiempo \\ \hline
-    \multirow{4}{*}{Banana} & Tabu     & \bf{13.28302}   & 3.937526     &  \bf{70.4}   & 53.3222 \\
-                            & ILS      & 13.31699   & \bf{1.305241}     &  70.58  & 591.7742 \\
-                            & Memetico & 14.02265   & 1.47631      &  74.32  & 172.797 \\
-                            & SGA      & 14.49056   & 2.111949     &  76.8   & \bf{35.232}\\ \hline
-\end{tabular}
-}
-\end{table}
+\clearpage
+\newpage
 
+\begin{figure}[h!]
 \begin{center}
-  \includegraphics[scale=0.4]{val_errors.jpeg}~\\[1cm]
+  \includegraphics[scale=0.4]{banana_error_val.jpeg}~\\[1cm]
 \end{center}
+\end{figure}
 
+\begin{figure}[h!]
 \begin{center}
-  \includegraphics[scale=0.4]{sizes.jpeg}~\\[1cm]
+  \includegraphics[scale=0.4]{banana_tam_reduc.jpeg}~\\[1cm]
 \end{center}
+\end{figure}
 
+\begin{figure}[h!]
+\begin{center}
+  \includegraphics[scale=0.4]{sonar_err_val.jpeg}~\\[1cm]
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+  \includegraphics[scale=0.4]{sonar_tam_red.jpeg}~\\[1cm]
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+  \includegraphics[scale=0.4]{wdbc_err_val.jpeg}~\\[1cm]
+\end{center}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[h!]
+\begin{center}
+  \includegraphics[scale=0.4]{wdbc_tam_red.jpeg}~\\[1cm]
+\end{center}
+\end{figure}
+
+% \newpage
 \section{Conclusiones y recomendaciones}
@@
 
+\bibitem[\protect\citename{Essentials of Metaheuristics}2013]{Sean:01}
+Luke, Sean
+\newblock 2013
+\newblock {\em Essentials Of Metaheuristics}. 
 \end{thebibliography}
